### 2022-09-13 fourmile_dataprocess.Rmd
precip and solar rad prism data ready to be shared

---
title: "Fourmile"
author: "Megan Sears"
date: "7/27/2022"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
editor_options: 
  chunk_output_type: console
---

```{r open package, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(rmdformats)
library(tidyverse)
library(dplyr)
library(lubridate)
library(plotly)
library(ggplot2)
library(here)
library(scales)
library(here)
library(rgdal)
library(sp)
library(sf)
library(mapview)
library(kableExtra)
library(dataRetrieval)
library(raster)
library(sf)
library(stringr)
library(hydroGOF)
library(Metrics)
```

```{r loading data, include=FALSE}
setwd('C:/Users/sears/Documents/Repos/fourmile/data')

#pull in all filenames with composite
filenames <- list.files('.', pattern='.csv', full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 4)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes)

```

## Gage locations

USGS2 and DS2 have the same coordinates. USGS1 and DS3 have the same coordinates.

```{r map, echo=FALSE}
#locations <- read_csv(here('data', 'Outlets.csv'))

#sp <- raster('MA_SP.01.18.tif')

#hrus <- st_read('hrus_final.shp')

#save.image(file = 'spatial.RData')
load('C:/Users/sears/Documents/Repos/fourmile/data/spatial.RData')

locations <- locations %>%
st_as_sf(., coords=c('y','x'), crs = 4326)

mapview(locations, zcol = 'type') + mapview(hrus)

mapview(sp, col.regions = list("brown", 'purple', 'blue', 'white'), at = seq(0, 100, 25)) + 
  mapview(locations, zcol = 'type') +
  mapview(hrus)

```

## Hydrographs

Notes on gages: So far DS2 and DS3 have not been used. US1 and DS1 are mostly point measurements.

```{r hydrographs, echo=F}
#daily streamflow
daily_q <- orun %>%
  dplyr::select(-1) %>%
  filter(!row_number() %in% c(1:9, 11:15)) %>%
  set_names(as.character(slice(., 1))) %>%
  slice(-1) %>%
  replace(.==-9999, NA) %>%
  rename(ds3 = 2,
         usgs1 = 3,
         ds2 = 4,
         usgs2 = 5,
         us1 = 6,
         ds1 = 7) %>%
  mutate(date = mdy(date)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s') %>%
  mutate(q_m3s = as.numeric(q_m3s))

hydro <- ggplot(daily_q, aes(x=date, y=q_m3s, color=site)) + geom_line() +
  ylim(0,8) + theme_bw() + 
  ggtitle('Daily streamflow (all)')

ggplotly(hydro)

#look at 2011-2013
hydro_datelimit <- ggplot(daily_q, aes(x=date, y=q_m3s, color=site)) + geom_line() + 
  theme_bw() + 
  scale_x_date(limits = as.Date(c('2011-01-01', '2014-01-01'))) +
  ylim(0,4) +
  ggtitle('Daily streamflow (2011-2013)')

ggplotly(hydro_datelimit)

```

## Streamflow differencing between each gage

### DS1 minus US1 (uppermost)

```{r DS1-US1, echo=FALSE}

diff_ds1 <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  mutate(ds1_us1 = ds1 - us1) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff1 <- ggplot(diff_ds1, aes(x=date, y=q_m3s)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2011-01-01', '2014-08-01'))) +
  #ylim(-1.1,0.6) +
  theme_bw() + 
  ggtitle('Daily streamflow DS1 - US1')

ggplotly(diff1)

```

#### Average DS1 minus US1 difference

```{r DS1-US1_avg, echo=FALSE}

mean(diff_ds1$q_m3s, na.rm=T)

diff_ds1_yr <- diff_ds1 %>%
  group_by(year(date)) %>%
  summarize(mean(q_m3s, na.rm=T))

kable(diff_ds1_yr, 
      col.names = c("Year", "Avg streamflow difference (m3/s)")) %>%
  kable_styling()

```

### USGS2 minus DS1

```{r USGS2-DS1, echo=FALSE}

diff_usgs2 <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  mutate(usgs2_ds1 = usgs2 - ds1) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff2 <- ggplot(diff_usgs2, aes(x=date, y=q_m3s)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2011-01-01', '2014-08-01'))) +
  #ylim(-1.1,0.6) +
  theme_bw() + 
  ggtitle('Daily streamflow USGS2 - DS1')

ggplotly(diff2)

```

#### Average USGS2 minus DS1 difference

```{r USGS2-DS1_avg, echo=FALSE}

mean(diff_usgs2$q_m3s, na.rm=T)

diff_usgs2_yr <- diff_usgs2 %>%
  group_by(year(date)) %>%
  summarize(mean(q_m3s, na.rm=T))

kable(diff_usgs2_yr, 
      col.names = c("Year", "Avg streamflow difference (m3/s)")) %>%
  kable_styling()

```

### USGS1 minus USGS2

```{r USGS1-USGS2, echo=FALSE}
diff_usgs1 <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  mutate(usgs2_ds1 = usgs1 - usgs2) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff3 <- ggplot(diff_usgs1, aes(x=date, y=q_m3s)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2011-01-01', '2013-12-01'))) +
  #ylim(-1.1,0.6) +
  theme_bw() + 
  ggtitle('Daily streamflow USGS1 - USGS2')

ggplotly(diff3)

```

#### Average USGS1 minus USGS2 difference

```{r USGS1-USGS2_avg, echo=FALSE}

mean(diff_usgs1$q_m3s, na.rm=T)

diff_usgs1_yr <- diff_usgs1 %>%
  group_by(year(date)) %>%
  summarize(mean(q_m3s, na.rm=T))

kable(diff_usgs1_yr, 
      col.names = c("Year", "Avg streamflow difference (m3/s)")) %>%
  kable_styling()

```

#### Sum of losses between USGS1 and USGS2

```{r USGS1-USGS2 sum of losses, echo=FALSE}

sum_diff_usgs1 <- diff_usgs1 %>%
  mutate(month = as.numeric(format(date, "%m")), year = as.numeric(format(date, "%Y"))) %>%
  mutate(neg_q = if_else(q_m3s < 0, q_m3s, 0)) %>%
  dplyr::group_by(year, month) %>%
  dplyr::summarize(mo_loss = sum(q_m3s)) %>%
  #distinct() %>%
  na.omit()

kable(sum_diff_usgs1, 
      col.names = c("Year", "Month", "Monthly sum of losses (m3/s)")) %>%
  kable_styling()

```

### Incorporating pumping

```{r USGS1+pumping-USGS2, echo=FALSE}

inlet <- inle %>%
  replace(.==-9999, NA) %>%
  mutate(date = mdy(date))

diff_usgs1_pump <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  left_join(., inlet, by='date') %>%
  mutate(usgs1_pump = (usgs1 + inflow_m3s),
         usgs1_pump_usgs2 = (usgs1_pump - usgs2)) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1, inflow_m3s)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff4 <- ggplot(diff_usgs1_pump, aes(x=date, y=q_m3s, color=site)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2012-04-01', '2012-10-01'))) +
  ylim(-0.5,0.5) +
  theme_bw() + 
  ggtitle('Daily streamflow: (USGS1 + pumping) - USGS2')

ggplotly(diff4)

```

## Precip

```{r precip, echo=F}
#daily precip
prec_update <- prec %>%
  mutate(date = dmy(date),
         doy = yday(date)) %>%
  mutate_at(c('Gordon_South', 'Gordon_North',
              'GH', 'LM', 'SE', 'SP', 'ST', 'NADP_SL'),
            ~(ifelse(!doy %in% c(152:258), 
                               NA, .))) %>%
  mutate_at(c('prism_W', 'prism_M', 'prism_E'), list(lead), n = 1) %>%
  mutate_at(c('prism_W', 'prism_M', 'prism_E'), ~(ifelse(doy %in% c(152:258),
                                                         NA, .))) %>%
  filter(date > '2009-12-31') %>%
  dplyr::select(-c(doy)) %>%
  mutate(date = format(as.Date(date), '%d.%m.%Y'))

#write_csv(prec_update, 'precipAll_update.csv')

daily_p <- prec %>%
  mutate(date = dmy(date)) %>%
  replace(.==-9999, NA) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'precip_mm') %>%
  mutate(precip_mm = as.numeric(precip_mm)) %>%
  mutate(WY = calcWaterYear(date)) %>%
  group_by(WY, site) %>%
  mutate(dailyp_cum_mm = cumsum(precip_mm))

precip <- ggplot(daily_p, aes(x=date, y=precip_mm, color=site)) + geom_line()  +
  theme_bw() +
  ggtitle('Daily precip (all)')

ggplotly(precip)

precip_cumu <- ggplot(daily_p, aes(x=date, y=dailyp_cum_mm, color=site)) + geom_line()  +
  theme_bw() +
  ggtitle('Daily precip cumulative sums (all)')

ggplotly(precip_cumu)

```

## Compare Niwot SNOTEL precip to prism pixel

### Original prism and Niwot SNOTEL precip comaprison

```{r, echo = F}
niwot_prism <- read.csv(here('data', 'niwot_prism_compare.csv')) %>% 
  filter(!row_number() %in% c(1:10)) %>%
   rename(date = 1,
          prism_mm = 2) %>%
   mutate(date = ymd(date),
          prism_mm = as.numeric(prism_mm))

niwot_snotel <- prec_update %>%
  dplyr::select(c(date, Niwot_Snotel)) %>%
  rename(snotel_mm = Niwot_Snotel) %>%
  mutate(date = dmy(date))

niwot_compare <- full_join(niwot_prism, niwot_snotel, by = 'date')

niwot_compare <- niwot_compare %>%
  mutate(diff_mm = snotel_mm - prism_mm) %>%
  na.omit()

#pbias(niwot_compare$prism_mm, niwot_compare$snotel_mm)

scatter <- ggplot(niwot_compare, aes(x=snotel_mm, y=prism_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism vs. Niwot SNOTEL precip')

ggplotly(scatter)

rmse(niwot_compare$prism_mm, niwot_compare$snotel_mm)
summary(niwot_compare$diff_mm)

#filter to look at sept - june
# niwot_winter_comp <- niwot_compare %>%
#   mutate(month = month(date)) %>%
#   filter(!month %in% c(6:8))
# 
# summary(niwot_winter_comp)
```

### Prism shift 1 day and Niwot SNOTEL precip

```{r, echo = F}
niwot_prismshift <- niwot_compare %>%
  mutate_at(c("prism_mm"), funs(lead), n = 1) %>%
  mutate(month = month(date)) %>%
  #filter(!month %in% c(6:8)) %>%
  mutate(diff_mm = snotel_mm - prism_mm) %>%
  na.omit()

#NSE(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
#pbias(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)

scatter_shiftday <- ggplot(niwot_prismshift, aes(x=snotel_mm, y=prism_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift 1 day versus Niwot SNOTEL precip')

ggplotly(scatter_shiftday)

rmse(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
summary(niwot_prismshift$diff_mm)
```

### Prism shift by fraction and Niwot SNOTEL precip

```{r, echo = F}
niwot_prismfrac <- niwot_compare %>%
  mutate(prism_most_mm = prism_mm * 0.974,
         prism_little_mm = prism_mm * 0.026) %>%
  mutate_at(c('prism_most_mm'), list(lead), n = 1) %>%
  mutate(prism_new_mm = prism_most_mm + prism_little_mm,
         diff_prismfrac = snotel_mm - prism_new_mm) %>%
  na.omit()

scatter_frac <- ggplot(niwot_prismfrac, aes(x=snotel_mm, y=prism_new_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift by 97.4% and Niwot SNOTEL precip')

ggplotly(scatter_frac)

rmse(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
summary(niwot_prismfrac$diff_prismfrac)
#NSE(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
#pbias(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)

```

## Compare prism solar rad to flux solar rad

### Original Prism and Niwot flux tower solar rad

```{r, echo = F}
srad <- solr %>%
  dplyr::select(-1) %>%
  filter(!row_number() %in% c(1:17)) %>%
  replace(.==-9999, NA) %>%
  rename(date = 1,
         prism_w = 2,
         prism_m = 3,
         prism_e = 4,
         gordon_s = 5,
         gordon_n = 6,
         greenlake = 7) %>%
  mutate(date = dmy(date))

niwot_srad <- Niwo %>%
  mutate(date = ymd(date)) %>%
  rename(niwot_flux_mjm2 = 3) %>%
  dplyr::select(-c(watts_per_m2)) %>%
  replace(. == -9999, NA)

## original  
srad_compare <- full_join(srad, niwot_srad, by = 'date') %>%
  dplyr::select(c(date, prism_w, niwot_flux_mjm2)) %>%
  mutate(diff = niwot_flux_mjm2 - prism_w) %>%
  na.omit()

scatter_srad_orig <- ggplot(srad_compare, aes(x=niwot_flux_mjm2, y=prism_w)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism versus Niwot flux tower solar rad')

ggplotly(scatter_srad_orig)

rmse(srad_compare$prism_w, srad_compare$niwot_flux_mjm2)
summary(srad_compare$diff)
```

### Prism shift by 1 day versus Niwot flux tower solar rad

```{r, echo = F}
## 1 day shift
srad_shift1day <- srad_compare %>%
  mutate_at(c("prism_w"),list(lead), n = 1) %>%
  na.omit() %>%
  mutate(diff = niwot_flux_mjm2 - prism_w)

scatter_srad_1day <- ggplot(srad_shift1day, aes(x=niwot_flux_mjm2, y=prism_w)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift 1 day vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_1day)
#scatter_srad_1day

rmse(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
summary(srad_shift1day$diff)
#pbias(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
```

### Prism shift by fraction versus Niwot flux tower solar rad

```{r, echo = F}
## shift by percentage
srad_fracshift <- srad_compare %>%
    mutate(prism_most = prism_w * 0.3,
         prism_little = prism_w * 0.7) %>%
  mutate_at(c('prism_most'), list(lead), n = 1) %>%
  mutate(prism_new = prism_most + prism_little,
         diff_prismfrac = niwot_flux_mjm2 - prism_new) %>%
  na.omit()

scatter_srad_frac <- ggplot(srad_fracshift, aes(x=niwot_flux_mjm2, y=prism_new)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Prism shift by 30% vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_frac)
#scatter_srad_frac

rmse(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
summary(srad_fracshift$diff_prismfrac)
#pbias(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
```

### Daymet (original) versus Niwot flux tower solar rad

```{r, echo = F}
## bring in daymet
## Daily total radiation (MJ/m2/day) can be calculated as follows: ((srad (W/m2) * dayl (s/day)) / l,000,000)
niwot_daymet <- daym %>%
  filter(!row_number() %in% c(1:7)) %>%
  rename(year = 1,
         doy = 2,
         srad_wm2 = 3) %>%
  mutate(date = as.Date(doy),
         month = month(date),
         day = day(date),
         date = make_date(year, month, day),
         daymet_mjm2 = (srad_wm2 * 86400) / 1000000) %>%
  dplyr::select(-c(year, doy, month, day))

niwot_daymet_comp <- full_join(niwot_srad, niwot_daymet, by = 'date') %>%
  dplyr::select(c(date, niwot_flux_mjm2, daymet_mjm2)) %>%
  na.omit() %>%
  mutate(diff = niwot_flux_mjm2 - daymet_mjm2)

scatter_srad_daymet <- ggplot(niwot_daymet_comp, aes(x=niwot_flux_mjm2, y=daymet_mjm2)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
  ggtitle('Daymet vs. Niwot flux tower solar rad')

ggplotly(scatter_srad_daymet)
#scatter_srad_frac

#summary(srad_fracshift)
rmse(niwot_daymet_comp$daymet_mjm2, niwot_daymet_comp$niwot_flux_mjm2)
summary(niwot_daymet_comp$diff)
```