### 2022-09-26 fourmile_inputs.Rmd
just need to finish precip

---
title: "Fourmile_inputs"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
editor_options: 
  chunk_output_type: console
---

```{r open package, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(rmdformats)
library(tidyverse)
library(dplyr)
library(lubridate)
library(plotly)
library(ggplot2)
library(here)
library(scales)
library(here)
library(rgdal)
library(sp)
library(sf)
library(mapview)
library(kableExtra)
library(dataRetrieval)
library(raster)
library(sf)
library(stringr)
library(hydroGOF)
library(Metrics)
library(ranger)
library(missRanger)
library(readxl)

```

```{r loading data, include = F}

setwd('C:/Users/sears/Documents/Repos/fourmile/data/files_to_edit')

#pull in all filenames with csv
filenames <- list.files('.', pattern='.csv', full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 4)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes)

bet_srad <- read_xlsx('bet_srad.xlsx')
updated_prec <- read_xlsx('updated_precip_2016to20.xlsx')
bet_airt <- read_xlsx('bt_Met1.xlsx')
bet_met <-read_xlsx('bt_Met2.xlsx')

```

# RH first

```{r}
#need to bring in Niwot and Betasso and remove prism
rh <- rhum %>%
  mutate(date = dmy(date)) %>%
  dplyr::select(-c(prism_W, prism_M, prism_E))

niwot_rh <- nime %>%
  mutate(date = make_date(Year, Month, Day)) %>%
  dplyr::select(c(date, RH_21m)) %>%
  rename(Niwot = RH_21m) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01')

bet_rh <- bet_met %>%
  rename(rh = 1,
    datetime = 2) %>%
  dplyr::select(c(datetime, rh)) %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(rh = mean(rh)) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01') %>%
  rename(Betasso = rh)

rhall <- full_join(rh, niwot_rh, by = 'date')

rhall <- full_join(rhall, bet_rh, by = 'date') %>% 
  mutate(date = format(as.Date(date), '%d.%m.%Y')) %>%
  replace(is.na(.), -9999)

#rh done
write.csv(rhall, 'rh_use.csv')

```

# wind

```{r}
wind <- wind %>%
  mutate(date = dmy(date))

niwot_wind <- nime %>%
  mutate(date = make_date(Year, Month, Day)) %>%
  dplyr::select(c(date, ws_21m)) %>%
  rename(Niwot = ws_21m) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01')

bet_wind <- bet_met %>%
  rename(windspeed = 3,
    datetime = 4) %>%
  dplyr::select(c(datetime, windspeed)) %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(wind = mean(windspeed)) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01') %>%
  rename(Betasso = wind)

windall <- full_join(wind, niwot_wind, by = 'date')

windall <- full_join(windall, bet_wind, by = 'date') %>% 
  mutate(date = format(as.Date(date), '%d.%m.%Y')) %>%
  replace(is.na(.), -9999)

write.csv(windall, 'wind_use.csv')

```

# Tmax

```{r}
tmax <- tmax %>%
  mutate(date = dmy(date)) %>%
  dplyr::select(-c(prism_W, prism_M, prism_E))

bet_tmax <- bet_airt %>%
    rename(tmax = 3,
    datetime = 4) %>%
  dplyr::select(c(datetime, tmax)) %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(tmax = max(tmax)) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01') %>%
  rename(Betasso = tmax)

tmaxall <- full_join(tmax, bet_tmax, by = 'date') %>%
  mutate(date = format(as.Date(date), '%d.%m.%Y')) %>%
  replace(is.na(.), -9999)

write.csv(tmaxall, 'tmax_use.csv')

```

# Tmin

```{r}
tmin <- tmin %>%
  mutate(date = dmy(date)) %>%
  dplyr::select(-c(prism_W, prism_M, prism_E))

bet_tmin <- bet_airt %>%
    rename(tmin = 1,
    datetime = 2) %>%
  dplyr::select(c(datetime, tmin)) %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(tmin = min(tmin)) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01') %>%
  rename(Betasso = tmin)

tminall <- full_join(tmin, bet_tmin, by = 'date') %>%
  mutate(date = format(as.Date(date), '%d.%m.%Y')) %>%
  replace(is.na(.), -9999)

write.csv(tminall, 'tmin_use.csv')

```

# Solar rad

```{r}
solar <- solr %>%
  mutate(date = dmy(date)) %>%
  dplyr::select(-c(prism_W, prism_M, prism_E))

niwot_solar <- niwo %>%
  dplyr::select(c(1, 3)) %>%
  filter(date > '2009-12-31') %>%
  filter(date < '2021-01-01') %>%
  rename(Niwot = 2)

bet_solar <- bet_srad %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(sw_in_avg_wm2= mean(in_swrad_wm2)) %>%
  filter(date > '2009-12-31') %>%
  mutate(bet_sw_in_mjm2 = (sw_in_avg_wm2 * 86400) / 1000000) %>%
  dplyr::select(-c(sw_in_avg_wm2)) %>%
  rename(Betasso = bet_sw_in_mjm2)

solarall <- full_join(solar, niwot_solar, by = 'date')

solarall <- full_join(solarall, bet_solar, by = 'date') %>% 
  mutate(date = format(as.Date(date), '%d.%m.%Y')) %>%
  replace(is.na(.), -9999)

write.csv(solarall, 'solar_use.csv')

```

# Precip

```{r}

#daily precip
precall <- prec %>%
  mutate(date = dmy(date),
         doy = yday(date)) %>%
  mutate_at(c('Gordon_South', 'Gordon_North',
              'GH', 'LM', 'SE', 'SP', 'ST', 'NADP_SL'),
            ~(ifelse(!doy %in% c(152:258),

  mutate_at(c('prism_W', 'prism_M', 'prism_E'), ~(ifelse(doy %in% c(152:258),
                                                         NA, .))) %>%
  filter(date > '2009-12-31') %>%
  dplyr::select(-c(doy)) %>%
  mutate(date = format(as.Date(date), '%d.%m.%Y'))

  mutate(prism_most_mm = prism_mm * 0.974,
         prism_little_mm = prism_mm * 0.026) %>%
  mutate_at(c('prism_most_mm'), list(lead), n = 1) %>%
  mutate(prism_new_mm = prism_most_mm + prism_little_mm

```