### 2022-10-09 fourmile_dataprocess.Rmd
added runoff ratios by WY

---
title: "Fourmile"
author: "Megan Sears"
date: "7/27/2022"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
editor_options: 
  chunk_output_type: console
---

```{r open package, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(rmdformats)
library(tidyverse)
library(dplyr)
library(lubridate)
library(plotly)
library(ggplot2)
library(here)
library(scales)
library(here)
library(rgdal)
library(sp)
library(sf)
library(mapview)
library(kableExtra)
library(dataRetrieval)
library(raster)
library(sf)
library(stringr)
library(hydroGOF)
library(Metrics)
library(ranger)
library(missRanger)
library(readxl)
```

```{r loading data, include=FALSE}
setwd('C:/Users/sears/Documents/Repos/fourmile/data')

#pull in all filenames with composite
filenames <- list.files('.', pattern='.csv', full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 4)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes)

```

## Gage locations

USGS2 and DS2 have the same coordinates. USGS1 and DS3 have the same coordinates.

```{r map, echo=FALSE}
#locations <- read_csv(here('data', 'Outlets.csv'))

#sp <- raster('MA_SP.01.18.tif')

#hrus <- st_read('hrus_final.shp')

#save.image(file = 'spatial.RData')
load('C:/Users/sears/Documents/Repos/fourmile/data/spatial.RData')

locations <- locations %>%
st_as_sf(., coords=c('y','x'), crs = 4326)

mapview(locations, zcol = 'type') + mapview(hrus)

mapview(sp, col.regions = list("brown", 'purple', 'blue', 'white'), at = seq(0, 100, 25)) + 
  mapview(locations, zcol = 'type') +
  mapview(hrus)

```

## Hydrographs

Notes on gages: So far DS2 and DS3 have not been used. US1 and DS1 are mostly point measurements.

```{r hydrographs, echo=F}
#daily streamflow
daily_q <- orun %>%
  dplyr::select(-1) %>%
  filter(!row_number() %in% c(1:9, 11:15)) %>%
  set_names(as.character(slice(., 1))) %>%
  slice(-1) %>%
  replace(.==-9999, NA) %>%
  rename(ds3 = 2,
         usgs1 = 3,
         ds2 = 4,
         usgs2 = 5,
         us1 = 6,
         ds1 = 7) %>%
  mutate(date = mdy(date)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s') %>%
  mutate(q_m3s = as.numeric(q_m3s))

hydro <- ggplot(daily_q, aes(x=date, y=q_m3s, color=site)) + geom_line() +
  ylim(0,8) + theme_bw() + 
  ggtitle('Daily streamflow (all)')

ggplotly(hydro)

#look at 2011-2013
hydro_datelimit <- ggplot(daily_q, aes(x=date, y=q_m3s, color=site)) + geom_line() + 
  theme_bw() + 
  scale_x_date(limits = as.Date(c('2011-01-01', '2014-01-01'))) +
  ylim(0,4) +
  ggtitle('Daily streamflow (2011-2013)')

ggplotly(hydro_datelimit)

```

## Streamflow differencing between each gage

### DS1 minus US1 (uppermost)

```{r DS1-US1, echo=FALSE}

diff_ds1 <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  mutate(ds1_us1 = ds1 - us1) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff1 <- ggplot(diff_ds1, aes(x=date, y=q_m3s)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2011-01-01', '2014-08-01'))) +
  #ylim(-1.1,0.6) +
  theme_bw() + 
  ggtitle('Daily streamflow DS1 - US1')

ggplotly(diff1)

```

#### Average DS1 minus US1 difference

```{r DS1-US1_avg, echo=FALSE}

mean(diff_ds1$q_m3s, na.rm=T)

diff_ds1_yr <- diff_ds1 %>%
  group_by(year(date)) %>%
  summarize(mean(q_m3s, na.rm=T))

kable(diff_ds1_yr, 
      col.names = c("Year", "Avg streamflow difference (m3/s)")) %>%
  kable_styling()

```

### USGS2 minus DS1

```{r USGS2-DS1, echo=FALSE}

diff_usgs2 <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  mutate(usgs2_ds1 = usgs2 - ds1) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff2 <- ggplot(diff_usgs2, aes(x=date, y=q_m3s)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2011-01-01', '2014-08-01'))) +
  #ylim(-1.1,0.6) +
  theme_bw() + 
  ggtitle('Daily streamflow USGS2 - DS1')

ggplotly(diff2)

```

#### Average USGS2 minus DS1 difference

```{r USGS2-DS1_avg, echo=FALSE}

mean(diff_usgs2$q_m3s, na.rm=T)

diff_usgs2_yr <- diff_usgs2 %>%
  group_by(year(date)) %>%
  summarize(mean(q_m3s, na.rm=T))

kable(diff_usgs2_yr, 
      col.names = c("Year", "Avg streamflow difference (m3/s)")) %>%
  kable_styling()

```

### USGS1 minus USGS2

```{r USGS1-USGS2, echo=FALSE}
diff_usgs1 <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  mutate(usgs2_ds1 = usgs1 - usgs2) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff3 <- ggplot(diff_usgs1, aes(x=date, y=q_m3s)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2011-01-01', '2013-12-01'))) +
  #ylim(-1.1,0.6) +
  theme_bw() + 
  ggtitle('Daily streamflow USGS1 - USGS2')

ggplotly(diff3)

```

#### Average USGS1 minus USGS2 difference

```{r USGS1-USGS2_avg, echo=FALSE}

mean(diff_usgs1$q_m3s, na.rm=T)

diff_usgs1_yr <- diff_usgs1 %>%
  group_by(year(date)) %>%
  summarize(mean(q_m3s, na.rm=T))

kable(diff_usgs1_yr, 
      col.names = c("Year", "Avg streamflow difference (m3/s)")) %>%
  kable_styling()

```

#### Sum of losses between USGS1 and USGS2

```{r USGS1-USGS2 sum of losses, echo=FALSE}

sum_diff_usgs1 <- diff_usgs1 %>%
  mutate(month = as.numeric(format(date, "%m")), year = as.numeric(format(date, "%Y"))) %>%
  mutate(neg_q = if_else(q_m3s < 0, q_m3s, 0)) %>%
  dplyr::group_by(year, month) %>%
  dplyr::summarize(mo_loss = sum(q_m3s)) %>%
  #distinct() %>%
  na.omit()

kable(sum_diff_usgs1, 
      col.names = c("Year", "Month", "Monthly sum of losses (m3/s)")) %>%
  kable_styling()

```

### Incorporating pumping

```{r USGS1+pumping-USGS2, echo=FALSE}

inlet <- inle %>%
  replace(.==-9999, NA) %>%
  mutate(date = mdy(date))

diff_usgs1_pump <- daily_q %>%
  pivot_wider(names_from = site, values_from = q_m3s) %>%
  left_join(., inlet, by='date') %>%
  mutate(usgs1_pump = (usgs1 + inflow_m3s),
         usgs1_pump_usgs2 = (usgs1_pump - usgs2)) %>%
  dplyr::select(-c(usgs2, us1, ds1, ds2, ds3, usgs1, inflow_m3s)) %>%
  pivot_longer(!date, names_to = 'site', values_to = 'q_m3s')

diff4 <- ggplot(diff_usgs1_pump, aes(x=date, y=q_m3s, color=site)) + 
  geom_point() +
  scale_x_date(limits = as.Date(c('2012-04-01', '2012-10-01'))) +
  ylim(-0.5,0.5) +
  theme_bw() + 
  ggtitle('Daily streamflow: (USGS1 + pumping) - USGS2')

ggplotly(diff4)

```

## Precip

```{r precip, echo=F}
#daily precip
# prec_update <- prec %>%
#   mutate(date = dmy(date),
#          doy = yday(date)) %>%
#   mutate_at(c('Gordon_South', 'Gordon_North',
#               'GH', 'LM', 'SE', 'SP', 'ST', 'NADP_SL'),
#             ~(ifelse(!doy %in% c(152:258), 
#                                NA, .))) %>%
#   mutate_at(c('prism_W', 'prism_M', 'prism_E'), list(lead), n = 1) %>%
#   mutate_at(c('prism_W', 'prism_M', 'prism_E'), ~(ifelse(doy %in% c(152:258),
#                                                          NA, .))) %>%
#   filter(date > '2009-12-31') %>%
#   dplyr::select(-c(doy)) %>%
#   mutate(date = format(as.Date(date), '%d.%m.%Y'))
# 
# #write_csv(prec_update, 'precipAll_update.csv')
# 
# daily_p <- prec %>%
#   mutate(date = dmy(date)) %>%
#   replace(.==-9999, NA) %>%
#   pivot_longer(!date, names_to = 'site', values_to = 'precip_mm') %>%
#   mutate(precip_mm = as.numeric(precip_mm)) %>%
#   mutate(WY = calcWaterYear(date)) %>%
#   group_by(WY, site) %>%
#   mutate(dailyp_cum_mm = cumsum(precip_mm))
# 
# precip <- ggplot(daily_p, aes(x=date, y=precip_mm, color=site)) + geom_line()  +
#   theme_bw() +
#   ggtitle('Daily precip (all)')
# 
# ggplotly(precip)
# 
# precip_cumu <- ggplot(daily_p, aes(x=date, y=dailyp_cum_mm, color=site)) + geom_line()  +
#   theme_bw() +
#   ggtitle('Daily precip cumulative sums (all)')
# 
# ggplotly(precip_cumu)

```

## Compare Niwot SNOTEL precip to prism pixel

### Original prism and Niwot SNOTEL precip comaprison

```{r, echo = F}
# niwot_prism <- read.csv(here('data', 'niwot_prism_compare.csv')) %>% 
#   filter(!row_number() %in% c(1:10)) %>%
#    rename(date = 1,
#           prism_mm = 2) %>%
#    mutate(date = ymd(date),
#           prism_mm = as.numeric(prism_mm))
# 
# niwot_snotel <- prec_update %>%
#   dplyr::select(c(date, Niwot_Snotel)) %>%
#   rename(snotel_mm = Niwot_Snotel) %>%
#   mutate(date = dmy(date))
# 
# niwot_compare <- full_join(niwot_prism, niwot_snotel, by = 'date')
# 
# niwot_compare <- niwot_compare %>%
#   mutate(diff_mm = snotel_mm - prism_mm) %>%
#   na.omit()
# 
# #pbias(niwot_compare$prism_mm, niwot_compare$snotel_mm)
# 
# scatter <- ggplot(niwot_compare, aes(x=snotel_mm, y=prism_mm)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism vs. Niwot SNOTEL precip')
#   
# ggplotly(scatter)
# 
# rmse(niwot_compare$prism_mm, niwot_compare$snotel_mm)
# bias(niwot_compare$snotel_mm, niwot_compare$prism_mm)
# summary(niwot_compare$diff_mm)
# #filter to look at sept - june
# niwot_winter_comp <- niwot_compare %>%
#   mutate(month = month(date)) %>%
#   filter(!month %in% c(6:8))
# 
# summary(niwot_winter_comp)
```

### Prism shift 1 day and Niwot SNOTEL precip

```{r, echo = F}
# niwot_prismshift <- niwot_compare %>%
#   mutate_at(c("prism_mm"), funs(lead), n = 1) %>%
#   mutate(month = month(date)) %>%
#   #filter(!month %in% c(6:8)) %>%
#   mutate(diff_mm = snotel_mm - prism_mm) %>%
#   na.omit()
# 
# #NSE(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
# #pbias(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
# 
# scatter_shiftday <- ggplot(niwot_prismshift, aes(x=snotel_mm, y=prism_mm)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift 1 day versus Niwot SNOTEL precip')
#   
# ggplotly(scatter_shiftday)
# 
# rmse(niwot_prismshift$prism_mm, niwot_prismshift$snotel_mm)
# bias(niwot_prismshift$snotel_mm, niwot_prismshift$prism_mm)
# summary(niwot_prismshift$diff_mm)
```

### Prism shift by fraction and Niwot SNOTEL precip

```{r, echo = F}
# niwot_prismfrac <- niwot_compare %>%
#   mutate(prism_most_mm = prism_mm * 0.974,
#          prism_little_mm = prism_mm * 0.026) %>%
#   mutate_at(c('prism_most_mm'), list(lead), n = 1) %>%
#   mutate(prism_new_mm = prism_most_mm + prism_little_mm,
#          diff_prismfrac = snotel_mm - prism_new_mm) %>%
#   na.omit()
# 
# scatter_frac <- ggplot(niwot_prismfrac, aes(x=snotel_mm, y=prism_new_mm)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift by 97.4% and Niwot SNOTEL precip')
#   
# ggplotly(scatter_frac)
# 
# rmse(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
# bias(niwot_prismfrac$snotel_mm, niwot_prismfrac$prism_new_mm)
# summary(niwot_prismfrac$diff_prismfrac)
#NSE(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)
#pbias(niwot_prismfrac$prism_new_mm, niwot_prismfrac$snotel_mm)

```

## Compare prism solar rad to flux solar rad

### Original Prism and Niwot flux tower solar rad

```{r, echo = F}
# srad <- solr %>%
#   dplyr::select(-1) %>%
#   filter(!row_number() %in% c(1:17)) %>%
#   replace(.==-9999, NA) %>%
#   rename(date = 1,
#          prism_w = 2,
#          prism_m = 3,
#          prism_e = 4,
#          gordon_s = 5,
#          gordon_n = 6,
#          greenlake = 7) %>%
#   mutate(date = dmy(date))
# 
# niwot_srad <- Niwo %>%
#   mutate(date = ymd(date)) %>%
#   rename(niwot_flux_mjm2 = 3) %>%
#   dplyr::select(-c(watts_per_m2)) %>%
#   replace(. == -9999, NA)
# 
# ## original  
# srad_compare <- full_join(srad, niwot_srad, by = 'date') %>%
#   dplyr::select(c(date, prism_w, niwot_flux_mjm2)) %>%
#   mutate(diff = niwot_flux_mjm2 - prism_w) %>%
#   na.omit()
# 
# scatter_srad_orig <- ggplot(srad_compare, aes(x=niwot_flux_mjm2, y=prism_w)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism versus Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_orig)
# 
# rmse(srad_compare$prism_w, srad_compare$niwot_flux_mjm2)
# bias(srad_compare$niwot_flux_mjm2, srad_compare$prism_w)
# summary(srad_compare$diff)
```

### Prism shift by 1 day versus Niwot flux tower solar rad

```{r, echo = F}
## 1 day shift
# srad_shift1day <- srad_compare %>%
#   mutate_at(c("prism_w"),list(lead), n = 1) %>%
#   na.omit() %>%
#   mutate(diff = niwot_flux_mjm2 - prism_w)
# 
# scatter_srad_1day <- ggplot(srad_shift1day, aes(x=niwot_flux_mjm2, y=prism_w)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift 1 day vs. Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_1day)
# #scatter_srad_1day
# 
# rmse(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
# bias(srad_shift1day$niwot_flux_mjm2, srad_shift1day$prism_w)
# summary(srad_shift1day$diff)
#pbias(srad_shift1day$prism_w, srad_shift1day$niwot_flux_mjm2)
```

### Prism shift by fraction versus Niwot flux tower solar rad

```{r, echo = F}
## shift by percentage
# srad_fracshift <- srad_compare %>%
#     mutate(prism_most = prism_w * 0.3,
#          prism_little = prism_w * 0.7) %>%
#   mutate_at(c('prism_most'), list(lead), n = 1) %>%
#   mutate(prism_new = prism_most + prism_little,
#          diff_prismfrac = niwot_flux_mjm2 - prism_new) %>%
#   na.omit()
#   
# scatter_srad_frac <- ggplot(srad_fracshift, aes(x=niwot_flux_mjm2, y=prism_new)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Prism shift by 30% vs. Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_frac)
# #scatter_srad_frac
# 
# rmse(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
# bias(srad_fracshift$niwot_flux_mjm2, srad_fracshift$prism_new)
# summary(srad_fracshift$diff_prismfrac)
#pbias(srad_fracshift$prism_new, srad_fracshift$niwot_flux_mjm2)
```

### Daymet (original) versus Niwot flux tower solar rad

```{r, echo = F}
## bring in daymet
## Daily total radiation (MJ/m2/day) can be calculated as follows: ((srad (W/m2) * dayl (s/day)) / l,000,000)
# niwot_daymet <- daym %>%
#   filter(!row_number() %in% c(1:7)) %>%
#   rename(year = 1,
#          doy = 2,
#          srad_wm2 = 3) %>%
#   mutate(date = as.Date(doy),
#          month = month(date),
#          day = day(date),
#          date = make_date(year, month, day),
#          daymet_mjm2 = (srad_wm2 * 86400) / 1000000) %>%
#   dplyr::select(-c(year, doy, month, day))
#   
# niwot_daymet_comp <- full_join(niwot_srad, niwot_daymet, by = 'date') %>%
#   dplyr::select(c(date, niwot_flux_mjm2, daymet_mjm2)) %>%
#   na.omit() %>%
#   mutate(diff = niwot_flux_mjm2 - daymet_mjm2)
#   
# scatter_srad_daymet <- ggplot(niwot_daymet_comp, aes(x=niwot_flux_mjm2, y=daymet_mjm2)) +
#   geom_point() +
#   geom_abline(intercept = 0, slope = 1, size=1.5, color="blue") +
#   ggtitle('Daymet vs. Niwot flux tower solar rad')
#   
# ggplotly(scatter_srad_daymet)
# #scatter_srad_frac
# 
# #summary(srad_fracshift)
# rmse(niwot_daymet_comp$daymet_mjm2, niwot_daymet_comp$niwot_flux_mjm2)
# bias(niwot_daymet_comp$niwot_flux_mjm2, niwot_daymet_comp$daymet_mjm2)
# summary(niwot_daymet_comp$diff)
```

### Solar rad gap filling

```{r, echo=F}
#bring in srad and niwot_srad
# srad_gapfil <- full_join(srad, niwot_srad) %>%
#   filter(date > '2009-12-31') %>%
#   dplyr::select(-c(prism_w, prism_m, prism_e))
# 
# #bat <- read_xlsx('bat_srad.xlsx')
# #save.image(file = 'bat.RData')
# load('C:/Users/sears/Documents/Repos/fourmile/data/bat.RData')
# 
# bet_day <- bat %>%
#   mutate(date = as.Date(datetime)) %>%
#   group_by(date) %>%
#   summarize(sw_in_avg_wm2= mean(in_swrad_wm2)) %>%
#   filter(date > '2009-12-31') %>%
#   mutate(bet_sw_in_mjm2 = (sw_in_avg_wm2 * 86400) / 1000000) %>%
#   dplyr::select(-c(sw_in_avg_wm2))
# 
# srad_gapfil_all <- full_join(srad_gapfil, bet_day, by='date') %>%
#   filter(date < '2020-12-31') %>%
#   dplyr::select(-c(bet_sw_in_mjm2.x)) %>%
#   rename(bet_sw = bet_sw_in_mjm2.y) %>%
#   mutate(doyh = yday(date))
# 
# summary(srad_gapfil_all)
# 
# srad_flip_nofill <- srad_gapfil %>%
#   pivot_longer(!date, names_to = 'site', values_to = 'sw_in_mjm2')
# 
# srad_nofill <- ggplot(srad_flip_nofill, aes(x=date, y=sw_in_mjm2, color=site)) +
#   geom_line() +
#   ggtitle('Solar rad, no fill')
# 
# ggplotly(srad_nofill)
# 
# new_srad <- missRanger(srad_gapfil_all, num.trees=100000)
# summary(new_srad)
# 
# srad_flip_fill <- new_srad %>%
#   dplyr::select(-c(doyh)) %>% 
#   pivot_longer(!date, names_to = 'site', values_to = 'sw_in_mjm2')
#   
# 
# srad_filled <- ggplot(srad_flip_fill, aes(x=date, y=sw_in_mjm2, color=site)) +
#   geom_line() +
#   ggtitle('Solar rad, fill')
# 
# ggplotly(srad_filled)

```

## Runoff ratios

```{r}
prec_runoff <- read_csv('C:/Users/sears/Documents/Repos/fourmile/data/prec_runoff.csv')

ratios_prep <- prec_runoff %>%
  mutate(prec_12 = hru5 + hru6,
         prec_13 = hru3 + hru11,
         prec_10 = hru1 + hru2 + hru8 + hru9 + hru4,
         prec_11 = hru7 + hru10,
         date = dmy(time)) %>%
  dplyr::select(c(date, prec_12, prec_13, prec_10, prec_11)) %>%
  mutate_at(c('prec_12', 'prec_13', 'prec_10', 'prec_11'), ~(./1000)) %>%
  mutate(prec_12_cmd = prec_12 * (14378900+16275100),
         prec_13_cmd = prec_13 * (6850100+4299900),
         prec_10_cmd = prec_10 * (4092400+2170400+2379900+1725600+1967700),
         prec_11_cmd = prec_11 * (7092000+5834400)) %>%
  dplyr::select(-c(prec_12, prec_13, prec_10, prec_11))

obs_q <- read_csv('C:/Users/sears/Documents/Repos/fourmile/data/orun.csv')

obs_q <- obs_q %>% 
  dplyr::select(-1) %>%
  filter(!row_number() %in% c(1:9, 11:15)) %>%
  set_names(as.character(slice(., 1))) %>%
  slice(-1) %>%
  replace(.==-9999, NA) %>%
  rename(ds3 = 2,
         usgs1 = 3,
         ds2 = 4,
         usgs2 = 5,
         us1 = 6,
         ds1 = 7) %>%
  mutate(date = mdy(date)) %>%
  mutate_if(is.character, as.numeric) %>%
  dplyr::select(-c(ds2, ds3)) %>%
  mutate(across(where(is.numeric), ~(.*86400))) %>%
  filter(!between(date, as.Date("2013-09-10"), as.Date("2014-04-01")))

ratios <- full_join(ratios_prep, obs_q, by = 'date')

summary(ratios)

ratios_wy <- ratios %>%
  mutate(wy = calcWaterYear(date)) %>%
  group_by(wy) %>%
  summarize_if(is.numeric, sum, na.rm=TRUE) %>%
  slice(-c(11:12)) %>%
  mutate(us1_rr = us1 / prec_12_cmd,
         ds1_rr = ds1 / prec_13_cmd,
         usgs2_rr = usgs2 / prec_10_cmd,
         usgs1_rr = usgs1 / prec_11_cmd)

kable(ratios_wy) %>%
  kable_styling()

ratios_2014 <- ratios %>%
  mutate(wy = calcWaterYear(date)) %>%
  filter(wy == 2014) %>%
  mutate(usgs1_rr = usgs1 / prec_11_cmd)

```