### 2023-04-25 38HRU_model/fourmile_model_outputs.Rmd
messed up

---
title: "Fourmile Calibrations and Output"
author: "Megan Sears"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F, 
  warning  = F)

library(tidyverse)
library(plotly)
library(lubridate)
library(knitr)
library(kableExtra)
library(dataRetrieval)
library(hydroGOF)
library(sp)
library(raster)
library(sf)
library(mapview)
library(here)
library(scales)
library(dataRetrieval)
library(tmap)
library(ggpubr)
library(gridExtra)

```

```{r load in files, include = F}

kge <- read_csv('calruns_KGE.csv')

```

# KGE for calibrations

```{r look at kge, echo = F}

kge_plot <- ggplot(kge, aes(year_missing, KGE, shape = input_data, color = input_data)) + geom_point(size = 3) +
  theme_bw()

ggplotly(kge_plot)

```

# Parameters

## Scatter plots

```{r read in params, echo = FALSE, message = FALSE, warning = FALSE, include = F}
setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_model/cals")

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 9)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

```

```{r params, echo = FALSE, message = FALSE, warning = FALSE}

# function to read files, remove top rows, paste site name
read_cals <- function(df, name){
    df <- df %>%
      slice(-c(1:9, 146:277)) %>%
      separate(Parameter, c("Parameter", "value"), sep = ",") %>%
      drop_na() %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)

}

grid_2011 <- read_cals(grid2011., name = 'grid_2011')
grid_2012 <- read_cals(grid2012., name = 'grid_2012')
grid_2013 <- read_cals(grid2013., name = 'grid_2013')
grid_2014 <- read_cals(grid2014., name = 'grid_2014')
grid_2015 <- read_cals(grid2015., name = 'grid_2015')
grid_2016 <- read_cals(grid2016., name = 'grid_2016')
grid_2017 <- read_cals(grid2017., name = 'grid_2017')
grid_2018 <- read_cals(grid2018., name = 'grid_2018')
grid_2019 <- read_cals(grid2019., name = 'grid_2019')
grid_2020 <- read_cals(grid2020., name = 'grid_2020')
grid_all <- read_cals(gridall.c, name = 'grid_all')

grid <- bind_rows(grid_2011, grid_2012, grid_2013, grid_2014, grid_2015,
                  grid_2016, grid_2017, grid_2018, grid_2019, grid_2020,
                  grid_all) %>%
  mutate(input_data = 'grid') %>%
  mutate(year_missing_almost = gsub('grid_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost)) %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

statn_2011 <- read_cals(statn2011, name = 'statn_2011')
statn_2012 <- read_cals(statn2012, name = 'statn_2012')
statn_2013 <- read_cals(statn2013, name = 'statn_2013')
statn_2014 <- read_cals(statn2014, name = 'statn_2014')
statn_2015 <- read_cals(statn2015, name = 'statn_2015')
statn_2016 <- read_cals(statn2016, name = 'statn_2016')
statn_2017 <- read_cals(statn2017, name = 'statn_2017')
statn_2018 <- read_cals(statn2018, name = 'statn_2018')
statn_2019 <- read_cals(statn2019, name = 'statn_2019')
statn_2020 <- read_cals(statn2020, name = 'statn_2020')
statn_all <- read_cals(statnall., name = 'statn_all')

statn <- bind_rows(statn_2011, statn_2012, statn_2013, statn_2014, statn_2015,
                  statn_2016, statn_2017, statn_2018, statn_2019, statn_2020,
                  statn_all) %>%
  mutate(input_data = 'statn') %>%
  mutate(year_missing_almost = gsub('statn_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost)) %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

params <- bind_rows(statn, grid)

params <- params %>%
  filter(!param %in% c('angstrom_a', 'angstrom_b', 
                      'Beta_min', 
                      'Beta_rsd', "Beta_trans", 'clat', 
                      'delayNitrification',
                      'denitrificationRateCoefficient', 
                      'denitrificationSoilSaturationThreshold',
                      'deposition_factor', 'drrad', 
                      'drspac', 'initLPS', 'initRG1',
                      'initRG2', 'Ksink', 'LExCoef', 
                      'longTZ', 'N_concRG1', 'N_concRG2', 'N_delay_RG1',
                      'N_delay_RG2', 
                      'nitrificationSoilTemperatureThreshold', 
                      'nitrificationSurfaceTemperatureThreshold',
                      'opti', 'piadin', 'rootfactor', 'sceno', 
                      'soilImpGT80', 'soilImpLT80',
                      'soilLinRed', 'soilMaxPerc', 'temp_lag')) %>%
    mutate(value = as.numeric(value),
         param = as.factor(param))

#function to plot all params
plot_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = year_missing,
               y = value,
               shape = input_data, 
               color = input_data)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plot_params)

#write.csv(params, 'params.csv')

```

## Density plots

```{r density plots, echo = F, message = F}

mean_params <- params %>%
  group_by(param, input_data) %>%
  summarize(mean = mean(value))

plotdens_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = value,
               fill = input_data, 
               color = input_data)) + 
    geom_density(alpha=0.5) +
    geom_vline(data=filter(mean_params, param == param_name), 
               aes(xintercept=mean, color=input_data),
             linetype="dashed", linewidth=0.75) +
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plotdens_params)

```

<!-- # Outlet results -->

<!-- ## Hydrograph -->

<!-- ```{r hydrographs, echo = F} -->
<!-- flow <- read_csv('discharge_all.csv') %>% -->
<!--   replace(.==-9999, NA) -->

<!-- flow_pivot <- flow %>% -->
<!--   pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>% -->
<!--   mutate(year = as.factor(year(date))) -->

<!-- q <- ggplot(flow_pivot, aes(date, q_m3s, color = cal_run)) + -->
<!--   geom_line() + -->
<!--   scale_color_manual(name="cal_run", values=c('obs' = 'black', -->
<!--                                             "grid_2011"="seagreen", -->
<!--                                             "grid_2012"="seagreen2", -->
<!--                                             "grid_2013"="steelblue", -->
<!--                                             "grid_2014"='dodgerblue', -->
<!--                                             "grid_2015" = "cyan3", -->
<!--                                             'grid_2016' = 'blue', -->
<!--                                             'grid_2017' = 'blue4', -->
<!--                                             'grid_2018' = 'cadetblue', -->
<!--                                             'grid_2019' = 'lightgreen', -->
<!--                                             'grid_2020' = 'olivedrab', -->
<!--                                             'grid_all' = 'steelblue3', -->
<!--                                             "statn_2011"="maroon", -->
<!--                                             "statn_2012"="deeppink4", -->
<!--                                             "statn_2013"="deeppink", -->
<!--                                             "statn_2014"='purple', -->
<!--                                             "statn_2015" = "mediumorchid1", -->
<!--                                             'statn_2016' = 'indianred', -->
<!--                                             'statn_2017' = 'sienna', -->
<!--                                             'statn_2018' = 'salmon2', -->
<!--                                             'statn_2019' = 'mediumpurple1', -->
<!--                                             'statn_2020' = 'violetred', -->
<!--                                             'statn_all' = 'red')) -->

<!-- ggplotly(q) -->

<!-- ``` -->

<!-- ## Annual hydrographs -->

<!-- ```{r cy hydrographs, echo = F} -->

<!-- plotq_cy <- function(calyear) { -->
<!--   {flow_pivot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = date, -->
<!--                y = q_m3s, -->
<!--                color = cal_run)) + -->
<!--     geom_line() + -->
<!--         scale_color_manual(name="cal_run", values=c('obs' = 'black', -->
<!--                                             "grid_2011"="seagreen",  -->
<!--                                             "grid_2012"="seagreen2",  -->
<!--                                             "grid_2013"="steelblue", -->
<!--                                             "grid_2014"='dodgerblue',  -->
<!--                                             "grid_2015" = "cyan3", -->
<!--                                             'grid_2016' = 'blue', -->
<!--                                             'grid_2017' = 'blue4', -->
<!--                                             'grid_2018' = 'cadetblue', -->
<!--                                             'grid_2019' = 'lightgreen', -->
<!--                                             'grid_2020' = 'olivedrab', -->
<!--                                             'grid_all' = 'steelblue3', -->
<!--                                             "statn_2011"="maroon",  -->
<!--                                             "statn_2012"="deeppink4",  -->
<!--                                             "statn_2013"="deeppink", -->
<!--                                             "statn_2014"='purple',  -->
<!--                                             "statn_2015" = "mediumorchid1", -->
<!--                                             'statn_2016' = 'indianred', -->
<!--                                             'statn_2017' = 'sienna', -->
<!--                                             'statn_2018' = 'salmon2', -->
<!--                                             'statn_2019' = 'mediumpurple1', -->
<!--                                             'statn_2020' = 'violetred', -->
<!--                                             'statn_all' = 'red')) + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw()} %>% -->
<!--     ggplotly() -->

<!-- } -->

<!-- name <- levels(flow_pivot$year) -->

<!-- try <- lapply(name, plotq_cy) -->

<!-- htmltools::tagList(try) -->

<!-- try -->
<!-- ``` -->

<!-- ## KGE -->

<!-- ### Overall KGE -->

<!-- ```{r kge overall, echo = F} -->

<!-- flow_kge_all <- flow %>% -->
<!--   summarize(grid2011 = KGE(grid_2011, obs), -->
<!--             grid2012 = KGE(grid_2012, obs), -->
<!--             grid2013 = KGE(grid_2013, obs), -->
<!--             grid2014 = KGE(grid_2014, obs), -->
<!--             grid2015 = KGE(grid_2015, obs), -->
<!--             grid2016 = KGE(grid_2016, obs), -->
<!--             grid2017 = KGE(grid_2017, obs), -->
<!--             grid2018 = KGE(grid_2018, obs), -->
<!--             grid2019 = KGE(grid_2019, obs), -->
<!--             grid2020 = KGE(grid_2020, obs), -->
<!--             gridall = KGE(grid_all, obs), -->
<!--             statn2011 = KGE(statn_2011, obs), -->
<!--             statn2012 = KGE(statn_2012, obs), -->
<!--             statn2013 = KGE(statn_2013, obs), -->
<!--             statn2014 = KGE(statn_2014, obs), -->
<!--             statn2015 = KGE(statn_2015, obs), -->
<!--             statn2016 = KGE(statn_2016, obs), -->
<!--             statn2017 = KGE(statn_2017, obs), -->
<!--             statn2018 = KGE(statn_2018, obs), -->
<!--             statn2019 = KGE(statn_2019, obs), -->
<!--             statn2020 = KGE(statn_2020, obs), -->
<!--             statnall = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- flow_kge_all <- flow_kge_all %>% -->
<!--   pivot_longer(everything()) %>% -->
<!--   arrange(value) -->

<!-- kable(flow_kge_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### KGE by year -->

<!-- ```{r kge by year, echo = F} -->

<!-- flow_year <- flow %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid2011 = KGE(grid_2011, obs), -->
<!--             grid2012 = KGE(grid_2012, obs), -->
<!--             grid2013 = KGE(grid_2013, obs), -->
<!--             grid2014 = KGE(grid_2014, obs), -->
<!--             grid2015 = KGE(grid_2015, obs), -->
<!--             grid2016 = KGE(grid_2016, obs), -->
<!--             grid2017 = KGE(grid_2017, obs), -->
<!--             grid2018 = KGE(grid_2018, obs), -->
<!--             grid2019 = KGE(grid_2019, obs), -->
<!--             grid2020 = KGE(grid_2020, obs), -->
<!--             gridall = KGE(grid_all, obs), -->
<!--             statn2011 = KGE(statn_2011, obs), -->
<!--             statn2012 = KGE(statn_2012, obs), -->
<!--             statn2013 = KGE(statn_2013, obs), -->
<!--             statn2014 = KGE(statn_2014, obs), -->
<!--             statn2015 = KGE(statn_2015, obs), -->
<!--             statn2016 = KGE(statn_2016, obs), -->
<!--             statn2017 = KGE(statn_2017, obs), -->
<!--             statn2018 = KGE(statn_2018, obs), -->
<!--             statn2019 = KGE(statn_2019, obs), -->
<!--             statn2020 = KGE(statn_2020, obs), -->
<!--             statnall = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- flow_kge <- flow_year %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'kge') -->

<!-- flow_kge %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- flow_plot <- flow_year %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- kge_plot_cy <- ggplot(flow_plot, aes(year, kge, color = model_run)) +  -->
<!--   geom_point(size = 3) + -->
<!--   scale_color_manual(name="model_run", values=c("grid2011"="seagreen",  -->
<!--                                             "grid2012"="seagreen2",  -->
<!--                                             "grid2013"="steelblue", -->
<!--                                             "grid2014"='dodgerblue',  -->
<!--                                             "grid2015" = "cyan3", -->
<!--                                             'grid2016' = 'blue', -->
<!--                                             'grid2017' = 'blue4', -->
<!--                                             'grid2018' = 'cadetblue', -->
<!--                                             'grid2019' = 'lightgreen', -->
<!--                                             'grid2020' = 'olivedrab', -->
<!--                                             'gridall' = 'steelblue3', -->
<!--                                             "statn2011"="maroon",  -->
<!--                                             "statn2012"="deeppink4",  -->
<!--                                             "statn2013"="deeppink", -->
<!--                                             "statn2014"='purple',  -->
<!--                                             "statn2015" = "mediumorchid1", -->
<!--                                             'statn2016' = 'indianred', -->
<!--                                             'statn2017' = 'sienna', -->
<!--                                             'statn2018' = 'salmon2', -->
<!--                                             'statn2019' = 'mediumpurple1', -->
<!--                                             'statn2020' = 'violetred', -->
<!--                                             'statnall' = 'red')) + -->
<!--   theme_bw() -->

<!-- ggplotly(kge_plot_cy) -->

<!-- ``` -->

<!-- ### Annual KGE plots -->

<!-- ```{r wy plots, echo = F} -->

<!-- plot_kge <- function(calyear) { -->
<!--   flow_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = kge)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(flow_plot$year) -->

<!-- lapply(name, plot_kge) -->

<!-- ``` -->

<!-- # Ensembles -->

<!-- ## Hydrograph -->

<!-- ```{r boulder cr gage, echo = F} -->

<!-- #pull in Boulder Creek @ N 75th gage -->

<!-- bouldcr <- readNWISdv(siteNumbers = "06730200", -->
<!--                           parameterCd = "00060",  -->
<!--                           startDate = "2011-01-01",  -->
<!--                           endDate = "2020-12-31")  %>% -->
<!--   dplyr::select(date = Date, -->
<!--          q_cfs = X_00060_00003) %>% -->
<!--   mutate(bouldcr = q_cfs/35.3146662) %>% -->
<!--   dplyr::select(-q_cfs) %>% -->
<!--   mutate(year = year(date)) -->

<!-- ``` -->

<!-- ```{r ensemble hydrographs, echo = F} -->

<!-- ensem <- flow %>% -->
<!--   rowwise() %>% -->
<!--   mutate(grid_ensemble = mean(c(grid_2011, grid_2012, -->
<!--                                 grid_2013, grid_2014, -->
<!--                                 grid_2015, grid_2016, -->
<!--                                 grid_2017, grid_2018, -->
<!--                                 grid_2019, grid_2020)), -->
<!--          statn_ensemble = mean(c(statn_2011, statn_2012, -->
<!--                                  statn_2013, statn_2014, -->
<!--                                  statn_2015, statn_2016, -->
<!--                                  statn_2017, statn_2018, -->
<!--                                  statn_2019, statn_2020)), -->
<!--          all_ensemble = mean(c(grid_2011, grid_2012, -->
<!--                                grid_2013, grid_2014, -->
<!--                                grid_2015, grid_2016, -->
<!--                                grid_2017, grid_2018, -->
<!--                                grid_2019, grid_2020, -->
<!--                                statn_2011, statn_2012, -->
<!--                                statn_2013, statn_2014, -->
<!--                                statn_2015, statn_2016, -->
<!--                                statn_2017, statn_2018, -->
<!--                                statn_2019, statn_2020, -->
<!--                                grid_all, statn_all))) %>% -->
<!--   dplyr::select(date, grid_ensemble, statn_ensemble,  -->
<!--          all_ensemble, grid_all,  -->
<!--          statn_all, obs)  -->

<!-- ensem_pivot <- ensem %>% -->
<!--   pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>% -->
<!--   mutate(year = as.factor(year(date))) -->

<!-- q_ensemble <- ggplot(ensem_pivot, aes(date, q_m3s, color = cal_run)) + -->
<!--   geom_line() + -->
<!--   scale_color_manual(name="cal_run", values=c('obs'='black', -->
<!--                                             "grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   geom_line(data = bouldcr, aes(date, bouldcr/10), color = 'darkgrey') + -->
<!--   scale_y_continuous(name = 'Q_m3s', -->
<!--                     sec.axis = sec_axis(~.*10, name='BoulderCr_cms')) + -->
<!--   theme_bw() -->

<!-- ggplotly(q_ensemble) %>% -->
<!--   add_lines(data=bouldcr, x=~date, y=~bouldcr, colors=NULL, yaxis="y2",  -->
<!--               inherit=FALSE, showlegend = FALSE) %>% -->
<!--   layout(yaxis2 = list(overlaying = "y", side = "right", -->
<!--                        color = 'darkgray', -->
<!--                        title = "BoulderCr_cms"), -->
<!--          legend = list(x = 1.1, y = 0.95)) -->

<!-- ``` -->

<!-- ## Annual hydrographs -->

<!-- ```{r hydrogs by cy, echo = F} -->

<!-- plotqensem_cy <- function(calyear) { -->
<!--   {ensem_pivot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = date, -->
<!--                y = q_m3s, -->
<!--                color = cal_run)) + -->
<!--     geom_line() + -->
<!--     scale_color_manual(name="cal_run", values=c('obs' = 'black', -->
<!--                                             "grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() -->
<!--     } %>% -->
<!--     ggplotly() -->
<!-- } -->

<!-- name <- levels(ensem_pivot$year) -->

<!-- try <- lapply(name, plotqensem_cy) -->

<!-- htmltools::tagList(try) -->

<!-- try -->

<!-- ``` -->

<!-- ## KGE -->

<!-- ### KGE for ensembles (entire period) -->

<!-- ```{r ensemble kge, echo = F} -->

<!-- ensem_kge_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = KGE(grid_ensemble, obs), -->
<!--             statn_ensemble = KGE(statn_ensemble, obs), -->
<!--             all_ensemble = KGE(all_ensemble, obs), -->
<!--             grid_all = KGE(grid_all, obs), -->
<!--             statn_all = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- kable(ensem_kge_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual KGEs for ensembles -->

<!-- ```{r annual kge ensemble, echo = F} -->

<!-- yr_kge_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = KGE(grid_ensemble, obs), -->
<!--             statn_ensemble = KGE(statn_ensemble, obs), -->
<!--             all_ensemble = KGE(all_ensemble, obs), -->
<!--             grid_all = KGE(grid_all, obs), -->
<!--             statn_all = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- yr_kge_table <- yr_kge_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'kge') -->

<!-- yr_kge_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_kge_plot <- yr_kge_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual kge vals for ensembles -->
<!-- plot_kge_yr_ensem <- function(calyear) { -->
<!--   yr_kge_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = kge)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_kge_plot$year) -->

<!-- lapply(name, plot_kge_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly KGEs for ensembles -->

<!-- ```{r monthly kge ensemble, echo = F} -->

<!-- mo_kge <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>%  -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = KGE(grid_ensemble, obs), -->
<!--             statn_ensemble = KGE(statn_ensemble, obs), -->
<!--             all_ensemble = KGE(all_ensemble, obs), -->
<!--             grid_all = KGE(grid_all, obs), -->
<!--             statn_all = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- kable(mo_kge) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot <- mo_kge %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'KGE') %>% -->
<!--   drop_na() -->

<!-- kge_plot_mo <- ggplot(mo_pivot, aes(month, KGE, color = cal_run)) + geom_point(size = 3) + -->
<!--     scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   theme_bw() -->

<!-- ggplotly(kge_plot_mo) -->

<!-- ``` -->

<!-- ## RMSE -->

<!-- ### Ensemble RMSE for entire period -->

<!-- ```{r ensemble rmse, echo = F} -->

<!-- ensem_rmse_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = rmse(grid_ensemble, obs), -->
<!--             statn_ensemble = rmse(statn_ensemble, obs), -->
<!--             all_ensemble = rmse(all_ensemble, obs), -->
<!--             grid_all = rmse(grid_all, obs), -->
<!--             statn_all = rmse(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(ensem_rmse_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual RMSE for ensembles -->

<!-- ```{r annual rmse ensem, echo = F} -->

<!-- yr_rmse_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = rmse(grid_ensemble, obs), -->
<!--             statn_ensemble = rmse(statn_ensemble, obs), -->
<!--             all_ensemble = rmse(all_ensemble, obs), -->
<!--             grid_all = rmse(grid_all, obs), -->
<!--             statn_all = rmse(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- yr_rmse_table <- yr_rmse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'rmse') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'rmse') -->

<!-- yr_rmse_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_rmse_plot <- yr_rmse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'rmse') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual rmse vals for ensembles -->
<!-- plot_rmse_yr_ensem <- function(calyear) { -->
<!--   yr_rmse_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = rmse)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_rmse_plot$year) -->

<!-- lapply(name, plot_rmse_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly RMSEs for ensembles -->

<!-- ```{r month ensem rmse, echo = F} -->

<!-- mo_rmse <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>%  -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = rmse(grid_ensemble, obs), -->
<!--             statn_ensemble = rmse(statn_ensemble, obs), -->
<!--             all_ensemble = rmse(all_ensemble, obs), -->
<!--             grid_all = rmse(grid_all, obs), -->
<!--             statn_all = rmse(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(mo_rmse) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot2 <- mo_rmse %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'RMSE') %>% -->
<!--   drop_na() -->

<!-- rmse_plot_mo <- ggplot(mo_pivot2, aes(month, RMSE, color = cal_run)) + geom_point(size = 3) + -->
<!--     scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   theme_bw() -->

<!-- ggplotly(rmse_plot_mo) -->

<!-- ``` -->

<!-- ## NSE -->

<!-- ### Ensemble NSE for entire period -->

<!-- ```{r nse all, echo = F} -->

<!-- ensem_nse_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = NSE(grid_ensemble, obs), -->
<!--             statn_ensemble = NSE(statn_ensemble, obs), -->
<!--             all_ensemble = NSE(all_ensemble, obs), -->
<!--             grid_all = NSE(grid_all, obs), -->
<!--             statn_all = NSE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- kable(ensem_nse_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual NSE for ensembles -->

<!-- ```{r annual nse, echo = F} -->

<!-- yr_nse_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = NSE(grid_ensemble, obs), -->
<!--             statn_ensemble = NSE(statn_ensemble, obs), -->
<!--             all_ensemble = NSE(all_ensemble, obs), -->
<!--             grid_all = NSE(grid_all, obs), -->
<!--             statn_all = NSE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- yr_nse_table <- yr_nse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'nse') -->

<!-- yr_nse_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_nse_plot <- yr_nse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual nse vals for ensembles -->
<!-- plot_nse_yr_ensem <- function(calyear) { -->
<!--   yr_nse_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = nse)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_nse_plot$year) -->

<!-- lapply(name, plot_nse_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly NSE for ensembles -->

<!-- ```{r monthly rmse, echo =F} -->

<!-- mo_nse <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>%  -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = NSE(grid_ensemble, obs), -->
<!--             statn_ensemble = NSE(statn_ensemble, obs), -->
<!--             all_ensemble = NSE(all_ensemble, obs), -->
<!--             grid_all = NSE(grid_all, obs), -->
<!--             statn_all = NSE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(mo_nse) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot2 <- mo_nse %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'nse') %>% -->
<!--   drop_na() -->

<!-- nse_plot_mo <- ggplot(mo_pivot2, aes(month, nse, color = cal_run)) + geom_point(size = 3) + -->
<!--   theme_bw() + -->
<!--   scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange"))  -->

<!-- ggplotly(nse_plot_mo) -->

<!-- ``` -->

<!-- ## Percent bias -->

<!-- Positive values indicate model overestimation and negative values model underestimation. -->

<!-- ### Ensemble pbias for entire period -->

<!-- ```{r pbias all, echo = F} -->

<!-- ensem_pbias_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = pbias(grid_ensemble, obs), -->
<!--             statn_ensemble = pbias(statn_ensemble, obs), -->
<!--             all_ensemble = pbias(all_ensemble, obs), -->
<!--             grid_all = pbias(grid_all, obs), -->
<!--             statn_all = pbias(statn_all, obs)) -->

<!-- kable(ensem_pbias_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual pbias for ensembles -->

<!-- ```{r annualpbias, echo = F} -->

<!-- yr_pbias_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = pbias(grid_ensemble, obs), -->
<!--             statn_ensemble = pbias(statn_ensemble, obs), -->
<!--             all_ensemble = pbias(all_ensemble, obs), -->
<!--             grid_all = pbias(grid_all, obs), -->
<!--             statn_all = pbias(statn_all, obs)) -->

<!-- yr_pbias_table <- yr_pbias_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'pbias') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'pbias') -->

<!-- yr_pbias_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_pbias_plot <- yr_pbias_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'pbias') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual pbias vals for ensembles -->
<!-- plot_pbias_yr_ensem <- function(calyear) { -->
<!--   yr_pbias_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = pbias)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_pbias_plot$year) -->

<!-- lapply(name, plot_pbias_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly pbias for ensembles -->

<!-- ```{r monthly pbias, echo =F} -->

<!-- mo_pbias <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>%  -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = pbias(grid_ensemble, obs), -->
<!--             statn_ensemble = pbias(statn_ensemble, obs), -->
<!--             all_ensemble = pbias(all_ensemble, obs), -->
<!--             grid_all = pbias(grid_all, obs), -->
<!--             statn_all = pbias(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(mo_pbias) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot2 <- mo_pbias %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'pbias') %>% -->
<!--   drop_na() -->

<!-- pbias_plot_mo <- ggplot(mo_pivot2, aes(month, pbias, color = cal_run)) +  -->
<!--   geom_point(size = 3) + -->
<!--   scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   theme_bw() -->

<!-- ggplotly(pbias_plot_mo) -->

<!-- ``` -->

<!-- ## Cumulative Q -->

<!-- ```{r cumq, echo = F} -->

<!-- ensem_cumq <- ensem %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   filter(month %in% c(4:9)) %>% -->
<!--   dplyr::select(-month) %>%  -->
<!--   pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>% -->
<!--   group_by(cal_run, year(date)) %>% -->
<!--   mutate(cumsum = cumsum(replace_na(q_m3s, 0))) %>% -->
<!--   rename(year = 4) %>% -->
<!--   mutate(year = as.factor(year)) %>% -->
<!--   ungroup() -->

<!-- #plot annual cumulative Q vals for ensembles -->
<!-- plot_cumulq_yr_ensem <- function(calyear) { -->
<!--   {ensem_cumq %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = date, -->
<!--                y = cumsum, -->
<!--                color = cal_run)) + -->
<!--     geom_line() + -->
<!--     scale_color_manual(name="cal_run", values=c("obs"="black", -->
<!--                                              "grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     labs(y = 'cumulative Q (m3s)')} %>% -->
<!--     ggplotly() -->
<!-- } -->

<!-- name <- levels(ensem_cumq$year) -->

<!-- cumulative_plots <- lapply(name, plot_cumulq_yr_ensem) -->

<!-- htmltools::tagList(cumulative_plots) -->

<!-- cumulative_plots -->

<!-- ``` -->

<!-- ## Percentiles -->

<!-- ```{r percentiles, echo = F} -->

<!-- quantile_calc <- function(x, probs, date) { -->
<!--   tibble(x = quantile(x, probs), probs = probs, -->
<!--          date = date[x = quantile(x, probs)]) -->
<!-- } -->

<!-- percent <- ensem_cumq %>% -->
<!--   group_by(year, cal_run) %>% -->
<!--   summarize(quantile_calc(cumsum, c(0.25, 0.5, 0.75), date)) %>% -->
<!--   mutate(date = as.Date(date)) %>% -->
<!--   mutate(probs = as.factor(probs)) -->

<!-- ### -->
<!-- quantile_cal <- function(data, prob) { -->
<!--   data %>% -->
<!--     dplyr::group_by(cal_run, year) %>% -->
<!--     mutate(quantile = paste0(prob), -->
<!--            value = quantile(cumsum, prob)) %>% -->
<!--     group_by(cal_run, year, quantile) %>%  -->
<!--     filter(cumsum >= value) %>% -->
<!--     filter(date == min(date)) -->
<!--   } -->

<!-- quantiles <- c(0.25, 0.5, 0.75) -->

<!-- quants <- quantiles %>% -->
<!--   map(~ quantile_cal(data = ensem_cumq, prob = .)) %>% -->
<!--   bind_rows() -->
<!-- #### -->

<!-- #plot percentiles -->
<!-- plot_perc_ensem <- function(perc) { -->
<!--   {percent %>% -->
<!--     filter(probs == perc) %>% -->
<!--     ggplot(aes(x = year, -->
<!--                y = x, -->
<!--                color = cal_run)) + -->
<!--     geom_point(size=3) + -->
<!--     scale_color_manual(name="cal_run", values=c("obs"="black", -->
<!--                                              "grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--     ggtitle(perc) + -->
<!--     ylab('Streamflow (cms)') + -->
<!--     theme_bw()} %>% -->
<!--     ggplotly() -->
<!-- } -->

<!-- name <- levels(percent$probs) -->

<!-- perc_plots <- lapply(name, plot_perc_ensem) -->

<!-- htmltools::tagList(perc_plots) -->

<!-- perc_plots -->

<!-- ``` -->

<!-- ## Max Q prior to April 1 -->

<!-- ```{r early max q, echo = F} -->

<!-- ensem_bould <- left_join(ensem, bouldcr, by = 'date') %>% -->
<!--   pivot_longer(!c(date, year), names_to = 'source', values_to = 'q_m3s') %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   filter(month %in% 3) %>% -->
<!--   group_by(source, year) %>% -->
<!--   slice_max(q_m3s, n=1) -->

<!-- early_peakq <- ensem_bould %>% -->
<!--   mutate(doy = yday(date)) %>% -->
<!--   dplyr::select(c(doy, source, q_m3s, year)) %>% -->
<!--   drop_na() %>%  -->
<!--   pivot_wider(names_from = 'source', values_from = c('q_m3s', 'doy')) -->

<!-- marpeakq <- ggplot(early_peakq, aes(x = doy_bouldcr)) + -->
<!--   geom_point(aes(y = doy_all_ensemble, shape = "all_ensemble",  -->
<!--                  color = 'all_ensemble'), size=2.5) + -->
<!--   geom_point(aes(y = doy_grid_all, shape = 'grid_all',  -->
<!--                  color = 'grid_all'), size=2.5) + -->
<!--   geom_point(aes(y = doy_grid_ensemble, shape = 'grid_ensemble',  -->
<!--                  color = 'grid_ensemble'), size=2.5) + -->
<!--   geom_point(aes(y = doy_statn_all, shape = 'statn_all',  -->
<!--                  color = 'statn_all'), size=2.5) + -->
<!--   geom_point(aes(y = doy_statn_ensemble, shape = 'statn_ensemble',  -->
<!--                  color = 'statn_ensemble'), size=2.5) + -->
<!--   geom_abline(intercept = 0, slope = 1, size=1, color='black') + -->
<!--   labs(x = "Boulder Creek DOY for March Peak Q",  -->
<!--        y = 'DOY for March Peak Q') + -->
<!--   theme_bw() + -->
<!--   scale_shape_manual(name="source", values=c("all_ensemble"=15,  -->
<!--                                             "grid_all"=16,  -->
<!--                                             "grid_ensemble"=17, -->
<!--                                             "statn_all"=18,  -->
<!--                                             "statn_ensemble" = 8)) + -->
<!--     scale_color_manual(name="source", values=c("all_ensemble"='orange',  -->
<!--                                             "grid_all"='purple1',  -->
<!--                                             "grid_ensemble"='maroon', -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "statn_ensemble" = 'darkturquoise')) + -->
<!--   ggtitle('DOY of peak Q in March') -->

<!-- ggplotly(marpeakq) -->

<!-- # ggplot(early_peakq, aes(doy_bouldcr, doy_grid_all)) + -->
<!-- #   geom_point(size=2.5) + -->
<!-- #   geom_abline(intercept = 0, slope = 1, size=1, color='red') + -->
<!-- #   theme_bw() + -->
<!-- #   ggtitle('DOY of peak Q prior to April 1st - Grid All') -->
<!-- #  -->
<!-- #  -->
<!-- # ggplot(early_peakq, aes(doy_bouldcr, doy_grid_ensemble)) + -->
<!-- #   geom_point(size=2.5) + -->
<!-- #   geom_abline(intercept = 0, slope = 1, size=1, color='red') + -->
<!-- #   theme_bw() + -->
<!-- #   ggtitle('DOY of peak Q prior to April 1st - Grid Ensemble') -->
<!-- #  -->
<!-- #  -->
<!-- # ggplot(early_peakq, aes(doy_bouldcr, doy_statn_all)) + -->
<!-- #   geom_point(size=2.5) + -->
<!-- #   geom_abline(intercept = 0, slope = 1, size=1, color='red') + -->
<!-- #   theme_bw() + -->
<!-- #   ggtitle('DOY of peak Q prior to April 1st - Station All') -->
<!-- #  -->
<!-- #  -->
<!-- # ggplot(early_peakq, aes(doy_bouldcr, doy_statn_ensemble)) + -->
<!-- #   geom_point(size=2.5) + -->
<!-- #   geom_abline(intercept = 0, slope = 1, size=1, color='red') + -->
<!-- #   theme_bw() + -->
<!-- #   ggtitle('DOY of peak Q prior to April 1st - Station Ensemble') -->

<!-- early_peakq_table <- ensem_bould %>% -->
<!--   mutate(doy = yday(date)) %>% -->
<!--   dplyr::arrange(date) -->

<!-- kable(early_peakq_table) %>% -->
<!--   kable_styling(full_width = F) %>%  -->
<!--   scroll_box(width = "100%", height = "500px") -->

<!-- bouldcr3 <- bouldcr %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   filter(month == 3, year == 2011) -->

<!-- test1 <- ensem_pivot %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   filter(year == 2011, -->
<!--          month == 3) %>%  -->
<!--   ggplot(aes(date, q_m3s, color = cal_run)) + -->
<!--   geom_line() + -->
<!--   scale_color_manual(name="cal_run", values=c('obs'='black', -->
<!--                                             "grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   geom_line(data = bouldcr3, aes(date, bouldcr/10), color = 'darkgrey') + -->
<!--   scale_y_continuous(name = 'Q_m3s', -->
<!--                     sec.axis = sec_axis(~.*10, name='BoulderCr_cms')) + -->
<!--   theme_bw() + -->
<!--   ggtitle('March 2011') -->

<!-- ggplotly(test1) %>% -->
<!--   add_lines(data=bouldcr, x=~date, y=~bouldcr, colors=NULL, yaxis="y2",  -->
<!--               inherit=FALSE, showlegend = FALSE) %>% -->
<!--   layout(yaxis2 = list(overlaying = "y", side = "right", -->
<!--                        color = 'darkgray', -->
<!--                        title = "BoulderCr_cms"), -->
<!--          legend = list(x = 1.1, y = 0.95)) -->

<!-- ################################################ -->

<!-- bouldcr3_2 <- bouldcr %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   filter(month == 3, year == 2019) -->

<!-- test2 <- ensem_pivot %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   filter(year == 2019, -->
<!--          month == 3) %>%  -->
<!--   ggplot(aes(date, q_m3s, color = cal_run)) + -->
<!--   geom_line() + -->
<!--   scale_color_manual(name="cal_run", values=c('obs'='black', -->
<!--                                             "grid_ensemble"="maroon",  -->
<!--                                             "statn_ensemble"="darkturquoise",  -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4',  -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   geom_line(data = bouldcr3_2, aes(date, bouldcr/10), color = 'darkgrey') + -->
<!--   scale_y_continuous(name = 'Q_m3s', -->
<!--                     sec.axis = sec_axis(~.*10, name='BoulderCr_cms')) + -->
<!--   theme_bw() + -->
<!--   ggtitle('March 2019') -->

<!-- ggplotly(test2) %>% -->
<!--   add_lines(data=bouldcr, x=~date, y=~bouldcr, colors=NULL, yaxis="y2",  -->
<!--               inherit=FALSE, showlegend = FALSE) %>% -->
<!--   layout(yaxis2 = list(overlaying = "y", side = "right", -->
<!--                        color = 'darkgray', -->
<!--                        title = "BoulderCr_cms"), -->
<!--          legend = list(x = 1.1, y = 0.95)) -->

<!-- ``` -->

<!-- # Catchment results -->

<!-- ## Precip -->

<!-- ```{r catch precip, echo = F} -->
<!-- setwd("/Users/megansears/Documents/Repos/fourmile/catchment_output") -->

<!-- statn <- read_csv('statnall.csv') -->
<!-- grid <- read_csv('gridall.csv') -->

<!-- grid <- grid %>% mutate(site = 'grid') -->
<!-- statn <- statn %>% mutate(site = 'statn') -->

<!-- catchment <- bind_rows(grid, statn) %>% -->
<!--   mutate(year = as.factor(year(time))) %>% -->
<!--   group_by(year, site) %>% -->
<!--   mutate(cumulative_p = cumsum(precip)) -->

<!-- #plot cumulative precip by year -->
<!-- plot_cumprecip <- function(calyear) { -->
<!--   {catchment %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = time, -->
<!--                y = cumulative_p, color = site)) + -->
<!--     geom_line(size=1) + -->
<!--     ggtitle(calyear) + -->
<!--     ylab('Cumulative precip (mm)') + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1))} %>% -->
<!--     ggplotly() -->
<!-- } -->

<!-- name <- levels(catchment$year) -->

<!-- precip_plots <- lapply(name, plot_cumprecip) -->

<!-- htmltools::tagList(precip_plots) -->

<!-- precip_plots -->

<!-- ## total P -->
<!-- max_cumulativep <- catchment %>% -->
<!--   group_by(year, site) %>% -->
<!--   summarize(total_cumulativep = max(cumulative_p)) -->

<!-- ggplot(max_cumulativep, aes(x=year, y=total_cumulativep, color=site)) +  -->
<!--   geom_point(size=3) + -->
<!--   ggtitle('Total cumulative precip (mm) by year') -->

<!-- ``` -->

<!-- ## Tmean -->

<!-- ```{r tmean catch, echo = F} -->
<!-- ## tmean  -->
<!-- tmean <- catchment %>% -->
<!--   group_by(year, site) %>% -->
<!--   summarize(tmean = mean(tmean)) %>% -->
<!--   ggplot(aes(x = year, y=tmean, color = site)) + -->
<!--   geom_point(size=3) + -->
<!--   ggtitle('Average temp by year') -->

<!-- tmean -->

<!-- ``` -->

<!-- ## RD and RG -->

<!-- ```{r rd and rg, echo = F, fig.width = 14, fig.asp = 1, out.width = "100%"} -->

<!-- gw_grid <- read_csv(here('catchment_output/gridall_outlet.csv')) %>% -->
<!--   dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>% -->
<!--   mutate(source = 'grid') -->

<!-- gw_statn <- read_csv(here('catchment_output/statnall_outlet.csv')) %>% -->
<!--   dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>% -->
<!--   mutate(source = 'statn') -->

<!-- gw <- bind_rows(gw_grid, gw_statn) %>% -->
<!--   mutate(year = year(time)) %>% -->
<!--   rename(date = time) -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=catchmentRD1_w), size =1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol = 2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('RD1') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=catchmentRD2_w), size =1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol = 2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('RD2') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=catchmentRG1_w), size =1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol = 2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('RG1') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=catchmentRG2_w), size =1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol=2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('RG2') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- gw <- gw %>% -->
<!--   group_by(year, source) %>% -->
<!--   mutate(RD1cumsum = cumsum(catchmentRD1_w), -->
<!--          RD2cumsum = cumsum(catchmentRD2_w), -->
<!--          RG1cumsum = cumsum(catchmentRG1_w), -->
<!--          RG2cumsum = cumsum(catchmentRG2_w)) -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=RD1cumsum), size =1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol=2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('Cumulative RD1') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=RD2cumsum), size = 1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol=2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('Cumulative RD2') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=RG1cumsum), size  = 1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol = 2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('Cumulative RG1') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ggplot(gw, aes(x=date, color = source)) + -->
<!--   geom_line(aes(y=RG2cumsum), size =1) + -->
<!--   facet_wrap(~year, scales = 'free', ncol = 2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   ggtitle('Cumulative RG2') + -->
<!--   theme(text = element_text(size = 20))    -->

<!-- ``` -->

<!-- ```{r rd and rg sums, echo = F, fig.width = 22, fig.asp = 1.5, out.width = "100%"} -->

<!-- ################## -->
<!-- # all 4 with facets -->

<!-- ggplot(gw, aes(x=date)) + -->
<!--   geom_line(aes(y=RG2cumsum, color = 'cumulative RG2'), size =1) + -->
<!--   geom_line(aes(y=RG1cumsum, color = 'cumulative RG1'), size =1) + -->
<!--   geom_line(aes(y=RD1cumsum, color = 'cumulative RD1'), size =1) + -->
<!--   geom_line(aes(y=RD2cumsum, color = 'cumulative RD2'), size =1) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   facet_wrap(~year + source, scales = 'free_x', ncol=2) + -->
<!--   ggtitle('Cumulative RD and RG') + -->
<!--   theme(text = element_text(size = 24))    -->

<!-- ``` -->

<!-- ```{r gw boxplot, echo = F} -->

<!-- gw_maxcumsum <- gw %>% -->
<!--   group_by(source, year) %>% -->
<!--   summarize(yr_max_RD1 = max(RD1cumsum), -->
<!--             yr_max_RD2 = max(RD2cumsum), -->
<!--             yr_max_RG1 = max(RG1cumsum), -->
<!--             yr_max_RG2 = max(RG2cumsum)) %>% -->
<!--   mutate(year = as.factor(year)) %>% -->
<!--   pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms') -->

<!-- ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) + -->
<!--   geom_boxplot(position = position_dodge(1)) +  -->
<!--   geom_point(position=position_jitterdodge(),alpha=0.3) + -->
<!--   ggtitle('Max annual cumulative sums') -->

<!-- ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) + -->
<!--   geom_boxplot(position = position_dodge(1)) + -->
<!--   geom_point(position=position_jitterdodge(),alpha=0.3) + -->
<!--   #geom_point(aes(color = source)) + -->
<!--   facet_wrap(~flow_type, scales = 'free') + -->
<!--   ggtitle('Max annual cumulative sums') -->

<!-- ``` -->

<!-- ### RD and RG by season -->

<!-- ```{r flow season, echo = F, fig.width = 12, fig.asp = 0.75, out.width = "100%"} -->

<!-- #group by season and get max sums by season -->
<!-- flow_season <- gw %>% -->
<!--   dplyr::select(RD1cumsum, RD2cumsum, RG1cumsum, RG2cumsum, date, year, source) %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   mutate(season = case_when(month %in% c(3,4,5,6) ~ 'snowmelt', -->
<!--                             month %in% c(7,8,9,10) ~ 'monsoon', -->
<!--                             month %in% c(11,12,1,2) ~ 'baseflow')) %>% -->
<!--   group_by(source, year, season) %>% -->
<!--   summarize(yr_season_RD1 = max(RD1cumsum), -->
<!--             yr_season_RD2 = max(RD2cumsum), -->
<!--             yr_season_RG1 = max(RG1cumsum), -->
<!--             yr_season_RG2 = max(RG2cumsum)) %>% -->
<!--             pivot_longer(!c(year, source, season), names_to = 'flow_type', values_to = 'flow_cms') -->

<!-- flow_season %>% -->
<!--   ggplot(aes(x = flow_type, y = flow_cms, fill = source)) + -->
<!--   geom_boxplot(position = position_dodge(1)) + -->
<!--   geom_point(position=position_jitterdodge(),alpha=0.3) + -->
<!--   #geom_point(aes(color = source)) + -->
<!--   facet_wrap(~season, scales = 'free') + -->
<!--   ggtitle('Max annual season cumulative sums') + -->
<!--   theme_bw() + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) + -->
<!--   theme(text = element_text(size = 20))  -->

<!-- flow_season %>% -->
<!--   ggplot(aes(x = flow_type, y = flow_cms, fill = source)) + -->
<!--   geom_boxplot(position = position_dodge(1)) + -->
<!--   geom_point(position=position_jitterdodge(),alpha=0.3) + -->
<!--   #geom_point(aes(color = source)) + -->
<!--   facet_wrap(~season + flow_type, scales = 'free') + -->
<!--   ggtitle('Max annual season cumulative sums') + -->
<!--   theme_bw() + -->
<!--   theme(text = element_text(size = 20)) -->

<!-- ``` -->

<!-- # HRU results -->

<!-- ## HRU Map -->

<!-- ```{r map, echo = F} -->

<!-- locations <- read_csv(here('data/GIS', 'fourmile_sensors.csv')) -->

<!-- hrus <- st_read(here('data/GIS', 'hrus_final.shp')) -->

<!-- locations <- locations %>% -->
<!--   st_as_sf(., coords=c('y','x'), crs = 4326) -->

<!-- mapviewOptions(basemaps = c("Esri.WorldShadedRelief", "OpenStreetMap.DE"), -->
<!--                layers.control.pos = "topright") -->

<!-- mapview(locations, zcol = 'type') + mapview(hrus) -->

<!-- ``` -->

<!-- ## Niwot and HRU7 -->

<!-- ### SWE and precip -->

<!-- ```{r niwot, echo = F, fig.width = 14, fig.asp = 1, out.width = "100%"} -->
<!-- library(snotelr) -->

<!-- niwot_snotel <- snotel_download(site_id = 663, internal = T) -->

<!-- niwot_snotel <- niwot_snotel %>% -->
<!--   dplyr::select(date, snow_water_equivalent, precipitation) %>% -->
<!--   mutate(date = as.Date(date)) %>% -->
<!--   rename(swe_mm_snotel = snow_water_equivalent, -->
<!--          precip_mm_snotel = precipitation) %>% -->
<!--   #mutate(swe_mm_snotel = swe_mm_snotel*25.4, -->
<!--          #precip_mm_snotel = precip_mm_snotel*25.4) %>% -->
<!--   filter(date > '2010-12-31' & date < '2020-10-01') -->

<!-- hru7_grid <- read_csv(here('HRU_output/HRU_7_grid.csv')) %>% -->
<!--   dplyr::select(date, precip_mm, swe_mm) %>% -->
<!--   rename(precip_mm_hrugrid = precip_mm, -->
<!--          swe_mm_hrugrid = swe_mm) -->

<!-- hru7_statn <- read_csv(here('HRU_output/HRU_7_statn.csv')) %>% -->
<!--   dplyr::select(date, precip_mm, swe_mm) %>% -->
<!--   rename(precip_mm_hrustatn = precip_mm, -->
<!--          swe_mm_hrustatn = swe_mm) -->

<!-- niwot <- niwot_snotel %>% -->
<!--   left_join(hru7_grid, by = 'date') %>% -->
<!--   left_join(hru7_statn, by = 'date') -->

<!-- # look at swe -->
<!-- niwot_swe <- niwot %>% -->
<!--   dplyr::select(date, swe_mm_snotel, swe_mm_hrugrid, swe_mm_hrustatn) -->

<!-- kable(niwot_swe) %>% -->
<!--   kable_styling(full_width = F) %>%  -->
<!--   scroll_box(width = "100%", height = "500px") -->

<!-- # cumulative swe -->
<!-- niwot_swe2 <- niwot_swe %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   pivot_longer(!c(date,year), names_to = 'source', values_to = 'swe_mm') %>% -->
<!--   group_by(year, source) %>% -->
<!--   mutate(swe_cumsum_mm = cumsum(swe_mm)) -->

<!-- cum_swe_niwot <- ggplot(niwot_swe2, aes(date, swe_cumsum_mm, color = source)) + -->
<!--   geom_line(size =1) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   facet_wrap(~year, scales = 'free', ncol=2) + -->
<!--   theme(text = element_text(size = 20))   -->

<!-- cum_swe_niwot -->

<!-- #look at precip -->
<!-- niwot_p <- niwot %>% -->
<!--   dplyr::select(date, precip_mm_snotel, precip_mm_hrugrid, precip_mm_hrustatn) -->

<!-- kable(niwot_p) %>% -->
<!--   kable_styling(full_width = F) %>%  -->
<!--   scroll_box(width = "100%", height = "500px") -->

<!-- # cumulative precip -->
<!-- niwot_p2 <- niwot_p %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   pivot_longer(!c(date,year), names_to = 'source', values_to = 'precip_mm') %>% -->
<!--   group_by(year, source) %>% -->
<!--   mutate(p_cumsum_mm = cumsum(precip_mm)) -->

<!-- cum_precip_niwot <- ggplot(niwot_p2, aes(date, p_cumsum_mm, color = source)) + -->
<!--   geom_line(size=1) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   facet_wrap(~year, scales = 'free', ncol=2) + -->
<!--   theme(text = element_text(size = 20))   -->

<!-- cum_precip_niwot -->

<!-- ``` -->

<!-- ### ET -->

<!-- ```{r ET, echo = F} -->

<!-- niwot_et <- read_csv(here('HRU_output/Niwot_daily_et.csv')) %>% -->
<!--   filter(year > 2010) %>% -->
<!--   mutate(date = make_date(year = year, month = month, day = day)) %>% -->
<!--   dplyr::select(date, et) %>% -->
<!--   rename(et_niwot = et) -->

<!-- hru7_grid <- read_csv(here('HRU_output/HRU_7_grid.csv')) %>% -->
<!--   dplyr::select(date, AET_mm) %>% -->
<!--   rename(et_hru7grid = AET_mm) -->

<!-- hru7_statn <- read_csv(here('HRU_output/HRU_7_statn.csv')) %>% -->
<!--   dplyr::select(date, AET_mm) %>% -->
<!--   rename(et_hru7statn = AET_mm) -->

<!-- niwot_et_compare <- niwot_et %>% -->
<!--   left_join(hru7_grid, by = 'date') %>% -->
<!--   left_join(hru7_statn, by = 'date') %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   filter(year < 2021) %>%  -->
<!--   pivot_longer(!c(date, year), names_to= 'source', values_to = 'et_mm') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot ET by year -->
<!-- plot_ET <- function(calyear) { -->
<!--   {niwot_et_compare %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = date, -->
<!--                y = et_mm, color = source)) + -->
<!--     geom_line(size=1) + -->
<!--     ggtitle(calyear) + -->
<!--     ylab('ET (mm/day)') + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1))} %>% -->
<!--     ggplotly() -->
<!-- } -->

<!-- name <- levels(niwot_et_compare$year) -->

<!-- ET_plots <- lapply(name, plot_ET) -->

<!-- htmltools::tagList(ET_plots) -->

<!-- ET_plots -->

<!-- # 1:1 scatter plots for Niwot vs grid and Niwot vs station -->
<!-- niwot_et_scatter <- niwot_et_compare %>% -->
<!--   filter(et_mm > 0) -->

<!-- # Niwot vs grid -->
<!-- et_grid_obs <- niwot_et_scatter %>% -->
<!--   filter(source == c('et_niwot','et_hru7grid')) %>% -->
<!--   pivot_wider(names_from = 'source', values_from = 'et_mm') -->

<!-- #scatter of grid vs niwot -->
<!-- et_scatter_grid_obs <- ggplot(et_grid_obs, aes(x = et_niwot, y = et_hru7grid)) + -->
<!--   geom_point() + -->
<!--   geom_abline(intercept = 0, slope = 1, size=1, color='red') +  -->
<!--   ggtitle('Niwot ET vs. HRU7 grid ET') -->

<!-- et_scatter_grid_obs -->

<!-- # Niwot vs statn inputs -->
<!-- et_statn_obs <- niwot_et_scatter %>% -->
<!--   filter(source == c('et_niwot','et_hru7statn')) %>% -->
<!--   pivot_wider(names_from = 'source', values_from = 'et_mm') -->

<!-- #scatter of grid vs niwot -->
<!-- et_scatter_statn_obs <- ggplot(et_statn_obs, aes(x = et_niwot, y = et_hru7statn)) + -->
<!--   geom_point() + -->
<!--   geom_abline(intercept = 0, slope = 1, size=1, color='red') +  -->
<!--   ggtitle('Niwot ET vs. HRU7 station ET') -->

<!-- et_scatter_statn_obs -->

<!-- ``` -->

<!-- ## Tmean for HRUs 7, 9, 22, 36 -->

<!-- ```{r hru7 temp, echo = F, fig.width = 14, fig.asp = 1, out.width = "100%"} -->

<!-- tmean7_grid <- read_csv(here('HRU_output/HRU_7_grid.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t7_grid = tmean) -->

<!-- tmean7_statn <- read_csv(here('HRU_output/HRU_7_statn.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t7_statn = tmean) -->

<!-- tmean9_grid <- read_csv(here('HRU_output/HRU_9_grid.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t9_grid = tmean) -->

<!-- tmean9_statn <- read_csv(here('HRU_output/HRU_9_statn.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t9_statn = tmean) -->

<!-- tmean22_grid <- read_csv(here('HRU_output/HRU_22_grid.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t22_grid = tmean) -->

<!-- tmean22_statn <- read_csv(here('HRU_output/HRU_22_statn.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t22_statn = tmean) -->

<!-- tmean36_grid <- read_csv(here('HRU_output/HRU_36_grid.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t36_grid = tmean) -->

<!-- tmean36_statn <- read_csv(here('HRU_output/HRU_36_statn.csv')) %>% -->
<!--   dplyr::select(date, tmean) %>% -->
<!--   rename(t36_statn = tmean) -->

<!-- tmean_hrus <- list(tmean7_grid, tmean7_statn, tmean9_grid,  -->
<!--      tmean9_statn, tmean22_grid, tmean22_statn, -->
<!--      tmean36_grid, tmean36_statn) -->

<!-- t_hrus <- tmean_hrus %>% reduce(left_join, by='date') -->

<!-- t_hrus <- t_hrus %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   mutate(t7_diff = t7_statn - t7_grid, -->
<!--          t9_diff = t9_statn - t9_grid, -->
<!--          t22_diff = t22_statn - t22_grid, -->
<!--          t36_diff = t36_statn - t36_grid) %>% -->
<!--   group_by(year) %>% -->
<!--   mutate(t7_cumdiff = cumsum(t7_diff), -->
<!--          t9_cumdiff = cumsum(t9_diff), -->
<!--          t22_cumdiff = cumsum(t22_diff), -->
<!--          t36_cumdiff = cumsum(t36_diff)) -->

<!-- tmean_cumdiff_plot <- ggplot(t_hrus, aes(x = date)) + -->
<!--   geom_line(aes(y = t7_cumdiff, color = 'HRU7'), size =1) +  -->
<!--   geom_line(aes(y = t9_cumdiff, color = 'HRU9'), size =1) +  -->
<!--   geom_line(aes(y = t22_cumdiff, color = 'HRU22'), size =1) +  -->
<!--   geom_line(aes(y = t36_cumdiff, color = 'HRU36'), size =1) +  -->
<!--   facet_wrap(~year, scales = 'free',ncol=2) + -->
<!--   scale_x_date(labels = date_format("%b")) + -->
<!--   labs(y = 'Tmean cumulative difference (deg C)') + -->
<!--   ggtitle('Tmean station and grid cumulative difference (Station - Grid)') + -->
<!--   theme(text = element_text(size = 20))   -->

<!-- tmean_cumdiff_plot -->

<!-- ``` -->

<!-- ## Tmean difference -->

<!-- 2011-2020 average of daily station-grid Tmean  -->

<!-- ```{r cumulative tmean for HRUs, echo = F} -->
<!-- #grid first -->
<!-- setwd(here("HRU_output/grid")) -->

<!-- read_ages_grid <- function(dataframe){ -->
<!--   read_csv(dataframe, skip = 3) %>% -->
<!--   slice(3:n()) %>% -->
<!--   dplyr::select(-starts_with('@'), -->
<!--                 -1) %>% -->
<!--   mutate(time = ymd(time)) %>% -->
<!--   mutate_if(is.character, as.numeric) %>% -->
<!--   mutate(source = 'grid') -->

<!-- } -->

<!-- #pull in all filenames with HRU in the name -->
<!-- filenames <- list.files(".", pattern="HRU", full.names=F) -->

<!-- #read all the csvs using read_ages function -->
<!-- dataframes <- lapply(filenames, read_ages_grid)  -->

<!-- #name all the csvs based on first 6 characters -->
<!-- names(dataframes) <- substr(filenames, 1, 6) -->

<!-- #bind all and make 1 df -->
<!-- hru_grids <- dataframes %>% -->
<!--   bind_rows()  -->

<!-- #statn next -->
<!-- setwd(here("HRU_output/statn")) -->

<!-- read_ages_statn <- function(dataframe){ -->
<!--   read_csv(dataframe, skip = 3) %>% -->
<!--   slice(3:n()) %>% -->
<!--   dplyr::select(-starts_with('@'), -->
<!--                 -1) %>% -->
<!--   mutate(time = ymd(time)) %>% -->
<!--   mutate_if(is.character, as.numeric) %>% -->
<!--   mutate(source = 'statn') -->

<!-- } -->

<!-- #pull in all filenames with HRU in the name -->
<!-- filenames <- list.files(".", pattern="HRU", full.names=F) -->

<!-- #read all the csvs using read_ages function -->
<!-- dataframes <- lapply(filenames, read_ages_statn)  -->

<!-- #name all the csvs based on first 6 characters -->
<!-- names(dataframes) <- substr(filenames, 1, 6) -->

<!-- #bind all and make 1 df -->
<!-- hru_statn <- dataframes %>% -->
<!--   bind_rows()  -->

<!-- #get rid of a bunch of data -->
<!-- hru_grid_tmean <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, tmean)) %>% -->
<!--   rename(tmean_grid = tmean)  %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(tmean_all_grid = mean(tmean_grid)) -->

<!-- hru_statn_tmean <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, tmean)) %>% -->
<!--   rename(tmean_statn = tmean) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(tmean_all_statn = mean(tmean_statn)) -->

<!-- tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('ID')) %>% -->
<!--   mutate(diff = tmean_all_statn - tmean_all_grid) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_tmean <- left_join(hrus, tmean_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- locations_t <- locations %>%  -->
<!--   dplyr::filter(name %in% c('Gordon_South', 'Gordon_North', -->
<!--                         'Betasso', 'Niwot_Snotel', -->
<!--                         'University Camp SNTL')) -->

<!-- tm_shape(hrus_map_tmean) + -->
<!--   tm_polygons('diff', id = 'value', -->
<!--               breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ## Tmax difference -->

<!-- ```{r tmax, echo = F} -->

<!-- tmax_grid <- read_csv(here('HRU_output/reg_tmaxAll_prism.csv'))%>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmax_grid') %>% -->
<!--   mutate(time = dmy(time)) %>% -->
<!--   group_by(value) %>% -->
<!--   summarize(Tmax_mean_grid = mean(tmax_grid)) -->

<!-- tmax_statn <- read_csv(here('HRU_output/reg_tmaxAll.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmax_statn') %>% -->
<!--   mutate(time = dmy(time)) %>% -->
<!--   group_by(value) %>% -->
<!--   summarize(Tmax_mean_statn = mean(tmax_statn)) -->

<!-- tmax_hrus <- left_join(tmax_grid, tmax_statn, by = c('value')) %>% -->
<!--   mutate(diff = Tmax_mean_statn - Tmax_mean_grid, -->
<!--          value = as.numeric(value)) -->

<!-- hrus_map_tmax <- left_join(hrus, tmax_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- tm_shape(hrus_map_tmax) + -->
<!--   tm_polygons('diff', id = 'value', -->
<!--               breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ## Tmax monthly average -->

<!-- ```{r tmax month, echo = F} -->

<!-- tmax_grid_mo <- read_csv(here('HRU_output/reg_tmaxAll_prism.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmax_grid') %>% -->
<!--   mutate(time = dmy(time), -->
<!--          month = month(time)) %>% -->
<!--   filter(month %in% c(3:6)) %>% -->
<!--   group_by(value, month) %>% -->
<!--   summarize(avg_tmax_grid = mean(tmax_grid)) -->

<!-- tmax_statn_mo <- read_csv(here('HRU_output/reg_tmaxAll.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmax_statn') %>% -->
<!--   mutate(time = dmy(time), -->
<!--          month = month(time)) %>% -->
<!--   filter(month %in% c(3:6)) %>% -->
<!--   group_by(value, month) %>% -->
<!--   summarize(avg_tmax_statn = mean(tmax_statn)) -->

<!-- tmax_hrus_mo <- left_join(tmax_grid_mo, tmax_statn_mo, by = c('value', 'month')) %>% -->
<!--   mutate(mo_tmax_diff = avg_tmax_statn - avg_tmax_grid, -->
<!--          value = as.numeric(value)) -->

<!-- tmap_mode('plot') -->

<!-- #march -->
<!-- tmax_hrus_mar<- tmax_hrus_mo %>% -->
<!--   filter(month == 3) -->

<!-- hrus_map_tmax_mar <- left_join(hrus, tmax_hrus_mar, by = 'value') -->

<!-- mar_tmax <- tm_shape(hrus_map_tmax_mar) + -->
<!--   #tm_polygons('mo_tmax_diff', id = 'value') + -->
<!--   tm_fill('mo_tmax_diff', id = 'value', -->
<!--           breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!--   tm_borders() + -->
<!--   tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'March avg tmax', -->
<!--             legend.outside = T) -->

<!-- #april -->
<!-- tmax_hrus_apr<- tmax_hrus_mo %>% -->
<!--   filter(month == 4) -->

<!-- hrus_map_tmax_apr <- left_join(hrus, tmax_hrus_apr, by = 'value') -->

<!-- apr_tmax <- tm_shape(hrus_map_tmax_apr) + -->
<!--   tm_fill('mo_tmax_diff', id = 'value', -->
<!--           breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!--   tm_borders() + -->
<!--   tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'April avg tmax', -->
<!--             legend.outside = T) -->

<!-- #may -->
<!-- tmax_hrus_may<- tmax_hrus_mo %>% -->
<!--   filter(month == 5) -->

<!-- hrus_map_tmax_may <- left_join(hrus, tmax_hrus_may, by = 'value') -->

<!-- may_tmax <- tm_shape(hrus_map_tmax_may) + -->
<!--   tm_fill('mo_tmax_diff', id = 'value', -->
<!--          breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!--   tm_borders() + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'May avg tmax', -->
<!--             legend.outside = T) -->

<!-- #june -->
<!-- tmax_hrus_jun<- tmax_hrus_mo %>% -->
<!--   filter(month == 6) -->

<!-- hrus_map_tmax_jun <- left_join(hrus, tmax_hrus_jun, by = 'value') -->

<!-- jun_tmax <- tm_shape(hrus_map_tmax_jun) + -->
<!--   tm_fill('mo_tmax_diff', id = 'value', -->
<!--          breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!--   tm_borders() + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'June avg tmax', -->
<!--             legend.outside = T) -->

<!-- tmap_arrange(mar_tmax, apr_tmax, may_tmax, jun_tmax) -->

<!-- ``` -->

<!-- ## Tmin difference -->

<!-- ```{r tmin, echo = F} -->

<!-- tmin_grid <- read_csv(here('HRU_output/reg_tminAll_prism.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmin_grid') %>% -->
<!--   mutate(time = dmy(time)) %>% -->
<!--   group_by(value) %>% -->
<!--   summarize(Tmin_mean_grid = mean(tmin_grid)) -->

<!-- tmin_statn <- read_csv(here('HRU_output/reg_tminAll.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmin_statn') %>% -->
<!--   mutate(time = dmy(time)) %>% -->
<!--   group_by(value) %>% -->
<!--   summarize(Tmin_mean_statn = mean(tmin_statn)) -->

<!-- tmin_hrus <- left_join(tmin_grid, tmin_statn, by = c('value')) %>% -->
<!--   mutate(diff = Tmin_mean_statn - Tmin_mean_grid, -->
<!--          value = as.numeric(value)) -->

<!-- hrus_map_tmin <- left_join(hrus, tmin_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- tm_shape(hrus_map_tmin) + -->
<!--   tm_polygons('diff', id = 'value', -->
<!--               breaks = c(-6,-4,-2,0,2,4,6)) + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ## Tmin monthly average -->

<!-- ```{r tmin month, echo = F} -->
<!-- tmin_grid_mo <- read_csv(here('HRU_output/reg_tminAll_prism.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmin_grid') %>% -->
<!--   mutate(time = dmy(time), -->
<!--          month = month(time)) %>% -->
<!--   filter(month %in% c(3:6)) %>% -->
<!--   group_by(value, month) %>% -->
<!--   summarize(avg_tmin_grid = mean(tmin_grid)) -->

<!-- tmin_statn_mo <- read_csv(here('HRU_output/reg_tminAll.csv')) %>%  -->
<!--   pivot_longer(cols = `3`:`19`, -->
<!--                names_to = 'value',  -->
<!--                values_to = 'tmin_statn') %>% -->
<!--   mutate(time = dmy(time), -->
<!--          month = month(time)) %>% -->
<!--   filter(month %in% c(3:6)) %>% -->
<!--   group_by(value, month) %>% -->
<!--   summarize(avg_tmin_statn = mean(tmin_statn)) -->

<!-- tmin_hrus_mo <- left_join(tmin_grid_mo, tmin_statn_mo, by = c('value', 'month')) %>% -->
<!--   mutate(mo_tmin_diff = avg_tmin_statn - avg_tmin_grid, -->
<!--          value = as.numeric(value)) -->

<!-- tmap_mode('plot') -->

<!-- #march -->
<!-- tmin_hrus_mar<- tmin_hrus_mo %>% -->
<!--   filter(month == 3) -->

<!-- hrus_map_tmin_mar <- left_join(hrus, tmin_hrus_mar, by = 'value') -->

<!-- mar_tmin <- tm_shape(hrus_map_tmin_mar) + -->
<!--   #tm_polygons('mo_tmin_diff', id = 'value') + -->
<!--   tm_fill('mo_tmin_diff', id = 'value', -->
<!--           breaks = c(-5,-2.5,0,2.5,5)) + -->
<!--   tm_borders() + -->
<!--   tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'March avg tmin', -->
<!--             legend.outside = T) -->

<!-- #april -->
<!-- tmin_hrus_apr<- tmin_hrus_mo %>% -->
<!--   filter(month == 4) -->

<!-- hrus_map_tmin_apr <- left_join(hrus, tmin_hrus_apr, by = 'value') -->

<!-- apr_tmin <- tm_shape(hrus_map_tmin_apr) + -->
<!--   tm_fill('mo_tmin_diff', id = 'value', -->
<!--           breaks = c(-5,-2.5,0,2.5,5)) + -->
<!--   tm_borders() + -->
<!--   tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'April avg tmin', -->
<!--             legend.outside = T) -->

<!-- #may -->
<!-- tmin_hrus_may<- tmin_hrus_mo %>% -->
<!--   filter(month == 5) -->

<!-- hrus_map_tmin_may <- left_join(hrus, tmin_hrus_may, by = 'value') -->

<!-- may_tmin <- tm_shape(hrus_map_tmin_may) + -->
<!--   tm_fill('mo_tmin_diff', id = 'value', -->
<!--          breaks = c(-5,-2.5,0,2.5,5)) + -->
<!--   tm_borders() + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'May avg tmin', -->
<!--             legend.outside = T) -->

<!-- #june -->
<!-- tmin_hrus_jun<- tmin_hrus_mo %>% -->
<!--   filter(month == 6) -->

<!-- hrus_map_tmin_jun <- left_join(hrus, tmin_hrus_jun, by = 'value') -->

<!-- jun_tmin <- tm_shape(hrus_map_tmin_jun) + -->
<!--   tm_fill('mo_tmin_diff', id = 'value', -->
<!--           breaks = c(-5,-2.5,0,2.5,5)) + -->
<!--   tm_borders() + -->
<!-- tm_shape(locations_t) + -->
<!--   tm_symbols(col='blue', size = 0.25) + -->
<!--   tm_layout(title = 'June avg tmin', -->
<!--             legend.outside = T) -->

<!-- tmap_arrange(mar_tmin, apr_tmin, may_tmin, jun_tmin) -->

<!-- ``` -->

<!-- ## Cumulative Precip difference -->

<!-- 2011-2020 average of daily station-grid precip for DOY 365 (or 366) -->

<!-- ```{r cum precip, echo = F} -->
<!-- #get rid of a bunch of data -->
<!-- hru_grid_precip <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, precip)) %>% -->
<!--   rename(precip_grid = precip) -->

<!-- hru_statn_precip <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, precip)) %>% -->
<!--   rename(precip_statn = precip) -->

<!-- precip_hrus <- left_join(hru_grid_precip, hru_statn_precip, by = c('time', 'ID')) %>% -->
<!--   mutate(diff = precip_statn - precip_grid, -->
<!--          year = year(time)) %>% -->
<!--   group_by(year, ID) %>% -->
<!--   mutate(total_diff = cumsum(diff)) %>%  -->
<!--   filter(time == max(time)) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(precip_avg_cumdiff = mean(total_diff)) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_precip <- left_join(hrus, precip_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- locations_p <- locations %>% -->
<!--   filter(type == 'precip') -->

<!-- tm_shape(hrus_map_precip) + -->
<!--   tm_polygons('precip_avg_cumdiff', midpoint = NA, palette = 'YlGnBu', id = 'value') + -->
<!-- tm_shape(locations_p) + -->
<!--   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ```{r relative p, echo = F} -->

<!-- # reg_p <- read_csv(here('HRU_output/reg_precipAll.csv'), -->
<!-- #                       skip = 21) %>% -->
<!-- #   dplyr::select(-starts_with('@')) %>% -->
<!-- #   mutate(time = dmy(time)) %>% -->
<!-- #   summarize(across(`3`:`19`, ~ mean(.))) %>% -->
<!-- #   pivot_longer(cols = `3`:`19`, -->
<!-- #                names_to = 'value',  -->
<!-- #                values_to = 'precip_avg') %>% -->
<!-- #   mutate(value = as.numeric(value)) -->
<!-- #  -->
<!-- #  -->
<!-- # hrus_map_precip <- left_join(hrus_map_precip, reg_p, by = 'value')  -->
<!-- #  -->
<!-- # hrus_map_precip <- hrus_map_precip %>% -->
<!-- #   mutate(relative_avgcum_diff = precip_avg_cumdiff / precip_avg) -->
<!-- #  -->
<!-- # tm_shape(hrus_map_precip) + -->
<!-- #   tm_polygons('relative_avgcum_diff', midpoint = NA, palette = 'YlGnBu', id = 'value') + -->
<!-- # tm_shape(locations_p) + -->
<!-- #   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ## Solar radiation difference -->

<!-- 2011-2020 average of daily station-grid solar rad -->

<!-- ```{r cum sol rad, echo = F} -->

<!-- #get rid of a bunch of data -->
<!-- hru_grid_solrad <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, solRad)) %>% -->
<!--   rename(solrad_grid = solRad)  %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(solrad_all_grid = mean(solrad_grid)) -->

<!-- hru_statn_solrad <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, solRad)) %>% -->
<!--   rename(solrad_statn = solRad) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(solrad_all_statn = mean(solrad_statn)) -->

<!-- solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('ID')) %>% -->
<!--   mutate(diff = solrad_all_statn - solrad_all_grid) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_solrad <- left_join(hrus, solrad_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- locations_solrad <- locations %>%  -->
<!--   dplyr::filter(name %in% c('Gordon_South', -->
<!--                         'Betasso', 'Niwot_Snotel')) -->

<!-- tm_shape(hrus_map_solrad) + -->
<!--   tm_polygons('diff', midpoint = NA, palette = 'YlOrRd', id = 'value') + -->
<!-- tm_shape(locations_solrad) + -->
<!--   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ## RH difference -->

<!-- ```{r rh, echo = F} -->

<!-- #get rid of a bunch of data -->
<!-- hru_grid_rh <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, rhum)) %>% -->
<!--   rename(rh_grid = rhum)  %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(rh_all_grid = mean(rh_grid)) -->

<!-- hru_statn_rh <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, rhum)) %>% -->
<!--   rename(rh_statn = rhum) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(rh_all_statn = mean(rh_statn)) -->

<!-- rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('ID')) %>% -->
<!--   mutate(diff = rh_all_statn - rh_all_grid) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_rh <- left_join(hrus, rh_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- locations_rh <- locations %>%  -->
<!--   dplyr::filter(name %in% c('Gordon_South', 'Gordon_North', -->
<!--                         'Betasso', 'Niwot_Snotel')) -->

<!-- tm_shape(hrus_map_rh) + -->
<!--   tm_polygons('diff', midpoint = NA, id = 'value') + -->
<!-- tm_shape(locations_rh) + -->
<!--   tm_symbols(col='blue', size=4) -->

<!-- ``` -->

<!-- ## Average soil sat comparison -->

<!-- ```{r soilsat, echo = F} -->

<!-- #get rid of a bunch of data -->
<!-- hru_grid_soilsat <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, soilSat)) %>% -->
<!--   rename(soilsat_grid = soilSat)  %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(soilsat_all_grid = mean(soilsat_grid)) -->

<!-- hru_statn_soilsat <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, soilSat)) %>% -->
<!--   rename(soilsat_statn = soilSat) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(soilsat_all_statn = mean(soilsat_statn)) -->

<!-- soilsat_hrus <- left_join(hru_grid_soilsat, hru_statn_soilsat, by = c('ID')) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_soilsat <- left_join(hrus, soilsat_hrus, by = 'value') -->

<!-- tmap_mode('plot') -->

<!-- soilavg_grid <- tm_shape(hrus_map_soilsat) + -->
<!--   tm_polygons('soilsat_all_grid', midpoint = NA, id = 'value', -->
<!--               breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3)) + -->
<!--   tm_legend(legend.outside = T) + -->
<!--   tm_layout(main.title = 'Average soil sat') -->

<!-- soilavg_statn <- tm_shape(hrus_map_soilsat) + -->
<!--   tm_polygons('soilsat_all_statn', midpoint = NA, id = 'value', -->
<!--               breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3)) + -->
<!--   tm_legend(legend.outside = T) -->

<!-- tmap_arrange(soilavg_grid, soilavg_statn)  -->

<!-- ``` -->

<!-- ## Average max soil sat comparison -->

<!-- ```{r max soilsat, echo = F} -->

<!-- #get rid of a bunch of data -->
<!-- hru_grid_soilsat_max <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, soilSat)) %>% -->
<!--   rename(soilsat_grid = soilSat)  %>% -->
<!--   mutate(year = year(time)) %>%  -->
<!--   group_by(ID, year) %>% -->
<!--   summarize(soilsat_yrmax_grid = max(soilsat_grid)) %>% -->
<!--   summarize(avg_max_grid = mean(soilsat_yrmax_grid)) -->

<!-- hru_statn_soilsat_max <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, soilSat)) %>% -->
<!--   rename(soilsat_statn = soilSat) %>% -->
<!--   mutate(year = year(time)) %>%  -->
<!--   group_by(ID, year) %>% -->
<!--   summarize(soilsat_yrmax_statn = max(soilsat_statn)) %>% -->
<!--   summarize(avg_max_statn = mean(soilsat_yrmax_statn)) -->

<!-- soilsat_hrus_max <- left_join(hru_grid_soilsat_max, hru_statn_soilsat_max, by = c('ID')) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_soilsat_max <- left_join(hrus, soilsat_hrus_max, by = 'value') -->

<!-- tmap_mode('plot') -->

<!-- soilavg_grid_max <- tm_shape(hrus_map_soilsat_max) + -->
<!--   tm_polygons('avg_max_grid', midpoint = NA, id = 'value', -->
<!--               breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7)) + -->
<!--   tm_legend(legend.outside = T) + -->
<!--   tm_layout(main.title = 'Average max soil sat', -->
<!--             title.position = c('left', 'top')) -->

<!-- soilavg_statn_max <- tm_shape(hrus_map_soilsat_max) + -->
<!--   tm_polygons('avg_max_statn', midpoint = NA, id = 'value', -->
<!--               breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7)) + -->
<!--   tm_legend(legend.outside = T) -->

<!-- tmap_arrange(soilavg_grid_max, soilavg_statn_max)  -->

<!-- ``` -->

<!-- ## Average ET comparison -->

<!-- ```{r et, echo = F} -->
<!-- #get rid of a bunch of data -->
<!-- hru_grid_ET <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, actET)) %>% -->
<!--   rename(ET_grid = actET)  %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(ET_all_grid = mean(ET_grid)) -->

<!-- hru_statn_ET <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, actET)) %>% -->
<!--   rename(ET_statn = actET) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(ET_all_statn = mean(ET_statn)) -->

<!-- ET_hrus <- left_join(hru_grid_ET, hru_statn_ET, by = c('ID')) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_ET <- left_join(hrus, ET_hrus, by = 'value') -->

<!-- tmap_mode('plot') -->

<!-- ETavg_grid <- tm_shape(hrus_map_ET) + -->
<!--   tm_polygons('ET_all_grid', midpoint = NA, id = 'value', -->
<!--               breaks = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) + -->
<!--   tm_legend(legend.outside = T) + -->
<!--   tm_layout(main.title = 'Average ET') -->

<!-- ETavg_statn <- tm_shape(hrus_map_ET) + -->
<!--   tm_polygons('ET_all_statn', midpoint = NA, id = 'value', -->
<!--              breaks = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) + -->
<!--   tm_legend(legend.outside = T) -->

<!-- tmap_arrange(ETavg_grid, ETavg_statn) -->

<!-- ``` -->

<!-- ## Average ET difference -->

<!-- ```{r et diff, echo = F} -->

<!-- ET_hrus <- ET_hrus %>% -->
<!--   mutate(diff = ET_all_statn - ET_all_grid) -->

<!-- hrus_map_et_diff <- left_join(hrus, ET_hrus, by = 'value') -->

<!-- tmap_mode('plot') -->

<!-- tm_shape(hrus_map_et_diff) + -->
<!--   tm_polygons('diff', midpoint = NA, id = 'value')  + -->
<!--   tm_legend(legend.outside = T) + -->
<!--   tm_layout(main.title = 'Average ET difference') -->

<!-- ``` -->

<!-- ## Average LAI comparison -->

<!-- ```{r lai, echo = F} -->

<!-- #get rid of a bunch of data -->
<!-- hru_grid_LAI <- hru_grids %>% -->
<!--   dplyr::select(c(time, ID, LAI)) %>% -->
<!--   rename(LAI_grid = LAI)  %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(LAI_all_grid = mean(LAI_grid)) -->

<!-- hru_statn_LAI <- hru_statn %>% -->
<!--   dplyr::select(c(time, ID, LAI)) %>% -->
<!--   rename(LAI_statn = LAI) %>% -->
<!--   group_by(ID) %>% -->
<!--   summarize(LAI_all_statn = mean(LAI_statn)) -->

<!-- LAI_hrus <- left_join(hru_grid_LAI, hru_statn_LAI, by = c('ID')) %>% -->
<!--   rename(value = ID) -->

<!-- hrus_map_LAI <- left_join(hrus, LAI_hrus, by = 'value') -->

<!-- tmap_mode('view') -->

<!-- LAIavg_grid <- tm_shape(hrus_map_LAI) + -->
<!--   tm_polygons('LAI_all_grid', midpoint = NA, id = 'value', -->
<!--               breaks = c(1,2,3,4,5)) + -->
<!--   tm_legend(legend.outside = T) + -->
<!--   tm_layout(main.title = 'Average LAI') -->

<!-- LAIavg_statn <- tm_shape(hrus_map_LAI) + -->
<!--   tm_polygons('LAI_all_statn', midpoint = NA, id = 'value', -->
<!--               breaks = c(1,2,3,4,5)) + -->
<!--   tm_legend(legend.outside = T) -->

<!-- tmap_arrange(LAIavg_grid, LAIavg_statn) -->

<!-- ``` -->

<!-- ## LAI over time  -->

<!-- ```{r lai over time, echo = F} -->

<!-- # plot time series of LAI over for HRU 1 and HRU 12 -->
<!-- hru_grid_select <- hru_grids %>% -->
<!--   filter(ID == 1 | ID == 12) %>% -->
<!--   dplyr::select(time, ID, LAI_grid = LAI) -->

<!-- hru_statn_select <- hru_statn %>% -->
<!--   filter(ID == 1 | ID == 12) %>% -->
<!--   dplyr::select(time, ID, LAI_statn = LAI) -->

<!-- hru_select <- left_join(hru_grid_select, hru_statn_select, by = c('time', 'ID')) %>% -->
<!--   rename(value = ID) -->

<!-- # plot time series of LAI over for HRU 1 -->
<!-- hru_select %>% -->
<!--   mutate(year = year(time)) %>%  -->
<!--   filter(value == 1) %>%  -->
<!-- ggplot(aes(x = time)) + -->
<!--   geom_line(aes(y = LAI_grid, color = "HRU1_burn_grid"), size = 1) + -->
<!--   geom_line(aes(y = LAI_statn, color = 'HRU1_burn_statn'), size = 1) + -->
<!--   ggtitle('LAI over time for HRU 1 (burned)') + -->
<!--   scale_color_manual(name = 'HRU' , values=c("HRU1_burn_grid"='orange',  -->
<!--                                             "HRU1_burn_statn"='springgreen4')) + -->
<!--   theme_bw() -->

<!-- # plot time series of LAI over for HRU 12 -->
<!-- {hru_select %>% -->
<!--   mutate(year = year(time)) %>%  -->
<!--   filter(value == 12) %>%  -->
<!-- ggplot(aes(x = time)) + -->
<!--   geom_line(aes(y = LAI_grid, color = "HRU12_unburn_grid"), size = 1) + -->
<!--   geom_line(aes(y = LAI_statn, color = 'HRU12_unburn_statn'), size = 1) + -->
<!--   ggtitle('LAI over time for HRU 12 (unburned)') + -->
<!--   scale_color_manual(name = 'HRU' , values=c("HRU12_unburn_grid"='orange',  -->
<!--                                             "HRU12_unburn_statn"='springgreen4')) + -->
<!--   theme_bw()} %>% -->
<!--   ggplotly() -->

<!-- ``` -->

<!-- ## HRU 29 LAI and biomass -->

<!-- ```{r hru29 lai and biomass} -->

<!-- # plot time series of LAI 29 -->
<!-- hru_grid29 <- hru_grids %>% -->
<!--   filter(ID == 29) %>% -->
<!--   dplyr::select(time, ID, LAI_grid = LAI) -->

<!-- hru_statn29 <- hru_statn %>% -->
<!--   filter(ID == 29) %>% -->
<!--   dplyr::select(time, ID, LAI_statn = LAI) -->

<!-- hru_29 <- left_join(hru_grid29, hru_statn29, by = c('time', 'ID'))  -->

<!-- # plot time series of LAI over for HRU 29 -->
<!-- {hru_29 %>% -->
<!--   ggplot(aes(x = time)) + -->
<!--   geom_line(aes(y = LAI_grid, color = "HRU29_grid"), size = 1) + -->
<!--   geom_line(aes(y = LAI_statn, color = 'HRU29_statn'), size = 1) + -->
<!--   ggtitle('LAI - HRU 29') + -->
<!--   scale_color_manual(name = 'HRU' , values=c("HRU29_grid"='orange',  -->
<!--                                             "HRU29_statn"='springgreen4')) + -->
<!--   theme_bw()} %>% -->
<!--   ggplotly() -->

<!-- hru29grid_bio <- read_csv('/Users/megansears/Documents/Repos/fourmile/HRU_output/HRU_29_grid.csv') %>% -->
<!--   mutate(time = mdy(time)) %>% -->
<!--   dplyr::select(time, BioAct_grid = BioAct_kgha) -->

<!-- hru29statn_bio <- read_csv('/Users/megansears/Documents/Repos/fourmile/HRU_output/HRU_29_statn.csv') %>% -->
<!--   mutate(time = mdy(time)) %>% -->
<!--   dplyr::select(time, BioAct_statn= BioAct_kgha) -->

<!-- hru_29_bio <- left_join(hru29grid_bio, hru29statn_bio, by = c('time'))  -->

<!-- #plot time series of biomass over time for HRU 29 -->
<!-- {hru_29_bio %>% -->
<!--   ggplot(aes(x = time)) + -->
<!--   geom_line(aes(y = BioAct_grid, color = "HRU29_grid"), size = 1) + -->
<!--   geom_line(aes(y = BioAct_statn, color = 'HRU29_statn'), size = 1) + -->
<!--   ggtitle('Biomass') + -->
<!--   scale_color_manual(name = 'HRU' , values=c("HRU29_grid"='orange',  -->
<!--                                             "HRU29_statn"='springgreen4')) + -->
<!--   theme_bw()} %>% -->
<!--   ggplotly() -->

<!-- ``` -->

<!-- # Mauscript figs -->

<!-- ## Figure 1 -->

<!-- ```{r fig1} -->

<!-- library(usmap) -->

<!-- dem <- raster('/Users/megansears/Documents/Repos/fourmile/data/GIS/cutDEM.tif') -->

<!-- hru_bnds <- st_read('/Users/megansears/Documents/Repos/fourmile/data/GIS/hrus_final.shp') -->

<!-- tmap_mode('plot') -->

<!-- tm_shape(dem) + -->
<!--   tm_raster(palette = terrain.colors(5)) + -->
<!--   tm_shape(hru_bnds) + -->
<!--   tm_polygons(alpha =0, col = 'black') + -->
<!--   tm_compass() + -->
<!--   tm_scale_bar() + -->
<!--   tm_layout(legend.outside = T, -->
<!--             main.title = 'A') -->

<!-- dem2 <- as.data.frame(dem, xy = TRUE) -->

<!-- # ggplot() + -->
<!-- #  geom_polygon(data = hru_bnds, aes(x = x, y = y, fill = value)) + -->
<!-- #   geom_raster(data = dem2, aes(x = x, y = y, -->
<!-- #                                       fill = cutDEM)) + -->
<!-- #   theme_bw() + -->
<!-- #   coord_quickmap() -->

<!-- ################################################### -->

<!-- catchment_mo_mean <- catchment %>% -->
<!--   mutate(month = month(time)) %>% -->
<!--   group_by(site, month) %>% -->
<!--   summarise(across(precip:solRad, ~ mean(.x, na.rm = TRUE))) %>% -->
<!--   dplyr::select(site, month, -->
<!--                 precip_mean = precip, tmean_mean = tmean,  -->
<!--                 solrad_mean = solRad, rh_mean = rhum) -->

<!-- catchment_mo_sd <- catchment %>% -->
<!--   mutate(month = month(time)) %>% -->
<!--   group_by(site, month) %>% -->
<!--   summarise(across(precip:solRad, ~ sd(.x, na.rm = TRUE))) %>% -->
<!--   dplyr::select(site, month, -->
<!--                 precip_sd = precip, tmean_sd = tmean,  -->
<!--                 solrad_sd = solRad, rh_sd = rhum) -->

<!-- catchment_mo <- left_join(catchment_mo_mean, catchment_mo_sd,  -->
<!--                           by = c('month', 'site')) -->

<!-- catchment_mo2 <- catchment_mo %>% -->
<!--   mutate(month_name = month.abb[month]) %>% -->
<!--   mutate(month_name = factor(month_name,  -->
<!--                              levels = c('Jan', 'Feb', 'Mar', -->
<!--                                         'Apr', 'May', 'Jun', -->
<!--                                         'Jul', 'Aug', 'Sep', -->
<!--                                         'Oct', 'Nov', 'Dec'))) -->

<!-- p_cols <- ggplot(catchment_mo2, aes(x = month_name, y = precip_mean, -->
<!--                           color = site, fill = site)) + -->
<!--   geom_col(position='dodge') + -->
<!--   geom_errorbar(aes(ymin = 0, ymax = precip_mean + precip_sd),  -->
<!--                 position = position_dodge(width = 0.6), width = 0.2) + -->
<!--   theme_bw(base_size = 14) + -->
<!--   labs(y = 'Precipitation (mm)', -->
<!--        x = element_blank())  -->

<!-- ggsave("P.png", plot = p_cols, width = 8, height = 6, dpi = 300) -->

<!-- t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean,  -->
<!--                                       color = site, group = site)) + -->
<!--   geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) + -->
<!--   geom_line() + -->
<!--   geom_point() + -->
<!--   theme_bw(base_size=14) + -->
<!--   labs(y = expression("Mean temperature " ( degree~C)), -->
<!--        x = element_blank()) -->

<!-- t_moplot -->

<!-- sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean,  -->
<!--                                       color = site, group = site)) + -->
<!--   geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) + -->
<!--   geom_line() + -->
<!--   geom_point() + -->
<!--   theme_bw(base_size = 14) + -->
<!--   labs(y = expression("Solar radiation (MJ/m"^"2" * ")"), -->
<!--        x = element_blank())  -->

<!-- sol_moplot -->

<!-- rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean,  -->
<!--                                       color = site, group = site)) + -->
<!--   geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) + -->
<!--   geom_line() + -->
<!--   geom_point() + -->
<!--   theme_bw(base_size = 14) + -->
<!--   labs(y = 'RH (%)', x= element_blank())  -->
<!-- rh_moplot -->

<!-- quad <- ggarrange(p_cols,  -->
<!--           t_moplot, -->
<!--           sol_moplot, -->
<!--           rh_moplot, -->
<!--           common.legend = T, -->
<!--           legend = 'bottom', -->
<!--              nrow = 2, -->
<!--           ncol= 2) -->

<!-- ggsave("quad.png", plot = quad, width = 8, height = 6, dpi = 300) -->

<!-- ``` -->