### 2023-05-03 38HRU_model/fourmile_model_outputs.Rmd
added in burn params

---
title: "Fourmile Calibrations and Output"
author: "Megan Sears"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F, 
  warning  = F)

library(tidyverse)
library(plotly)
library(lubridate)
library(knitr)
library(kableExtra)
library(dataRetrieval)
library(hydroGOF)
library(sp)
library(raster)
library(sf)
library(mapview)
library(here)
library(scales)
library(dataRetrieval)
library(tmap)
library(ggpubr)
library(gridExtra)
library(stringr)

```

```{r load in files, include = F}

setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_model')

kge <- read_csv('calruns_KGE.csv')

```

# KGE for calibrations

```{r look at kge, echo = F}

kge_plot <- ggplot(kge, aes(year_missing, KGE, shape = input_data, color = input_data)) + geom_point(size = 3) +
  theme_bw()

ggplotly(kge_plot)

```

# Parameters

## Scatter plots

```{r read in params, echo = FALSE, message = FALSE, warning = FALSE, include = F}
setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_model/cals")

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 10)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

```

```{r params, echo = FALSE, message = FALSE, warning = FALSE}

# function to read files, remove top rows, paste site name
read_cals <- function(df, name){
    df <- df %>%
      slice(-c(1:9, 146:277)) %>%
      separate(Parameter, c("Parameter", "value"), sep = ",") %>%
      drop_na() %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)

}

grid_2011 <- read_cals(grid_2011., name = 'grid_2011')
grid_2012 <- read_cals(grid_2012., name = 'grid_2012')
grid_2013 <- read_cals(grid_2013., name = 'grid_2013')
grid_2014 <- read_cals(grid_2014., name = 'grid_2014')
grid_2015 <- read_cals(grid_2015., name = 'grid_2015')
grid_2016 <- read_cals(grid_2016., name = 'grid_2016')
grid_2017 <- read_cals(grid_2017., name = 'grid_2017')
grid_2018 <- read_cals(grid_2018., name = 'grid_2018')
grid_2019 <- read_cals(grid_2019., name = 'grid_2019')
grid_2020 <- read_cals(grid_2020., name = 'grid_2020')
grid_all <- read_cals(grid_all.c, name = 'grid_all')

grid <- bind_rows(grid_2011, grid_2012, grid_2013, grid_2014, grid_2015,
                  grid_2016, grid_2017, grid_2018, grid_2019, grid_2020,
                  grid_all) %>%
  mutate(input_data = 'grid') %>%
  mutate(year_missing_almost = gsub('grid_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost)) %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

statn_2011 <- read_cals(statn_2011, name = 'statn_2011')
statn_2012 <- read_cals(statn_2012, name = 'statn_2012')
statn_2013 <- read_cals(statn_2013, name = 'statn_2013')
statn_2014 <- read_cals(statn_2014, name = 'statn_2014')
statn_2015 <- read_cals(statn_2015, name = 'statn_2015')
statn_2016 <- read_cals(statn_2016, name = 'statn_2016')
statn_2017 <- read_cals(statn_2017, name = 'statn_2017')
statn_2018 <- read_cals(statn_2018, name = 'statn_2018')
statn_2019 <- read_cals(statn_2019, name = 'statn_2019')
statn_2020 <- read_cals(statn_2020, name = 'statn_2020')
statn_all <- read_cals(statn_all., name = 'statn_all')

statn <- bind_rows(statn_2011, statn_2012, statn_2013, statn_2014, statn_2015,
                  statn_2016, statn_2017, statn_2018, statn_2019, statn_2020,
                  statn_all) %>%
  mutate(input_data = 'statn') %>%
  mutate(year_missing_almost = gsub('statn_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost)) %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

params <- bind_rows(statn, grid)

params <- params %>%
  filter(!param %in% c('angstrom_a', 'angstrom_b', 
                      'Beta_min', 
                      'Beta_rsd', "Beta_trans", 'clat', 
                      'delayNitrification',
                      'denitrificationRateCoefficient', 
                      'denitrificationSoilSaturationThreshold',
                      'deposition_factor', 'drrad', 
                      'drspac', 'initLPS', 'initRG1',
                      'initRG2', 'Ksink', 'LExCoef', 
                      'longTZ', 'N_concRG1', 'N_concRG2', 'N_delay_RG1',
                      'N_delay_RG2', 
                      'nitrificationSoilTemperatureThreshold', 
                      'nitrificationSurfaceTemperatureThreshold',
                      'opti', 'piadin', 'rootfactor', 'sceno', 
                      'soilImpGT80', 'soilImpLT80',
                      'soilLinRed', 'soilMaxPerc', 'temp_lag')) %>%
    mutate(value = as.numeric(value),
         param = as.factor(param))

#function to plot all params
plot_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = year_missing,
               y = value,
               shape = input_data, 
               color = input_data)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plot_params)

#write.csv(params, 'params.csv')

```

## Density plots

```{r density plots, echo = F, message = F}

mean_params <- params %>%
  group_by(param, input_data) %>%
  summarize(mean = mean(value))

plotdens_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = value,
               fill = input_data, 
               color = input_data)) + 
    geom_density(alpha=0.5) +
    geom_vline(data=filter(mean_params, param == param_name), 
               aes(xintercept=mean, color=input_data),
             linetype="dashed", linewidth=0.75) +
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plotdens_params)

```

## Burn parameters

```{echo = FALSE, message = FALSE, warning = FALSE, include = F}

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=TRUE)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 10)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

# read in all the grid burn params
read_cals_burn <- function(df, name){
    df <- df %>%
      slice(-c(1:274)) %>%
      separate_wider_delim(Parameter, 
                       delim = ',',
                       names = c('soilID',
                                 'luca_a_rain',
                                 'luca_a_snow',
                                 'luca_soilMaxInfSummer',
                                 'luca_soilMaxDPS',
                                 'luca_lagSurfaceRunoff',
                                 'luca_soilDistMPSLPS')) %>%
      slice(-1) %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)
}

grid_2011_burn <- read_cals_burn(grid_2011., name = 'grid_2011')
grid_2012_burn <- read_cals_burn(grid_2012., name = 'grid_2012')
grid_2013_burn <- read_cals_burn(grid_2013., name = 'grid_2013')
grid_2014_burn <- read_cals_burn(grid_2014., name = 'grid_2014')
grid_2015_burn <- read_cals_burn(grid_2015., name = 'grid_2015')
grid_2016_burn <- read_cals_burn(grid_2016., name = 'grid_2016')
grid_2017_burn <- read_cals_burn(grid_2017., name = 'grid_2017')
grid_2018_burn <- read_cals_burn(grid_2018., name = 'grid_2018')
grid_2019_burn <- read_cals_burn(grid_2019., name = 'grid_2019')
grid_2020_burn <- read_cals_burn(grid_2020., name = 'grid_2020')
grid_all_burn <- read_cals_burn(grid_all.c, name = 'grid_all')

grid_burn <- bind_rows(grid_2011_burn, 
                  grid_2012_burn, 
                  grid_2013_burn, 
                  grid_2014_burn, 
                  grid_2015_burn,
                  grid_2016_burn, 
                  grid_2017_burn, 
                  grid_2018_burn, 
                  grid_2019_burn, 
                  grid_2020_burn,
                  grid_all_burn) %>%
  mutate(input_data = 'grid') %>%
  mutate(year_missing_almost = gsub('grid_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost))

# now do statn burn params
statn_2011_burn <- read_cals_burn(statn_2011, name = 'statn_2011')
statn_2012_burn <- read_cals_burn(statn_2012, name = 'statn_2012')
statn_2013_burn <- read_cals_burn(statn_2013, name = 'statn_2013')
statn_2014_burn <- read_cals_burn(statn_2014, name = 'statn_2014')
statn_2015_burn <- read_cals_burn(statn_2015, name = 'statn_2015')
statn_2016_burn <- read_cals_burn(statn_2016, name = 'statn_2016')
statn_2017_burn <- read_cals_burn(statn_2017, name = 'statn_2017')
statn_2018_burn <- read_cals_burn(statn_2018, name = 'statn_2018')
statn_2019_burn <- read_cals_burn(statn_2019, name = 'statn_2019')
statn_2020_burn <- read_cals_burn(statn_2020, name = 'statn_2020')
statn_all_burn <- read_cals_burn(statn_all., name = 'statn_all')

statn_burn <- bind_rows(statn_2011_burn, 
                  statn_2012_burn, 
                  statn_2013_burn, 
                  statn_2014_burn, 
                  statn_2015_burn,
                  statn_2016_burn, 
                  statn_2017_burn, 
                  statn_2018_burn, 
                  statn_2019_burn, 
                  statn_2020_burn,
                  statn_all_burn) %>%
  mutate(input_data = 'statn') %>%
  mutate(year_missing_almost = gsub('statn_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost))

burn_params <- bind_rows(grid_burn, statn_burn) %>%
  ungroup() %>%
  mutate(across(c(luca_a_rain:luca_soilDistMPSLPS), as.numeric)) %>%
  group_by(cal) %>%
  mutate(across(c(luca_a_rain:luca_soilDistMPSLPS), mean)) %>%
  filter(soilID == 22) %>%
  dplyr::select(-soilID) %>% 
  pivot_longer(!c(cal, input_data, year_missing), 
               names_to = 'params',
               values_to = 'value') %>% 
  mutate(params = str_remove(params, 'luca_')) %>%
  mutate(params = as.factor(params),
         burn = 'yes')

```

```{r}

plot_burnparams <- function(param_name) {
  burn_params %>%
    filter(params == param_name) %>%
    ggplot(aes(y = value,
               x = year_missing,
               color = input_data)) +
    geom_point(size = 3) +
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(burn_params$params)

lapply(name, plot_burnparams)

```

## Combined parameters

The two burned HRU parameters were averaged together.

```{r}
params <- params %>%
  mutate(burn = 'no')

params_combo <- bind_rows(params,
                          burn_params %>%
                            rename(param = params))

#function to plot all params
plot_params <- function(param_name) {
  params_combo %>%
    filter(param == param_name) %>%
    ggplot(aes(x = year_missing,
               y = value,
               shape = burn, 
               color = input_data)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params_combo$param)

lapply(name, plot_params)

```

# Outlet results

## Hydrograph

```{r hydrographs, echo = F}
flow <- read_csv('discharge_all.csv') %>%
  replace(.==-9999, NA)

flow_pivot <- flow %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date)))

q <- ggplot(flow_pivot, aes(date, q_m3s, color = cal_run)) +
  geom_line() +
  scale_color_manual(name="cal_run", values=c('obs' = 'black',
                                            "grid_2011"="seagreen",
                                            "grid_2012"="seagreen2",
                                            "grid_2013"="steelblue",
                                            "grid_2014"='dodgerblue',
                                            "grid_2015" = "cyan3",
                                            'grid_2016' = 'blue',
                                            'grid_2017' = 'blue4',
                                            'grid_2018' = 'cadetblue',
                                            'grid_2019' = 'lightgreen',
                                            'grid_2020' = 'olivedrab',
                                            'grid_all' = 'steelblue3',
                                            "statn_2011"="maroon",
                                            "statn_2012"="deeppink4",
                                            "statn_2013"="deeppink",
                                            "statn_2014"='purple',
                                            "statn_2015" = "mediumorchid1",
                                            'statn_2016' = 'indianred',
                                            'statn_2017' = 'sienna',
                                            'statn_2018' = 'salmon2',
                                            'statn_2019' = 'mediumpurple1',
                                            'statn_2020' = 'violetred',
                                            'statn_all' = 'red'))

ggplotly(q)

```

## KGE

### Overall KGE

```{r kge overall, echo = F}

flow_kge_all <- flow %>%
  summarize(grid2011 = KGE(grid_2011, obs),
            grid2012 = KGE(grid_2012, obs),
            grid2013 = KGE(grid_2013, obs),
            grid2014 = KGE(grid_2014, obs),
            grid2015 = KGE(grid_2015, obs),
            grid2016 = KGE(grid_2016, obs),
            grid2017 = KGE(grid_2017, obs),
            grid2018 = KGE(grid_2018, obs),
            grid2019 = KGE(grid_2019, obs),
            grid2020 = KGE(grid_2020, obs),
            gridall = KGE(grid_all, obs),
            statn2011 = KGE(statn_2011, obs),
            statn2012 = KGE(statn_2012, obs),
            statn2013 = KGE(statn_2013, obs),
            statn2014 = KGE(statn_2014, obs),
            statn2015 = KGE(statn_2015, obs),
            statn2016 = KGE(statn_2016, obs),
            statn2017 = KGE(statn_2017, obs),
            statn2018 = KGE(statn_2018, obs),
            statn2019 = KGE(statn_2019, obs),
            statn2020 = KGE(statn_2020, obs),
            statnall = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_kge_all <- flow_kge_all %>%
  pivot_longer(everything()) %>%
  arrange(value)

kable(flow_kge_all) %>%
  kable_styling(full_width = F)

```

### KGE by year

```{r kge by year, echo = F}

flow_year <- flow %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(grid2011 = KGE(grid_2011, obs),
            grid2012 = KGE(grid_2012, obs),
            grid2013 = KGE(grid_2013, obs),
            grid2014 = KGE(grid_2014, obs),
            grid2015 = KGE(grid_2015, obs),
            grid2016 = KGE(grid_2016, obs),
            grid2017 = KGE(grid_2017, obs),
            grid2018 = KGE(grid_2018, obs),
            grid2019 = KGE(grid_2019, obs),
            grid2020 = KGE(grid_2020, obs),
            gridall = KGE(grid_all, obs),
            statn2011 = KGE(statn_2011, obs),
            statn2012 = KGE(statn_2012, obs),
            statn2013 = KGE(statn_2013, obs),
            statn2014 = KGE(statn_2014, obs),
            statn2015 = KGE(statn_2015, obs),
            statn2016 = KGE(statn_2016, obs),
            statn2017 = KGE(statn_2017, obs),
            statn2018 = KGE(statn_2018, obs),
            statn2019 = KGE(statn_2019, obs),
            statn2020 = KGE(statn_2020, obs),
            statnall = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_kge <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>%
  pivot_wider(names_from = 'year', values_from = 'kge')

flow_kge %>%
  kable(.) %>%
  kable_styling()

flow_plot <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>%
  mutate(year = as.factor(year))

kge_plot_cy <- ggplot(flow_plot, aes(year, kge, color = model_run)) +
  geom_point(size = 3) +
  scale_color_manual(name="model_run", values=c("grid2011"="seagreen",
                                            "grid2012"="seagreen2",
                                            "grid2013"="steelblue",
                                            "grid2014"='dodgerblue',
                                            "grid2015" = "cyan3",
                                            'grid2016' = 'blue',
                                            'grid2017' = 'blue4',
                                            'grid2018' = 'cadetblue',
                                            'grid2019' = 'lightgreen',
                                            'grid2020' = 'olivedrab',
                                            'gridall' = 'steelblue3',
                                            "statn2011"="maroon",
                                            "statn2012"="deeppink4",
                                            "statn2013"="deeppink",
                                            "statn2014"='purple',
                                            "statn2015" = "mediumorchid1",
                                            'statn2016' = 'indianred',
                                            'statn2017' = 'sienna',
                                            'statn2018' = 'salmon2',
                                            'statn2019' = 'mediumpurple1',
                                            'statn2020' = 'violetred',
                                            'statnall' = 'red')) +
  theme_bw()

ggplotly(kge_plot_cy)

```

### Annual KGE plots

```{r wy plots, echo = F}

plot_kge <- function(calyear) {
  flow_plot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = model_run,
               y = kge)) +
    geom_col() +
    ggtitle(calyear) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
}

name <- levels(flow_plot$year)

lapply(name, plot_kge)

```

# Ensembles

## Hydrograph

```{r ensemble hydrographs, echo = F}

ensem <- flow %>%
  rowwise() %>%
  mutate(grid_ensemble = mean(c(grid_2011, grid_2012,
                                grid_2013, grid_2014,
                                grid_2015, grid_2016,
                                grid_2017, grid_2018,
                                grid_2019, grid_2020)),
         statn_ensemble = mean(c(statn_2011, statn_2012,
                                 statn_2013, statn_2014,
                                 statn_2015, statn_2016,
                                 statn_2017, statn_2018,
                                 statn_2019, statn_2020)),
         all_ensemble = mean(c(grid_2011, grid_2012,
                               grid_2013, grid_2014,
                               grid_2015, grid_2016,
                               grid_2017, grid_2018,
                               grid_2019, grid_2020,
                               statn_2011, statn_2012,
                               statn_2013, statn_2014,
                               statn_2015, statn_2016,
                               statn_2017, statn_2018,
                               statn_2019, statn_2020,
                               grid_all, statn_all))) %>%
  dplyr::select(date, grid_ensemble, statn_ensemble,
         all_ensemble, grid_all,
         statn_all, obs)

ensem_pivot <- ensem %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date)))

q_ensemble <- ggplot(ensem_pivot, aes(date, q_m3s, color = cal_run)) +
  geom_line() +
  scale_color_manual(name="cal_run", values=c('obs'='black',
                                            "grid_ensemble"="maroon",
                                            "statn_ensemble"="darkturquoise",
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4',
                                            "all_ensemble" = "orange")) +
  geom_line(data = bouldcr, aes(date, bouldcr/10), color = 'darkgrey') +
  scale_y_continuous(name = 'Q_m3s',
                    sec.axis = sec_axis(~.*10, name='BoulderCr_cms')) +
  theme_bw()

ggplotly(q_ensemble) %>%
  add_lines(data=bouldcr, x=~date, y=~bouldcr, colors=NULL, yaxis="y2",
              inherit=FALSE, showlegend = FALSE) %>%
  layout(yaxis2 = list(overlaying = "y", side = "right",
                       color = 'darkgray',
                       title = "BoulderCr_cms"),
         legend = list(x = 1.1, y = 0.95))

```

## KGE

### KGE for ensembles (entire period)

```{r ensemble kge, echo = F}

ensem_kge_all <- ensem %>%
  ungroup() %>%
  summarize(grid_ensemble = KGE(grid_ensemble, obs),
            statn_ensemble = KGE(statn_ensemble, obs),
            all_ensemble = KGE(all_ensemble, obs),
            grid_all = KGE(grid_all, obs),
            statn_all = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

kable(ensem_kge_all) %>%
  kable_styling(full_width = F)

```

### Annual KGEs for ensembles

```{r annual kge ensemble, echo = F}

yr_kge_ensem <- ensem %>%
  ungroup() %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(grid_ensemble = KGE(grid_ensemble, obs),
            statn_ensemble = KGE(statn_ensemble, obs),
            all_ensemble = KGE(all_ensemble, obs),
            grid_all = KGE(grid_all, obs),
            statn_all = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

yr_kge_table <- yr_kge_ensem %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>%
  pivot_wider(names_from = 'year', values_from = 'kge')

yr_kge_table %>%
  kable(.) %>%
  kable_styling()

yr_kge_plot <- yr_kge_ensem %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>%
  mutate(year = as.factor(year))

#plot annual kge vals for ensembles
plot_kge_yr_ensem <- function(calyear) {
  yr_kge_plot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = model_run,
               y = kge)) +
    geom_col() +
    ggtitle(calyear) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
}

name <- levels(yr_kge_plot$year)

lapply(name, plot_kge_yr_ensem)

```

### Monthly KGEs for ensembles

```{r monthly kge ensemble, echo = F}

mo_kge <- ensem %>%
  ungroup() %>%
  drop_na() %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(grid_ensemble = KGE(grid_ensemble, obs),
            statn_ensemble = KGE(statn_ensemble, obs),
            all_ensemble = KGE(all_ensemble, obs),
            grid_all = KGE(grid_all, obs),
            statn_all = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

kable(mo_kge) %>%
  kable_styling(full_width = F)

mo_pivot <- mo_kge %>%
  pivot_longer(!month, names_to = 'cal_run', values_to = 'KGE') %>%
  drop_na()

kge_plot_mo <- ggplot(mo_pivot, aes(month, KGE, color = cal_run)) + geom_point(size = 3) +
    scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",
                                            "statn_ensemble"="darkturquoise",
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4',
                                            "all_ensemble" = "orange")) +
  theme_bw()

ggplotly(kge_plot_mo)

```

## RMSE

### Ensemble RMSE for entire period

```{r ensemble rmse, echo = F}

ensem_rmse_all <- ensem %>%
  ungroup() %>%
  summarize(grid_ensemble = rmse(grid_ensemble, obs),
            statn_ensemble = rmse(statn_ensemble, obs),
            all_ensemble = rmse(all_ensemble, obs),
            grid_all = rmse(grid_all, obs),
            statn_all = rmse(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 4)

kable(ensem_rmse_all) %>%
  kable_styling(full_width = F)

```

### Annual RMSE for ensembles

```{r annual rmse ensem, echo = F}

yr_rmse_ensem <- ensem %>%
  ungroup() %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(grid_ensemble = rmse(grid_ensemble, obs),
            statn_ensemble = rmse(statn_ensemble, obs),
            all_ensemble = rmse(all_ensemble, obs),
            grid_all = rmse(grid_all, obs),
            statn_all = rmse(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 4)

yr_rmse_table <- yr_rmse_ensem %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'rmse') %>%
  pivot_wider(names_from = 'year', values_from = 'rmse')

yr_rmse_table %>%
  kable(.) %>%
  kable_styling()

yr_rmse_plot <- yr_rmse_ensem %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'rmse') %>%
  mutate(year = as.factor(year))

#plot annual rmse vals for ensembles
plot_rmse_yr_ensem <- function(calyear) {
  yr_rmse_plot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = model_run,
               y = rmse)) +
    geom_col() +
    ggtitle(calyear) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
}

name <- levels(yr_rmse_plot$year)

lapply(name, plot_rmse_yr_ensem)

```

### Monthly RMSEs for ensembles

```{r month ensem rmse, echo = F}

mo_rmse <- ensem %>%
  ungroup() %>%
  drop_na() %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(grid_ensemble = rmse(grid_ensemble, obs),
            statn_ensemble = rmse(statn_ensemble, obs),
            all_ensemble = rmse(all_ensemble, obs),
            grid_all = rmse(grid_all, obs),
            statn_all = rmse(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 4)

kable(mo_rmse) %>%
  kable_styling(full_width = F)

mo_pivot2 <- mo_rmse %>%
  pivot_longer(!month, names_to = 'cal_run', values_to = 'RMSE') %>%
  drop_na()

rmse_plot_mo <- ggplot(mo_pivot2, aes(month, RMSE, color = cal_run)) + geom_point(size = 3) +
    scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",
                                            "statn_ensemble"="darkturquoise",
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4',
                                            "all_ensemble" = "orange")) +
  theme_bw()

ggplotly(rmse_plot_mo)

```

## NSE

### Ensemble NSE for entire period

```{r nse all, echo = F}

ensem_nse_all <- ensem %>%
  ungroup() %>%
  summarize(grid_ensemble = NSE(grid_ensemble, obs),
            statn_ensemble = NSE(statn_ensemble, obs),
            all_ensemble = NSE(all_ensemble, obs),
            grid_all = NSE(grid_all, obs),
            statn_all = NSE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

kable(ensem_nse_all) %>%
  kable_styling(full_width = F)

```

### Annual NSE for ensembles

```{r annual nse, echo = F}

yr_nse_ensem <- ensem %>%
  ungroup() %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(grid_ensemble = NSE(grid_ensemble, obs),
            statn_ensemble = NSE(statn_ensemble, obs),
            all_ensemble = NSE(all_ensemble, obs),
            grid_all = NSE(grid_all, obs),
            statn_all = NSE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 4)

yr_nse_table <- yr_nse_ensem %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>%
  pivot_wider(names_from = 'year', values_from = 'nse')

yr_nse_table %>%
  kable(.) %>%
  kable_styling()

yr_nse_plot <- yr_nse_ensem %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>%
  mutate(year = as.factor(year))

#plot annual nse vals for ensembles
plot_nse_yr_ensem <- function(calyear) {
  yr_nse_plot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = model_run,
               y = nse)) +
    geom_col() +
    ggtitle(calyear) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
}

name <- levels(yr_nse_plot$year)

lapply(name, plot_nse_yr_ensem)

```

### Monthly NSE for ensembles

```{r monthly rmse, echo =F}

mo_nse <- ensem %>%
  ungroup() %>%
  drop_na() %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(grid_ensemble = NSE(grid_ensemble, obs),
            statn_ensemble = NSE(statn_ensemble, obs),
            all_ensemble = NSE(all_ensemble, obs),
            grid_all = NSE(grid_all, obs),
            statn_all = NSE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 4)

kable(mo_nse) %>%
  kable_styling(full_width = F)

mo_pivot2 <- mo_nse %>%
  pivot_longer(!month, names_to = 'cal_run', values_to = 'nse') %>%
  drop_na()

nse_plot_mo <- ggplot(mo_pivot2, aes(month, nse, color = cal_run)) + geom_point(size = 3) +
  theme_bw() +
  scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon",
                                            "statn_ensemble"="darkturquoise",
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4',
                                            "all_ensemble" = "orange"))

ggplotly(nse_plot_mo)

```

## Cumulative Q

```{r cumq, echo = F}

ensem_cumq <- ensem %>%
  mutate(month = month(date)) %>%
  filter(month %in% c(4:9)) %>%
  dplyr::select(-month) %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  group_by(cal_run, year(date)) %>%
  mutate(cumsum = cumsum(replace_na(q_m3s, 0))) %>%
  rename(year = 4) %>%
  mutate(year = as.factor(year)) %>%
  ungroup()

#plot annual cumulative Q vals for ensembles
plot_cumulq_yr_ensem <- function(calyear) {
  {ensem_cumq %>%
    filter(year == calyear) %>%
    ggplot(aes(x = date,
               y = cumsum,
               color = cal_run)) +
    geom_line() +
    scale_color_manual(name="cal_run", values=c("obs"="black",
                                             "grid_ensemble"="maroon",
                                            "statn_ensemble"="darkturquoise",
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4',
                                            "all_ensemble" = "orange")) +
    ggtitle(calyear) +
    theme_bw() +
    labs(y = 'cumulative Q (m3s)')} %>%
    ggplotly()
}

name <- levels(ensem_cumq$year)

cumulative_plots <- lapply(name, plot_cumulq_yr_ensem)

htmltools::tagList(cumulative_plots)

cumulative_plots

```

# Catchment results

## Precip

```{r catch precip, echo = F}
setwd("/Users/megansears/Documents/Repos/fourmile/catchment_output")

statn <- read_csv('statnall.csv')
grid <- read_csv('gridall.csv')

grid <- grid %>% mutate(site = 'grid')
statn <- statn %>% mutate(site = 'statn')

catchment <- bind_rows(grid, statn) %>%
  mutate(year = as.factor(year(time))) %>%
  group_by(year, site) %>%
  mutate(cumulative_p = cumsum(precip))

#plot cumulative precip by year
plot_cumprecip <- function(calyear) {
  {catchment %>%
    filter(year == calyear) %>%
    ggplot(aes(x = time,
               y = cumulative_p, color = site)) +
    geom_line(size=1) +
    ggtitle(calyear) +
    ylab('Cumulative precip (mm)') +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))} %>%
    ggplotly()
}

name <- levels(catchment$year)

precip_plots <- lapply(name, plot_cumprecip)

htmltools::tagList(precip_plots)

precip_plots

## total P
max_cumulativep <- catchment %>%
  group_by(year, site) %>%
  summarize(total_cumulativep = max(cumulative_p))

ggplot(max_cumulativep, aes(x=year, y=total_cumulativep, color=site)) +
  geom_point(size=3) +
  ggtitle('Total cumulative precip (mm) by year')

```

## Tmean

```{r tmean catch, echo = F}
## tmean
tmean <- catchment %>%
  group_by(year, site) %>%
  summarize(tmean = mean(tmean)) %>%
  ggplot(aes(x = year, y=tmean, color = site)) +
  geom_point(size=3) +
  ggtitle('Average temp by year')

tmean

```

## RD and RG

```{r rd and rg, echo = F, fig.width = 14, fig.asp = 1, out.width = "100%"}

gw_grid <- read_csv(here('catchment_output/gridall_outlet.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid')

gw_statn <- read_csv(here('catchment_output/statnall_outlet.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn')

gw <- bind_rows(gw_grid, gw_statn) %>%
  mutate(year = year(time)) %>%
  rename(date = time)

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRD1_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RD1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRD2_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RD2') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRG1_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RG1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRG2_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol=2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RG2') +
  theme(text = element_text(size = 20))

gw <- gw %>%
  group_by(year, source) %>%
  mutate(RD1cumsum = cumsum(catchmentRD1_w),
         RD2cumsum = cumsum(catchmentRD2_w),
         RG1cumsum = cumsum(catchmentRG1_w),
         RG2cumsum = cumsum(catchmentRG2_w))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RD1cumsum), size =1) +
  facet_wrap(~year, scales = 'free', ncol=2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RD1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RD2cumsum), size = 1) +
  facet_wrap(~year, scales = 'free', ncol=2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RD2') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RG1cumsum), size  = 1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RG1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RG2cumsum), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RG2') +
  theme(text = element_text(size = 20))

```

```{r rd and rg sums, echo = F, fig.width = 22, fig.asp = 1.5, out.width = "100%"}

##################
# all 4 with facets

ggplot(gw, aes(x=date)) +
  geom_line(aes(y=RG2cumsum, color = 'cumulative RG2'), size =1) +
  geom_line(aes(y=RG1cumsum, color = 'cumulative RG1'), size =1) +
  geom_line(aes(y=RD1cumsum, color = 'cumulative RD1'), size =1) +
  geom_line(aes(y=RD2cumsum, color = 'cumulative RD2'), size =1) +
  scale_x_date(labels = date_format("%b")) +
  facet_wrap(~year + source, scales = 'free_x', ncol=2) +
  ggtitle('Cumulative RD and RG') +
  theme(text = element_text(size = 24))

```

```{r gw boxplot, echo = F}

gw_maxcumsum <- gw %>%
  group_by(source, year) %>%
  summarize(yr_max_RD1 = max(RD1cumsum),
            yr_max_RD2 = max(RD2cumsum),
            yr_max_RG1 = max(RG1cumsum),
            yr_max_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  ggtitle('Max annual cumulative sums')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~flow_type, scales = 'free') +
  ggtitle('Max annual cumulative sums')

```

### RD and RG by season

```{r flow season, echo = F, fig.width = 12, fig.asp = 0.75, out.width = "100%"}

#group by season and get max sums by season
flow_season <- gw %>%
  dplyr::select(RD1cumsum, RD2cumsum, RG1cumsum, RG2cumsum, date, year, source) %>%
  mutate(month = month(date)) %>%
  mutate(season = case_when(month %in% c(3,4,5,6) ~ 'snowmelt',
                            month %in% c(7,8,9,10) ~ 'monsoon',
                            month %in% c(11,12,1,2) ~ 'baseflow')) %>%
  group_by(source, year, season) %>%
  summarize(yr_season_RD1 = max(RD1cumsum),
            yr_season_RD2 = max(RD2cumsum),
            yr_season_RG1 = max(RG1cumsum),
            yr_season_RG2 = max(RG2cumsum)) %>%
            pivot_longer(!c(year, source, season), names_to = 'flow_type', values_to = 'flow_cms')

flow_season %>%
  ggplot(aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~season, scales = 'free') +
  ggtitle('Max annual season cumulative sums') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(text = element_text(size = 20))

flow_season %>%
  ggplot(aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~season + flow_type, scales = 'free') +
  ggtitle('Max annual season cumulative sums') +
  theme_bw() +
  theme(text = element_text(size = 20))

```

# HRU results

## Tmean difference

2011-2020 average of daily station-grid Tmean

```{r cumulative tmean for HRUs, echo = F}
#grid first
setwd(here("HRU_output/grid"))

read_ages_grid <- function(dataframe){
  read_csv(dataframe, skip = 3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'grid')

}

#pull in all filenames with HRU in the name
filenames <- list.files(".", pattern="HRU", full.names=F)

#read all the csvs using read_ages function
dataframes <- lapply(filenames, read_ages_grid)

#name all the csvs based on first 6 characters
names(dataframes) <- substr(filenames, 1, 6)

#bind all and make 1 df
hru_grids <- dataframes %>%
  bind_rows()

#statn next
setwd(here("HRU_output/statn"))

read_ages_statn <- function(dataframe){
  read_csv(dataframe, skip = 3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'statn')

}

#pull in all filenames with HRU in the name
filenames <- list.files(".", pattern="HRU", full.names=F)

#read all the csvs using read_ages function
dataframes <- lapply(filenames, read_ages_statn)

#name all the csvs based on first 6 characters
names(dataframes) <- substr(filenames, 1, 6)

#bind all and make 1 df
hru_statn <- dataframes %>%
  bind_rows()

#get rid of a bunch of data
hru_grid_tmean <- hru_grids %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_grid = tmean)  %>%
  group_by(ID) %>%
  summarize(tmean_all_grid = mean(tmean_grid))

hru_statn_tmean <- hru_statn %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_statn = tmean) %>%
  group_by(ID) %>%
  summarize(tmean_all_statn = mean(tmean_statn))

tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('ID')) %>%
  mutate(diff = tmean_all_statn - tmean_all_grid) %>%
  rename(value = ID)

hrus_map_tmean <- left_join(hrus, tmean_hrus, by = 'value')

tmap_mode('view')

locations_t <- locations %>%
  dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
                        'Betasso', 'Niwot_Snotel',
                        'University Camp SNTL'))

tm_shape(hrus_map_tmean) +
  tm_polygons('diff', id = 'value',
              breaks = c(-6,-4,-2,0,2,4,6)) +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

```

## Tmax difference

```{r tmax, echo = F}

tmax_grid <- read_csv(here('HRU_output/reg_tmaxAll_prism.csv'))%>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmax_grid') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmax_mean_grid = mean(tmax_grid))

tmax_statn <- read_csv(here('HRU_output/reg_tmaxAll.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmax_statn') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmax_mean_statn = mean(tmax_statn))

tmax_hrus <- left_join(tmax_grid, tmax_statn, by = c('value')) %>%
  mutate(diff = Tmax_mean_statn - Tmax_mean_grid,
         value = as.numeric(value))

hrus_map_tmax <- left_join(hrus, tmax_hrus, by = 'value')

tmap_mode('view')

tm_shape(hrus_map_tmax) +
  tm_polygons('diff', id = 'value',
              breaks = c(-6,-4,-2,0,2,4,6)) +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

```

## Tmax monthly average

```{r tmax month, echo = F}

tmax_grid_mo <- read_csv(here('HRU_output/reg_tmaxAll_prism.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmax_grid') %>%
  mutate(time = dmy(time),
         month = month(time)) %>%
  filter(month %in% c(3:6)) %>%
  group_by(value, month) %>%
  summarize(avg_tmax_grid = mean(tmax_grid))

tmax_statn_mo <- read_csv(here('HRU_output/reg_tmaxAll.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmax_statn') %>%
  mutate(time = dmy(time),
         month = month(time)) %>%
  filter(month %in% c(3:6)) %>%
  group_by(value, month) %>%
  summarize(avg_tmax_statn = mean(tmax_statn))

tmax_hrus_mo <- left_join(tmax_grid_mo, tmax_statn_mo, by = c('value', 'month')) %>%
  mutate(mo_tmax_diff = avg_tmax_statn - avg_tmax_grid,
         value = as.numeric(value))

tmap_mode('plot')

#march
tmax_hrus_mar<- tmax_hrus_mo %>%
  filter(month == 3)

hrus_map_tmax_mar <- left_join(hrus, tmax_hrus_mar, by = 'value')

mar_tmax <- tm_shape(hrus_map_tmax_mar) +
  #tm_polygons('mo_tmax_diff', id = 'value') +
  tm_fill('mo_tmax_diff', id = 'value',
          breaks = c(-6,-4,-2,0,2,4,6)) +
  tm_borders() +
  tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'March avg tmax',
            legend.outside = T)

#april
tmax_hrus_apr<- tmax_hrus_mo %>%
  filter(month == 4)

hrus_map_tmax_apr <- left_join(hrus, tmax_hrus_apr, by = 'value')

apr_tmax <- tm_shape(hrus_map_tmax_apr) +
  tm_fill('mo_tmax_diff', id = 'value',
          breaks = c(-6,-4,-2,0,2,4,6)) +
  tm_borders() +
  tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'April avg tmax',
            legend.outside = T)

#may
tmax_hrus_may<- tmax_hrus_mo %>%
  filter(month == 5)

hrus_map_tmax_may <- left_join(hrus, tmax_hrus_may, by = 'value')

may_tmax <- tm_shape(hrus_map_tmax_may) +
  tm_fill('mo_tmax_diff', id = 'value',
         breaks = c(-6,-4,-2,0,2,4,6)) +
  tm_borders() +
tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'May avg tmax',
            legend.outside = T)

#june
tmax_hrus_jun<- tmax_hrus_mo %>%
  filter(month == 6)

hrus_map_tmax_jun <- left_join(hrus, tmax_hrus_jun, by = 'value')

jun_tmax <- tm_shape(hrus_map_tmax_jun) +
  tm_fill('mo_tmax_diff', id = 'value',
         breaks = c(-6,-4,-2,0,2,4,6)) +
  tm_borders() +
tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'June avg tmax',
            legend.outside = T)

tmap_arrange(mar_tmax, apr_tmax, may_tmax, jun_tmax)

```

## Tmin difference

```{r tmin, echo = F}

tmin_grid <- read_csv(here('HRU_output/reg_tminAll_prism.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmin_grid') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmin_mean_grid = mean(tmin_grid))

tmin_statn <- read_csv(here('HRU_output/reg_tminAll.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmin_statn') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmin_mean_statn = mean(tmin_statn))

tmin_hrus <- left_join(tmin_grid, tmin_statn, by = c('value')) %>%
  mutate(diff = Tmin_mean_statn - Tmin_mean_grid,
         value = as.numeric(value))

hrus_map_tmin <- left_join(hrus, tmin_hrus, by = 'value')

tmap_mode('view')

tm_shape(hrus_map_tmin) +
  tm_polygons('diff', id = 'value',
              breaks = c(-6,-4,-2,0,2,4,6)) +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

```

## Tmin monthly average

```{r tmin month, echo = F}
tmin_grid_mo <- read_csv(here('HRU_output/reg_tminAll_prism.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmin_grid') %>%
  mutate(time = dmy(time),
         month = month(time)) %>%
  filter(month %in% c(3:6)) %>%
  group_by(value, month) %>%
  summarize(avg_tmin_grid = mean(tmin_grid))

tmin_statn_mo <- read_csv(here('HRU_output/reg_tminAll.csv')) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value',
               values_to = 'tmin_statn') %>%
  mutate(time = dmy(time),
         month = month(time)) %>%
  filter(month %in% c(3:6)) %>%
  group_by(value, month) %>%
  summarize(avg_tmin_statn = mean(tmin_statn))

tmin_hrus_mo <- left_join(tmin_grid_mo, tmin_statn_mo, by = c('value', 'month')) %>%
  mutate(mo_tmin_diff = avg_tmin_statn - avg_tmin_grid,
         value = as.numeric(value))

tmap_mode('plot')

#march
tmin_hrus_mar<- tmin_hrus_mo %>%
  filter(month == 3)

hrus_map_tmin_mar <- left_join(hrus, tmin_hrus_mar, by = 'value')

mar_tmin <- tm_shape(hrus_map_tmin_mar) +
  #tm_polygons('mo_tmin_diff', id = 'value') +
  tm_fill('mo_tmin_diff', id = 'value',
          breaks = c(-5,-2.5,0,2.5,5)) +
  tm_borders() +
  tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'March avg tmin',
            legend.outside = T)

#april
tmin_hrus_apr<- tmin_hrus_mo %>%
  filter(month == 4)

hrus_map_tmin_apr <- left_join(hrus, tmin_hrus_apr, by = 'value')

apr_tmin <- tm_shape(hrus_map_tmin_apr) +
  tm_fill('mo_tmin_diff', id = 'value',
          breaks = c(-5,-2.5,0,2.5,5)) +
  tm_borders() +
  tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'April avg tmin',
            legend.outside = T)

#may
tmin_hrus_may<- tmin_hrus_mo %>%
  filter(month == 5)

hrus_map_tmin_may <- left_join(hrus, tmin_hrus_may, by = 'value')

may_tmin <- tm_shape(hrus_map_tmin_may) +
  tm_fill('mo_tmin_diff', id = 'value',
         breaks = c(-5,-2.5,0,2.5,5)) +
  tm_borders() +
tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'May avg tmin',
            legend.outside = T)

#june
tmin_hrus_jun<- tmin_hrus_mo %>%
  filter(month == 6)

hrus_map_tmin_jun <- left_join(hrus, tmin_hrus_jun, by = 'value')

jun_tmin <- tm_shape(hrus_map_tmin_jun) +
  tm_fill('mo_tmin_diff', id = 'value',
          breaks = c(-5,-2.5,0,2.5,5)) +
  tm_borders() +
tm_shape(locations_t) +
  tm_symbols(col='blue', size = 0.25) +
  tm_layout(title = 'June avg tmin',
            legend.outside = T)

tmap_arrange(mar_tmin, apr_tmin, may_tmin, jun_tmin)

```

## Cumulative Precip difference

2011-2020 average of daily station-grid precip for DOY 365 (or 366)

```{r cum precip, echo = F}
#get rid of a bunch of data
hru_grid_precip <- hru_grids %>%
  dplyr::select(c(time, ID, precip)) %>%
  rename(precip_grid = precip)

hru_statn_precip <- hru_statn %>%
  dplyr::select(c(time, ID, precip)) %>%
  rename(precip_statn = precip)

precip_hrus <- left_join(hru_grid_precip, hru_statn_precip, by = c('time', 'ID')) %>%
  mutate(diff = precip_statn - precip_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>%
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(precip_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

hrus_map_precip <- left_join(hrus, precip_hrus, by = 'value')

tmap_mode('view')

locations_p <- locations %>%
  filter(type == 'precip')

tm_shape(hrus_map_precip) +
  tm_polygons('precip_avg_cumdiff', midpoint = NA, palette = 'YlGnBu', id = 'value') +
tm_shape(locations_p) +
  tm_symbols(col='blue', size=4)

```

## Solar radiation difference

2011-2020 average of daily station-grid solar rad

```{r cum sol rad, echo = F}

#get rid of a bunch of data
hru_grid_solrad <- hru_grids %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_grid = solRad)  %>%
  group_by(ID) %>%
  summarize(solrad_all_grid = mean(solrad_grid))

hru_statn_solrad <- hru_statn %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_statn = solRad) %>%
  group_by(ID) %>%
  summarize(solrad_all_statn = mean(solrad_statn))

solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('ID')) %>%
  mutate(diff = solrad_all_statn - solrad_all_grid) %>%
  rename(value = ID)

hrus_map_solrad <- left_join(hrus, solrad_hrus, by = 'value')

tmap_mode('view')

locations_solrad <- locations %>%
  dplyr::filter(name %in% c('Gordon_South',
                        'Betasso', 'Niwot_Snotel'))

tm_shape(hrus_map_solrad) +
  tm_polygons('diff', midpoint = NA, palette = 'YlOrRd', id = 'value') +
tm_shape(locations_solrad) +
  tm_symbols(col='blue', size=4)

```

## RH difference

```{r rh, echo = F}

#get rid of a bunch of data
hru_grid_rh <- hru_grids %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_grid = rhum)  %>%
  group_by(ID) %>%
  summarize(rh_all_grid = mean(rh_grid))

hru_statn_rh <- hru_statn %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_statn = rhum) %>%
  group_by(ID) %>%
  summarize(rh_all_statn = mean(rh_statn))

rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('ID')) %>%
  mutate(diff = rh_all_statn - rh_all_grid) %>%
  rename(value = ID)

hrus_map_rh <- left_join(hrus, rh_hrus, by = 'value')

tmap_mode('view')

locations_rh <- locations %>%
  dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
                        'Betasso', 'Niwot_Snotel'))

tm_shape(hrus_map_rh) +
  tm_polygons('diff', midpoint = NA, id = 'value') +
tm_shape(locations_rh) +
  tm_symbols(col='blue', size=4)

```

## Average soil sat comparison

```{r soilsat, echo = F}

#get rid of a bunch of data
hru_grid_soilsat <- hru_grids %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_grid = soilSat)  %>%
  group_by(ID) %>%
  summarize(soilsat_all_grid = mean(soilsat_grid))

hru_statn_soilsat <- hru_statn %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_statn = soilSat) %>%
  group_by(ID) %>%
  summarize(soilsat_all_statn = mean(soilsat_statn))

soilsat_hrus <- left_join(hru_grid_soilsat, hru_statn_soilsat, by = c('ID')) %>%
  rename(value = ID)

hrus_map_soilsat <- left_join(hrus, soilsat_hrus, by = 'value')

tmap_mode('plot')

soilavg_grid <- tm_shape(hrus_map_soilsat) +
  tm_polygons('soilsat_all_grid', midpoint = NA, id = 'value',
              breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average soil sat')

soilavg_statn <- tm_shape(hrus_map_soilsat) +
  tm_polygons('soilsat_all_statn', midpoint = NA, id = 'value',
              breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3)) +
  tm_legend(legend.outside = T)

tmap_arrange(soilavg_grid, soilavg_statn)

```

## Average max soil sat comparison

```{r max soilsat, echo = F}

#get rid of a bunch of data
hru_grid_soilsat_max <- hru_grids %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_grid = soilSat)  %>%
  mutate(year = year(time)) %>%
  group_by(ID, year) %>%
  summarize(soilsat_yrmax_grid = max(soilsat_grid)) %>%
  summarize(avg_max_grid = mean(soilsat_yrmax_grid))

hru_statn_soilsat_max <- hru_statn %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_statn = soilSat) %>%
  mutate(year = year(time)) %>%
  group_by(ID, year) %>%
  summarize(soilsat_yrmax_statn = max(soilsat_statn)) %>%
  summarize(avg_max_statn = mean(soilsat_yrmax_statn))

soilsat_hrus_max <- left_join(hru_grid_soilsat_max, hru_statn_soilsat_max, by = c('ID')) %>%
  rename(value = ID)

hrus_map_soilsat_max <- left_join(hrus, soilsat_hrus_max, by = 'value')

tmap_mode('plot')

soilavg_grid_max <- tm_shape(hrus_map_soilsat_max) +
  tm_polygons('avg_max_grid', midpoint = NA, id = 'value',
              breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average max soil sat',
            title.position = c('left', 'top'))

soilavg_statn_max <- tm_shape(hrus_map_soilsat_max) +
  tm_polygons('avg_max_statn', midpoint = NA, id = 'value',
              breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7)) +
  tm_legend(legend.outside = T)

tmap_arrange(soilavg_grid_max, soilavg_statn_max)

```

## Average ET comparison

```{r et, echo = F}
#get rid of a bunch of data
hru_grid_ET <- hru_grids %>%
  dplyr::select(c(time, ID, actET)) %>%
  rename(ET_grid = actET)  %>%
  group_by(ID) %>%
  summarize(ET_all_grid = mean(ET_grid))

hru_statn_ET <- hru_statn %>%
  dplyr::select(c(time, ID, actET)) %>%
  rename(ET_statn = actET) %>%
  group_by(ID) %>%
  summarize(ET_all_statn = mean(ET_statn))

ET_hrus <- left_join(hru_grid_ET, hru_statn_ET, by = c('ID')) %>%
  rename(value = ID)

hrus_map_ET <- left_join(hrus, ET_hrus, by = 'value')

tmap_mode('plot')

ETavg_grid <- tm_shape(hrus_map_ET) +
  tm_polygons('ET_all_grid', midpoint = NA, id = 'value',
              breaks = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average ET')

ETavg_statn <- tm_shape(hrus_map_ET) +
  tm_polygons('ET_all_statn', midpoint = NA, id = 'value',
             breaks = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) +
  tm_legend(legend.outside = T)

tmap_arrange(ETavg_grid, ETavg_statn)

```

## Average ET difference

```{r et diff, echo = F}

ET_hrus <- ET_hrus %>%
  mutate(diff = ET_all_statn - ET_all_grid)

hrus_map_et_diff <- left_join(hrus, ET_hrus, by = 'value')

tmap_mode('plot')

tm_shape(hrus_map_et_diff) +
  tm_polygons('diff', midpoint = NA, id = 'value')  +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average ET difference')

```

## Average LAI comparison

```{r lai, echo = F}

#get rid of a bunch of data
hru_grid_LAI <- hru_grids %>%
  dplyr::select(c(time, ID, LAI)) %>%
  rename(LAI_grid = LAI)  %>%
  group_by(ID) %>%
  summarize(LAI_all_grid = mean(LAI_grid))

hru_statn_LAI <- hru_statn %>%
  dplyr::select(c(time, ID, LAI)) %>%
  rename(LAI_statn = LAI) %>%
  group_by(ID) %>%
  summarize(LAI_all_statn = mean(LAI_statn))

LAI_hrus <- left_join(hru_grid_LAI, hru_statn_LAI, by = c('ID')) %>%
  rename(value = ID)

hrus_map_LAI <- left_join(hrus, LAI_hrus, by = 'value')

tmap_mode('view')

LAIavg_grid <- tm_shape(hrus_map_LAI) +
  tm_polygons('LAI_all_grid', midpoint = NA, id = 'value',
              breaks = c(1,2,3,4,5)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average LAI')

LAIavg_statn <- tm_shape(hrus_map_LAI) +
  tm_polygons('LAI_all_statn', midpoint = NA, id = 'value',
              breaks = c(1,2,3,4,5)) +
  tm_legend(legend.outside = T)

tmap_arrange(LAIavg_grid, LAIavg_statn)

```