### 2024-06-28 sept2013_edits.Rmd
2013 flood edits

---
title: "2013 flood edits"
author: "Megan Sears"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F, 
  warning  = F)

library(tidyverse)
library(plotly)
library(lubridate)
library(knitr)
library(kableExtra)
library(hydroGOF)
library(here)

```

```{r, include = F}

setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_model/sept2013_edits')

kge_edits2013 <- read_csv('cal_KGE.csv')

```

- grid_all25 = grid_all (no years missing) but has 25 cms on 9/12/13<br>
- grid_allNA = grid_all (no years missing) but has NA on 9/12/13<br>
- grid_all = grid_all (no years missing) - original grid all<br>
- statn_all25 = statn_all (no years missing) but has 25 cms on 9/12/13<br>
- statn_allNA = statn_all (no years missing) but has NA on 9/12/13<br>
- statn_all = statn_all (no years missing) - original statn all<br>

# Parameters

## KGE for calibrations

```{r look at kge, echo = F}

kge_plot <- ggplot(kge_edits2013, aes(cal_run, KGE, shape = flood_edits, color = flood_edits)) + geom_point(size = 3) +
  theme_bw()

ggplotly(kge_plot)

```

## Unburned params

```{r, echo = F, message = F, warning = F, include = F}

setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_model/sept2013_edits/params')

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 12)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

```

```{r params, echo = FALSE, message = FALSE, warning = FALSE}

# function to read files, remove top rows, paste site name
read_cals <- function(df, name){
    df <- df %>%
      slice(-c(1:9, 146:277)) %>%
      separate(Parameter, c("Parameter", "value"), sep = ",") %>%
      drop_na() %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)

}

grid_all <- read_cals(grid_all_cal, name = 'grid_all')
grid_all25 <- read_cals(grid_all25_c, name = 'grid_all25')
grid_allNA <- read_cals(gril_allNA_c, name = 'grid_allNA')

grid <- bind_rows(grid_all25, grid_allNA,
                  grid_all) %>%
  mutate(input_data = 'grid') %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

statn_all <- read_cals(statn_all_ca, name = 'statn_all')
statn_all25 <- read_cals(statn_all25_, name = 'statn_all25')
statn_allNA <- read_cals(statn_allNA_, name = 'statn_allNA')

statn <- bind_rows(statn_all25, statn_allNA,
                  statn_all) %>%
  mutate(input_data = 'statn') %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

params <- bind_rows(statn, grid)

params <- params %>%
  filter(!param %in% c('angstrom_a', 'angstrom_b', 
                      'Beta_min', 
                      'Beta_rsd', "Beta_trans", 'clat', 
                      'delayNitrification',
                      'denitrificationRateCoefficient', 
                      'denitrificationSoilSaturationThreshold',
                      'deposition_factor', 'drrad', 
                      'drspac', 'initLPS', 'initRG1',
                      'initRG2', 'Ksink', 'LExCoef', 
                      'longTZ', 'N_concRG1', 'N_concRG2', 'N_delay_RG1',
                      'N_delay_RG2', 
                      'nitrificationSoilTemperatureThreshold', 
                      'nitrificationSurfaceTemperatureThreshold',
                      'opti', 'piadin', 'rootfactor', 'sceno', 
                      'soilImpGT80', 'soilImpLT80',
                      'soilLinRed', 'soilMaxPerc', 'temp_lag')) %>%
    mutate(value = as.numeric(value),
         param = as.factor(param))

#function to plot all params
plot_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = cal,
               y = value,
               color = cal)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plot_params)

```

## Burn params

```{r echo = FALSE, message = FALSE, warning = FALSE, include = F}

rm(list = ls())

setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_model/sept2013_edits/params")

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 10)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

```

```{r, echo = F}

# read in all the grid burn params
read_cals_burn <- function(df, name){
    df <- df %>%
      slice(-c(1:274)) %>%
      separate_wider_delim(Parameter, 
                       delim = ',',
                       names = c('soilID',
                                 'luca_a_rain',
                                 'luca_a_snow',
                                 'luca_soilMaxInfSummer',
                                 'luca_soilMaxDPS',
                                 'luca_lagSurfaceRunoff',
                                 'luca_soilDistMPSLPS')) %>%
      slice(-1) %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)
}

grid_all_burn <- read_cals_burn(grid_all_c, name = 'grid_all')
grid_all25_burn <- read_cals_burn(grid_all25, name = 'grid_all25')
grid_allNA_burn <- read_cals_burn(gril_allNA, name = 'grid_allNA')

grid_burn <- bind_rows(grid_all25_burn, grid_allNA_burn,
                  grid_all_burn) %>%
  mutate(input_data = 'grid')

# now do statn burn params
statn_all_burn <- read_cals_burn(statn_all_, name = 'statn_all')
statn_all25_burn <- read_cals_burn(statn_all2, name = 'statn_all25')
statn_allNA_burn <- read_cals_burn(statn_allN, name = 'statn_allNA')

statn_burn <- bind_rows(statn_all25_burn, statn_allNA_burn,
                  statn_all_burn) %>%
  mutate(input_data = 'statn')

burn_params <- bind_rows(grid_burn, statn_burn) %>%
  ungroup() %>%
  mutate(across(c(luca_a_rain:luca_soilDistMPSLPS), as.numeric)) %>%
  group_by(cal) %>%
  mutate(across(c(luca_a_rain:luca_soilDistMPSLPS), mean)) %>%
  filter(soilID == 22) %>%
  dplyr::select(-soilID) %>% 
  pivot_longer(!c(cal, input_data), 
               names_to = 'params',
               values_to = 'value') %>% 
  mutate(params = str_remove(params, 'luca_')) %>%
  mutate(params = as.factor(params),
         burn = 'yes')

```

```{r, echo = F}

plot_burnparams <- function(param_name) {
  burn_params %>%
    filter(params == param_name) %>%
    ggplot(aes(y = value,
               x = cal,
               color = cal)) +
    geom_point(size = 3) +
    ggtitle(param_name) +
    theme_bw()
}

name2 <- levels(burn_params$params)

lapply(name2, plot_burnparams)

```

# Outlet

## Hydrographs

```{r, echo = F}

setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_model/sept2013_edits')

flow <- read_csv('streamflow.csv') %>%
  replace(.==-9999, NA) %>%
  rename(date = time) %>%
  mutate(date = mdy(date))

flow_pivot <- flow %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date)))

q <- ggplot(flow_pivot, aes(date, q_m3s, color = cal_run)) +
  geom_line() +
  scale_color_manual(name="cal_run", values=c('obs' = 'black',
                                            "grid_allNA"="seagreen",
                                            'grid_all25' = 'blue',
                                            'grid_all' = 'steelblue3',
                                            "statn_all25"="maroon",
                                            "statn_allNA"='purple',
                                            'statn_all' = 'red'))

ggplotly(q)

```

## Q KGE

```{r, echo = F}

flow_kge_all <- flow %>%
  drop_na() %>%
  summarize(grid_allNA = KGE(grid_allNA, obs),
            grid_all25 = KGE(grid_all25, obs),
            gridall = KGE(grid_all, obs),
            statn_all25 = KGE(statn_all25, obs),
            statn_allNA = KGE(statn_allNA, obs),
            statnall = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_kge_all <- flow_kge_all %>%
  pivot_longer(everything()) %>%
  arrange(value)

kable(flow_kge_all) %>%
  kable_styling(full_width = F)

```

## Q percent bias

```{r, echo = F}

flow_pbias_all <- flow %>%
  drop_na() %>%
  summarize(grid_allNA = pbias(grid_allNA, obs),
            grid_all25 = pbias(grid_all25, obs),
            gridall = pbias(grid_all, obs),
            statn_all25 = pbias(statn_all25, obs),
            statn_allNA = pbias(statn_allNA, obs),
            statnall = pbias(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_pbias_all <- flow_pbias_all %>%
  pivot_longer(everything()) %>%
  arrange(value)

kable(flow_pbias_all) %>%
  kable_styling(full_width = F)

```