### 2024-07-16 fourmile_outputs_2013edits.Rmd
fixed spatial figs

---
title: "Fourmile Calibrations and Output"
author: "Megan Sears"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F, 
  warning  = F)

library(tidyverse)
library(plotly)
library(lubridate)
library(knitr)
library(kableExtra)
library(dataRetrieval)
library(hydroGOF)
library(sp)
library(raster)
library(sf)
library(mapview)
library(here)
library(scales)
library(dataRetrieval)
library(tmap)
library(ggpubr)
library(gridExtra)
library(stringr)

```

```{r load in files, include = F}

setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates')

#kge <- read_csv('calruns_KGE.csv') # skip for now - come back if needed

```

```{r look at kge, echo = F}

# kge_plot <- ggplot(kge, aes(year_missing, KGE, shape = input_data, color = input_data)) + geom_point(size = 3) +
#   theme_bw()
# 
# ggplotly(kge_plot)

```

# Parameters

## Scatter plots

```{r read in params, echo = FALSE, message = FALSE, warning = FALSE, include = F}

setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/cals")

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 10)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

```

```{r params, echo = FALSE, message = FALSE, warning = FALSE}

# function to read files, remove top rows, paste site name
read_cals <- function(df, name){
    df <- df %>%
      slice(-c(1:9, 146:277)) %>%
      separate(Parameter, c("Parameter", "value"), sep = ",") %>%
      drop_na() %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)

}

grid_2011 <- read_cals(grid_2011., name = 'grid_2011')
grid_2012 <- read_cals(grid_2012., name = 'grid_2012')
grid_2013 <- read_cals(grid_2013., name = 'grid_2013')
grid_2014 <- read_cals(grid_2014., name = 'grid_2014')
grid_2015 <- read_cals(grid_2015., name = 'grid_2015')
grid_2016 <- read_cals(grid_2016., name = 'grid_2016')
grid_2017 <- read_cals(grid_2017., name = 'grid_2017')
grid_2018 <- read_cals(grid_2018., name = 'grid_2018')
grid_2019 <- read_cals(grid_2019., name = 'grid_2019')
grid_2020 <- read_cals(grid_2020., name = 'grid_2020')
grid_all <- read_cals(grid_all.c, name = 'grid_all')

grid <- bind_rows(grid_2011, grid_2012, grid_2013, grid_2014, grid_2015,
                  grid_2016, grid_2017, grid_2018, grid_2019, grid_2020,
                  grid_all) %>%
  mutate(input_data = 'grid') %>%
  mutate(year_missing_almost = gsub('grid_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost)) %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

statn_2011 <- read_cals(statn_2011, name = 'statn_2011')
statn_2012 <- read_cals(statn_2012, name = 'statn_2012')
statn_2013 <- read_cals(statn_2013, name = 'statn_2013')
statn_2014 <- read_cals(statn_2014, name = 'statn_2014')
statn_2015 <- read_cals(statn_2015, name = 'statn_2015')
statn_2016 <- read_cals(statn_2016, name = 'statn_2016')
statn_2017 <- read_cals(statn_2017, name = 'statn_2017')
statn_2018 <- read_cals(statn_2018, name = 'statn_2018')
statn_2019 <- read_cals(statn_2019, name = 'statn_2019')
statn_2020 <- read_cals(statn_2020, name = 'statn_2020')
statn_all <- read_cals(statn_all., name = 'statn_all')

statn <- bind_rows(statn_2011, statn_2012, statn_2013, statn_2014, statn_2015,
                  statn_2016, statn_2017, statn_2018, statn_2019, statn_2020,
                  statn_all) %>%
  mutate(input_data = 'statn') %>%
  mutate(year_missing_almost = gsub('statn_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost)) %>%
  rename(param = Parameter) %>%
  filter(!param == 'locGrw',
         !param == 'tempRes')

params <- bind_rows(statn, grid)

params <- params %>%
  filter(!param %in% c('angstrom_a', 'angstrom_b', 
                      'Beta_min', 
                      'Beta_rsd', "Beta_trans", 'clat', 
                      'delayNitrification',
                      'denitrificationRateCoefficient', 
                      'denitrificationSoilSaturationThreshold',
                      'deposition_factor', 'drrad', 
                      'drspac', 'initLPS', 'initRG1',
                      'initRG2', 'Ksink', 'LExCoef', 
                      'longTZ', 'N_concRG1', 'N_concRG2', 'N_delay_RG1',
                      'N_delay_RG2', 
                      'nitrificationSoilTemperatureThreshold', 
                      'nitrificationSurfaceTemperatureThreshold',
                      'opti', 'piadin', 'rootfactor', 'sceno', 
                      'soilImpGT80', 'soilImpLT80',
                      'soilLinRed', 'soilMaxPerc', 'temp_lag')) %>%
    mutate(value = as.numeric(value),
         param = as.factor(param))

#function to plot all params
plot_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = year_missing,
               y = value,
               shape = input_data, 
               color = input_data)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plot_params)

#write.csv(params, 'params.csv')

# find min and max of grid and statn param for sensitivity analysis
# grid_minmax <- grid %>%
#   group_by(param) %>%
#   summarize(min = as.numeric(min(value)),
#             max = as.numeric(max(value))) %>%
#   mutate(input = 'grid_unburn')
# 
# statn_minmax <- statn %>%
#   group_by(param) %>%
#   summarize(min = as.numeric(min(value)),
#             max = as.numeric(max(value))) %>%
#   mutate(input = 'statn_unburn')
# 
# minmax_burn <- burn_params %>%
#   group_by(params, input_data) %>%
#   summarize(min = min(value),
#             max = max(value)) %>%
#   mutate(input = paste(input_data, '_burn')) %>%
#   dplyr::select(-input_data) %>%
#   rename(param = params) %>%
#   mutate(param = as.character(param))
# 
# minmax_params <- bind_rows(grid_minmax, statn_minmax, minmax_burn)
# write.csv(minmax_params, 'min_max_params.csv')

```

## Density plots

```{r density plots, echo = FALSE, message = FALSE, warning = FALSE}

mean_params <- params %>%
  group_by(param, input_data) %>%
  summarize(mean = mean(value))

plotdens_params <- function(param_name) {
  params %>%
    filter(param == param_name) %>%
    ggplot(aes(x = value,
               fill = input_data, 
               color = input_data)) + 
    geom_density(alpha=0.5) +
    geom_vline(data=filter(mean_params, param == param_name), 
               aes(xintercept=mean, color=input_data),
             linetype="dashed", linewidth=0.75) +
    ggtitle(param_name) +
    theme_bw()
}

name <- levels(params$param)

lapply(name, plotdens_params)

```

## Burn parameters

```{r echo = FALSE, message = FALSE, warning = FALSE, include = F}

rm(statn_2011, statn_2012, statn_2013, statn_2014, statn_2015,
                  statn_2016, statn_2017, statn_2018, statn_2019, statn_2020,
                  statn_all., statn_all)

setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/cals")

#pull in all filenames with composite
filenames <- list.files('.', pattern=".csv", full.names=F)

#read all the csvs
dataframes <- lapply(filenames, read_csv)

#name all the csv
names(dataframes) <- substr(filenames, 1, 10)

#extract csvs from list to global env
lapply(names(dataframes), function(x) assign(x, dataframes[[x]], envir = .GlobalEnv))
rm(dataframes) #remove dataframes list

```

```{r echo = FALSE, message = FALSE, warning = FALSE}

# read in all the grid burn params
read_cals_burn <- function(df, name){
    df <- df %>%
      slice(-c(1:274)) %>%
      separate_wider_delim(Parameter, 
                       delim = ',',
                       names = c('soilID',
                                 'luca_a_rain',
                                 'luca_a_snow',
                                 'luca_soilMaxInfSummer',
                                 'luca_soilMaxDPS',
                                 'luca_lagSurfaceRunoff',
                                 'luca_soilDistMPSLPS')) %>%
      slice(-1) %>%
      dplyr::select(-1) %>% 
      mutate(cal = name)
}

grid_2011_burn <- read_cals_burn(grid_2011., name = 'grid_2011')
grid_2012_burn <- read_cals_burn(grid_2012., name = 'grid_2012')
grid_2013_burn <- read_cals_burn(grid_2013., name = 'grid_2013')
grid_2014_burn <- read_cals_burn(grid_2014., name = 'grid_2014')
grid_2015_burn <- read_cals_burn(grid_2015., name = 'grid_2015')
grid_2016_burn <- read_cals_burn(grid_2016., name = 'grid_2016')
grid_2017_burn <- read_cals_burn(grid_2017., name = 'grid_2017')
grid_2018_burn <- read_cals_burn(grid_2018., name = 'grid_2018')
grid_2019_burn <- read_cals_burn(grid_2019., name = 'grid_2019')
grid_2020_burn <- read_cals_burn(grid_2020., name = 'grid_2020')
grid_all_burn <- read_cals_burn(grid_all.c, name = 'grid_all')

grid_burn <- bind_rows(grid_2011_burn, 
                  grid_2012_burn, 
                  grid_2013_burn, 
                  grid_2014_burn, 
                  grid_2015_burn,
                  grid_2016_burn, 
                  grid_2017_burn, 
                  grid_2018_burn, 
                  grid_2019_burn, 
                  grid_2020_burn,
                  grid_all_burn) %>%
  mutate(input_data = 'grid') %>%
  mutate(year_missing_almost = gsub('grid_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost))

# now do statn burn params
statn_2011_burn <- read_cals_burn(statn_2011, name = 'statn_2011')
statn_2012_burn <- read_cals_burn(statn_2012, name = 'statn_2012')
statn_2013_burn <- read_cals_burn(statn_2013, name = 'statn_2013')
statn_2014_burn <- read_cals_burn(statn_2014, name = 'statn_2014')
statn_2015_burn <- read_cals_burn(statn_2015, name = 'statn_2015')
statn_2016_burn <- read_cals_burn(statn_2016, name = 'statn_2016')
statn_2017_burn <- read_cals_burn(statn_2017, name = 'statn_2017')
statn_2018_burn <- read_cals_burn(statn_2018, name = 'statn_2018')
statn_2019_burn <- read_cals_burn(statn_2019, name = 'statn_2019')
statn_2020_burn <- read_cals_burn(statn_2020, name = 'statn_2020')
statn_all_burn <- read_cals_burn(statn_all., name = 'statn_all')

statn_burn <- bind_rows(statn_2011_burn, 
                  statn_2012_burn, 
                  statn_2013_burn, 
                  statn_2014_burn, 
                  statn_2015_burn,
                  statn_2016_burn, 
                  statn_2017_burn, 
                  statn_2018_burn, 
                  statn_2019_burn, 
                  statn_2020_burn,
                  statn_all_burn) %>%
  mutate(input_data = 'statn') %>%
  mutate(year_missing_almost = gsub('statn_', '', cal)) %>%
  mutate(year_missing = gsub('all', 'none', year_missing_almost)) %>%
  dplyr::select(-c(year_missing_almost))

burn_params <- bind_rows(grid_burn, statn_burn) %>%
  ungroup() %>%
  mutate(across(c(luca_a_rain:luca_soilDistMPSLPS), as.numeric)) %>%
  group_by(cal) %>%
  mutate(across(c(luca_a_rain:luca_soilDistMPSLPS), mean)) %>%
  filter(soilID == 22) %>%
  dplyr::select(-soilID) %>% 
  pivot_longer(!c(cal, input_data, year_missing), 
               names_to = 'params',
               values_to = 'value') %>% 
  mutate(params = str_remove(params, 'luca_')) %>%
  mutate(params = as.factor(params),
         burn = 'yes')

```

```{r echo = FALSE, message = FALSE, warning = FALSE}

plot_burnparams <- function(param_name) {
  burn_params %>%
    filter(params == param_name) %>%
    ggplot(aes(y = value,
               x = year_missing,
               color = input_data)) +
    geom_point(size = 3) +
    ggtitle(param_name) +
    theme_bw()
}

name2 <- levels(burn_params$params)

lapply(name2, plot_burnparams)

```

## Combined parameters

The two burned HRU parameters were averaged together.

```{r echo = FALSE, message = FALSE, warning = FALSE}
params2 <- params %>%
  mutate(burn = 'no') %>%
  filter(param %in% c('a_rain',
                        'a_snow',
                        'soilMaxInfSummer',
                        'soilMaxDPS',
                        'lagSurfaceRunoff',
                        'soilDistMPSLPS'))

params_combo <- bind_rows(params2,
                          burn_params %>%
                            rename(param = params))

#function to plot all params
plot_params <- function(param_name) {
  params_combo %>%
    filter(param == param_name) %>%
    ggplot(aes(x = year_missing,
               y = value,
               shape = burn, 
               color = input_data)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

#name <- levels(params_combo$param)

lapply(name2, plot_params)

# boxplots of burned and unburned

ggplot(params_combo, aes(x=burn, y=value, fill=input_data)) +
  geom_boxplot() +
  facet_wrap(~param, scales = 'free_y')

write_csv(params_combo, '/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/params_burn_unburn.csv')

```

## Parameters area weighted

```{r echo = FALSE, message = FALSE, warning = FALSE}

# add area weights to params combo
params_weighted <- params_combo %>%
  mutate(area = if_else(burn == 'no', 48243400,
                        18823000)) %>%
  group_by(param, cal) %>%
  summarize(numerator = sum(value * area),
    denominator = sum(area)) %>%
  mutate(area_weighted_average = numerator / denominator) %>%
  dplyr::select(-c(numerator, denominator)) %>%
  separate(col = cal, into = c("source", "year_missing"), sep = "_")

write_csv(params_weighted, '/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/params_weighted_sept2013_edits.csv')

#function to plot all params
plot_params <- function(param_name) {
  params_weighted %>%
    filter(param == param_name) %>%
    ggplot(aes(x = year_missing,
               y = area_weighted_average,
               color = source)) + 
    geom_point(size = 3) + 
    ggtitle(param_name) +
    theme_bw()
}

#name <- levels(params_combo$param)

lapply(name2, plot_params)

```

# Outlet results

## Hydrograph

```{r hydrographs, echo = FALSE, message = FALSE, warning = FALSE}
setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates')

flow <- read_csv('discharge_all.csv') %>%
  replace(.==-9999, NA) %>%
  rename(date = time) %>%
  mutate(date = mdy(date))

flow_pivot <- flow %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date)))

q <- ggplot(flow_pivot, aes(date, q_m3s, color = cal_run)) +
  geom_line() +
  scale_color_manual(name="cal_run", values=c('obs' = 'black',
                                            "grid_2011"="seagreen",
                                            "grid_2012"="seagreen2",
                                            "grid_2013"="steelblue",
                                            "grid_2014"='dodgerblue',
                                            "grid_2015" = "cyan3",
                                            'grid_2016' = 'blue',
                                            'grid_2017' = 'blue4',
                                            'grid_2018' = 'cadetblue',
                                            'grid_2019' = 'lightgreen',
                                            'grid_2020' = 'olivedrab',
                                            'grid_all' = 'steelblue3',
                                            "statn_2011"="maroon",
                                            "statn_2012"="deeppink4",
                                            "statn_2013"="deeppink",
                                            "statn_2014"='purple',
                                            "statn_2015" = "mediumorchid1",
                                            'statn_2016' = 'indianred',
                                            'statn_2017' = 'sienna',
                                            'statn_2018' = 'salmon2',
                                            'statn_2019' = 'mediumpurple1',
                                            'statn_2020' = 'violetred',
                                            'statn_all' = 'red'))

ggplotly(q)

```

## Cumulative Q

```{r cumq, echo = FALSE, message = FALSE, warning = FALSE}

flow_cumq <- flow_pivot %>%
  mutate(month = month(date)) %>%
  filter(month %in% c(4:9)) %>%
  dplyr::select(-month) %>%
  #pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  group_by(cal_run, year(date)) %>%
  mutate(cumsum = cumsum(replace_na(q_m3s, 0))) %>%
  #rename(year = 4) %>%
  mutate(year = as.factor(year)) %>%
  ungroup()

#plot annual cumulative Q vals for ensembles
plot_cumulq_yr_ <- function(calyear) {
  {flow_cumq %>%
    filter(year == calyear) %>%
    ggplot(aes(x = date,
               y = cumsum,
               color = cal_run)) +
    geom_line() +
scale_color_manual(name="cal_run", values=c('obs' = 'black',
                                            "grid_2011"="seagreen",
                                            "grid_2012"="seagreen2",
                                            "grid_2013"="steelblue",
                                            "grid_2014"='dodgerblue',
                                            "grid_2015" = "cyan3",
                                            'grid_2016' = 'blue',
                                            'grid_2017' = 'blue4',
                                            'grid_2018' = 'cadetblue',
                                            'grid_2019' = 'lightgreen',
                                            'grid_2020' = 'olivedrab',
                                            'grid_all' = 'steelblue3',
                                            "statn_2011"="maroon",
                                            "statn_2012"="deeppink4",
                                            "statn_2013"="deeppink",
                                            "statn_2014"='purple',
                                            "statn_2015" = "mediumorchid1",
                                            'statn_2016' = 'indianred',
                                            'statn_2017' = 'sienna',
                                            'statn_2018' = 'salmon2',
                                            'statn_2019' = 'mediumpurple1',
                                            'statn_2020' = 'violetred',
                                            'statn_all' = 'red')) +
    ggtitle(calyear) +
    theme_bw() +
    labs(y = 'cumulative Q (m3s)')} %>%
    ggplotly()
}

name <- levels(flow_cumq$year)

cumulative_plots <- lapply(name, plot_cumulq_yr_)

htmltools::tagList(cumulative_plots)

cumulative_plots

```

## KGE

### Overall KGE

```{r kge overall, echo = FALSE, message = FALSE, warning = FALSE}

flow_kge_all <- flow %>%
  summarize(grid2011 = KGE(grid_2011, obs),
            grid2012 = KGE(grid_2012, obs),
            grid2013 = KGE(grid_2013, obs),
            grid2014 = KGE(grid_2014, obs),
            grid2015 = KGE(grid_2015, obs),
            grid2016 = KGE(grid_2016, obs),
            grid2017 = KGE(grid_2017, obs),
            grid2018 = KGE(grid_2018, obs),
            grid2019 = KGE(grid_2019, obs),
            grid2020 = KGE(grid_2020, obs),
            gridall = KGE(grid_all, obs),
            statn2011 = KGE(statn_2011, obs),
            statn2012 = KGE(statn_2012, obs),
            statn2013 = KGE(statn_2013, obs),
            statn2014 = KGE(statn_2014, obs),
            statn2015 = KGE(statn_2015, obs),
            statn2016 = KGE(statn_2016, obs),
            statn2017 = KGE(statn_2017, obs),
            statn2018 = KGE(statn_2018, obs),
            statn2019 = KGE(statn_2019, obs),
            statn2020 = KGE(statn_2020, obs),
            statnall = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_kge_all <- flow_kge_all %>%
  pivot_longer(everything()) %>%
  arrange(value)

kable(flow_kge_all) %>%
  kable_styling(full_width = F)

```

### KGE by year

```{r kge by year, echo = FALSE, message = FALSE, warning = FALSE}

flow_year <- flow %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(grid2011 = KGE(grid_2011, obs),
            grid2012 = KGE(grid_2012, obs),
            grid2013 = KGE(grid_2013, obs),
            grid2014 = KGE(grid_2014, obs),
            grid2015 = KGE(grid_2015, obs),
            grid2016 = KGE(grid_2016, obs),
            grid2017 = KGE(grid_2017, obs),
            grid2018 = KGE(grid_2018, obs),
            grid2019 = KGE(grid_2019, obs),
            grid2020 = KGE(grid_2020, obs),
            gridall = KGE(grid_all, obs),
            statn2011 = KGE(statn_2011, obs),
            statn2012 = KGE(statn_2012, obs),
            statn2013 = KGE(statn_2013, obs),
            statn2014 = KGE(statn_2014, obs),
            statn2015 = KGE(statn_2015, obs),
            statn2016 = KGE(statn_2016, obs),
            statn2017 = KGE(statn_2017, obs),
            statn2018 = KGE(statn_2018, obs),
            statn2019 = KGE(statn_2019, obs),
            statn2020 = KGE(statn_2020, obs),
            statnall = KGE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_kge <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>%
  pivot_wider(names_from = 'year', values_from = 'kge')

flow_kge %>%
  kable(.) %>%
  kable_styling()

flow_plot <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>%
  mutate(year = as.factor(year))

ggplot(flow_plot, aes(x = model_run, y = year, fill = kge)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "viridis", name = "KGE", direction = -1) +
  geom_text(aes(label = round(kge, 2)), color = "grey", size = 2.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Year")

```

### Annual KGE plots

```{r wy plots, echo = FALSE, message = FALSE, warning = FALSE}

plot_kge <- function(calyear) {
  flow_plot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = model_run,
               y = kge)) +
    geom_point(size=3) +
    ggtitle(calyear) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
}

name <- levels(flow_plot$year)

lapply(name, plot_kge)

```

## Pbias

### Overall pbias

```{r pbias overall, echo = FALSE, message = FALSE, warning = FALSE}

flow_pbias_all <- flow %>%
  summarize(grid2011 = pbias(grid_2011, obs),
            grid2012 = pbias(grid_2012, obs),
            grid2013 = pbias(grid_2013, obs),
            grid2014 = pbias(grid_2014, obs),
            grid2015 = pbias(grid_2015, obs),
            grid2016 = pbias(grid_2016, obs),
            grid2017 = pbias(grid_2017, obs),
            grid2018 = pbias(grid_2018, obs),
            grid2019 = pbias(grid_2019, obs),
            grid2020 = pbias(grid_2020, obs),
            gridall = pbias(grid_all, obs),
            statn2011 = pbias(statn_2011, obs),
            statn2012 = pbias(statn_2012, obs),
            statn2013 = pbias(statn_2013, obs),
            statn2014 = pbias(statn_2014, obs),
            statn2015 = pbias(statn_2015, obs),
            statn2016 = pbias(statn_2016, obs),
            statn2017 = pbias(statn_2017, obs),
            statn2018 = pbias(statn_2018, obs),
            statn2019 = pbias(statn_2019, obs),
            statn2020 = pbias(statn_2020, obs),
            statnall = pbias(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_pbias_all <- flow_pbias_all %>%
  pivot_longer(everything()) %>%
  arrange(abs(value))

kable(flow_pbias_all) %>%
  kable_styling(full_width = F)

```

### Pbias by year

```{r pbias by year, echo = FALSE, message = FALSE, warning = FALSE}

flow_year <- flow %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(grid2011 = pbias(grid_2011, obs),
            grid2012 = pbias(grid_2012, obs),
            grid2013 = pbias(grid_2013, obs),
            grid2014 = pbias(grid_2014, obs),
            grid2015 = pbias(grid_2015, obs),
            grid2016 = pbias(grid_2016, obs),
            grid2017 = pbias(grid_2017, obs),
            grid2018 = pbias(grid_2018, obs),
            grid2019 = pbias(grid_2019, obs),
            grid2020 = pbias(grid_2020, obs),
            gridall = pbias(grid_all, obs),
            statn2011 = pbias(statn_2011, obs),
            statn2012 = pbias(statn_2012, obs),
            statn2013 = pbias(statn_2013, obs),
            statn2014 = pbias(statn_2014, obs),
            statn2015 = pbias(statn_2015, obs),
            statn2016 = pbias(statn_2016, obs),
            statn2017 = pbias(statn_2017, obs),
            statn2018 = pbias(statn_2018, obs),
            statn2019 = pbias(statn_2019, obs),
            statn2020 = pbias(statn_2020, obs),
            statnall = pbias(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_pbias <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'pbias') %>%
  pivot_wider(names_from = 'year', values_from = 'pbias')

flow_pbias %>%
  kable(.) %>%
  kable_styling()

flow_plot <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'pbias') %>%
  mutate(year = as.factor(year))

ggplot(flow_plot, aes(x = model_run, y = year, fill = pbias)) +
  geom_tile(color = "white") +
   scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Ppias") +
  geom_text(aes(label = round(pbias, 2)), color = "black", size = 2.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Year")

```

### Annual Pbias plots

```{r pbias plots, echo = FALSE, message = FALSE, warning = FALSE}

plot_pbias <- function(calyear) {
  flow_plot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = model_run,
               y = pbias)) +
    geom_point(size=3) +
    ggtitle(calyear) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
}

name <- levels(flow_plot$year)

lapply(name, plot_pbias)

```

<!-- # Ensembles -->

<!-- ## Hydrograph -->

<!-- ```{r ensemble hydrographs, echo = F} -->

<!-- ensem <- flow %>% -->
<!--   rowwise() %>% -->
<!--   mutate(grid_ensemble = mean(c(grid_2011, grid_2012, -->
<!--                                 grid_2013, grid_2014, -->
<!--                                 grid_2015, grid_2016, -->
<!--                                 grid_2017, grid_2018, -->
<!--                                 grid_2019, grid_2020)), -->
<!--          statn_ensemble = mean(c(statn_2011, statn_2012, -->
<!--                                  statn_2013, statn_2014, -->
<!--                                  statn_2015, statn_2016, -->
<!--                                  statn_2017, statn_2018, -->
<!--                                  statn_2019, statn_2020)), -->
<!--          all_ensemble = mean(c(grid_2011, grid_2012, -->
<!--                                grid_2013, grid_2014, -->
<!--                                grid_2015, grid_2016, -->
<!--                                grid_2017, grid_2018, -->
<!--                                grid_2019, grid_2020, -->
<!--                                statn_2011, statn_2012, -->
<!--                                statn_2013, statn_2014, -->
<!--                                statn_2015, statn_2016, -->
<!--                                statn_2017, statn_2018, -->
<!--                                statn_2019, statn_2020, -->
<!--                                grid_all, statn_all))) %>% -->
<!--   dplyr::select(date, grid_ensemble, statn_ensemble, -->
<!--          all_ensemble, grid_all, -->
<!--          statn_all, obs) -->

<!-- ensem_pivot <- ensem %>% -->
<!--   pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>% -->
<!--   mutate(year = as.factor(year(date))) -->

<!-- q_ensemble <- ggplot(ensem_pivot, aes(date, q_m3s, color = cal_run)) + -->
<!--   geom_line() + -->
<!--   scale_color_manual(name="cal_run", values=c('obs'='black', -->
<!--                                             "grid_ensemble"="maroon", -->
<!--                                             "statn_ensemble"="darkturquoise", -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4', -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   scale_y_continuous(name = 'Q_m3s') + -->
<!--   theme_bw() -->

<!-- q_ensemble -->

<!-- ``` -->

<!-- ## KGE -->

<!-- ### KGE for ensembles (entire period) -->

<!-- ```{r ensemble kge, echo = F} -->

<!-- ensem_kge_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = KGE(grid_ensemble, obs), -->
<!--             statn_ensemble = KGE(statn_ensemble, obs), -->
<!--             all_ensemble = KGE(all_ensemble, obs), -->
<!--             grid_all = KGE(grid_all, obs), -->
<!--             statn_all = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- kable(ensem_kge_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual KGEs for ensembles -->

<!-- ```{r annual kge ensemble, echo = F} -->

<!-- yr_kge_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = KGE(grid_ensemble, obs), -->
<!--             statn_ensemble = KGE(statn_ensemble, obs), -->
<!--             all_ensemble = KGE(all_ensemble, obs), -->
<!--             grid_all = KGE(grid_all, obs), -->
<!--             statn_all = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- yr_kge_table <- yr_kge_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'kge') -->

<!-- yr_kge_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_kge_plot <- yr_kge_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'kge') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual kge vals for ensembles -->
<!-- plot_kge_yr_ensem <- function(calyear) { -->
<!--   yr_kge_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = kge)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_kge_plot$year) -->

<!-- lapply(name, plot_kge_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly KGEs for ensembles -->

<!-- ```{r monthly kge ensemble, echo = F} -->

<!-- mo_kge <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = KGE(grid_ensemble, obs), -->
<!--             statn_ensemble = KGE(statn_ensemble, obs), -->
<!--             all_ensemble = KGE(all_ensemble, obs), -->
<!--             grid_all = KGE(grid_all, obs), -->
<!--             statn_all = KGE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- kable(mo_kge) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot <- mo_kge %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'KGE') %>% -->
<!--   drop_na() -->

<!-- kge_plot_mo <- ggplot(mo_pivot, aes(month, KGE, color = cal_run)) + geom_point(size = 3) + -->
<!--     scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon", -->
<!--                                             "statn_ensemble"="darkturquoise", -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4', -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   theme_bw() -->

<!-- ggplotly(kge_plot_mo) -->

<!-- ``` -->

<!-- ## RMSE -->

<!-- ### Ensemble RMSE for entire period -->

<!-- ```{r ensemble rmse, echo = F} -->

<!-- ensem_rmse_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = rmse(grid_ensemble, obs), -->
<!--             statn_ensemble = rmse(statn_ensemble, obs), -->
<!--             all_ensemble = rmse(all_ensemble, obs), -->
<!--             grid_all = rmse(grid_all, obs), -->
<!--             statn_all = rmse(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(ensem_rmse_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual RMSE for ensembles -->

<!-- ```{r annual rmse ensem, echo = F} -->

<!-- yr_rmse_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = rmse(grid_ensemble, obs), -->
<!--             statn_ensemble = rmse(statn_ensemble, obs), -->
<!--             all_ensemble = rmse(all_ensemble, obs), -->
<!--             grid_all = rmse(grid_all, obs), -->
<!--             statn_all = rmse(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- yr_rmse_table <- yr_rmse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'rmse') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'rmse') -->

<!-- yr_rmse_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_rmse_plot <- yr_rmse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'rmse') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual rmse vals for ensembles -->
<!-- plot_rmse_yr_ensem <- function(calyear) { -->
<!--   yr_rmse_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = rmse)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_rmse_plot$year) -->

<!-- lapply(name, plot_rmse_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly RMSEs for ensembles -->

<!-- ```{r month ensem rmse, echo = F} -->

<!-- mo_rmse <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = rmse(grid_ensemble, obs), -->
<!--             statn_ensemble = rmse(statn_ensemble, obs), -->
<!--             all_ensemble = rmse(all_ensemble, obs), -->
<!--             grid_all = rmse(grid_all, obs), -->
<!--             statn_all = rmse(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(mo_rmse) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot2 <- mo_rmse %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'RMSE') %>% -->
<!--   drop_na() -->

<!-- rmse_plot_mo <- ggplot(mo_pivot2, aes(month, RMSE, color = cal_run)) + geom_point(size = 3) + -->
<!--     scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon", -->
<!--                                             "statn_ensemble"="darkturquoise", -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4', -->
<!--                                             "all_ensemble" = "orange")) + -->
<!--   theme_bw() -->

<!-- ggplotly(rmse_plot_mo) -->

<!-- ``` -->

<!-- ## NSE -->

<!-- ### Ensemble NSE for entire period -->

<!-- ```{r nse all, echo = F} -->

<!-- ensem_nse_all <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   summarize(grid_ensemble = NSE(grid_ensemble, obs), -->
<!--             statn_ensemble = NSE(statn_ensemble, obs), -->
<!--             all_ensemble = NSE(all_ensemble, obs), -->
<!--             grid_all = NSE(grid_all, obs), -->
<!--             statn_all = NSE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 2) -->

<!-- kable(ensem_nse_all) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- ``` -->

<!-- ### Annual NSE for ensembles -->

<!-- ```{r annual nse, echo = F} -->

<!-- yr_nse_ensem <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   mutate(year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(grid_ensemble = NSE(grid_ensemble, obs), -->
<!--             statn_ensemble = NSE(statn_ensemble, obs), -->
<!--             all_ensemble = NSE(all_ensemble, obs), -->
<!--             grid_all = NSE(grid_all, obs), -->
<!--             statn_all = NSE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- yr_nse_table <- yr_nse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>% -->
<!--   pivot_wider(names_from = 'year', values_from = 'nse') -->

<!-- yr_nse_table %>% -->
<!--   kable(.) %>% -->
<!--   kable_styling() -->

<!-- yr_nse_plot <- yr_nse_ensem %>% -->
<!--   pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>% -->
<!--   mutate(year = as.factor(year)) -->

<!-- #plot annual nse vals for ensembles -->
<!-- plot_nse_yr_ensem <- function(calyear) { -->
<!--   yr_nse_plot %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = model_run, -->
<!--                y = nse)) + -->
<!--     geom_col() + -->
<!--     ggtitle(calyear) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1)) -->
<!-- } -->

<!-- name <- levels(yr_nse_plot$year) -->

<!-- lapply(name, plot_nse_yr_ensem) -->

<!-- ``` -->

<!-- ### Monthly NSE for ensembles -->

<!-- ```{r monthly rmse, echo =F} -->

<!-- mo_nse <- ensem %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>% -->
<!--   mutate(month = month(date)) %>% -->
<!--   group_by(month) %>% -->
<!--   summarize(grid_ensemble = NSE(grid_ensemble, obs), -->
<!--             statn_ensemble = NSE(statn_ensemble, obs), -->
<!--             all_ensemble = NSE(all_ensemble, obs), -->
<!--             grid_all = NSE(grid_all, obs), -->
<!--             statn_all = NSE(statn_all, obs)) %>% -->
<!--     mutate_if(is.numeric, -->
<!--             round, -->
<!--             digits = 4) -->

<!-- kable(mo_nse) %>% -->
<!--   kable_styling(full_width = F) -->

<!-- mo_pivot2 <- mo_nse %>% -->
<!--   pivot_longer(!month, names_to = 'cal_run', values_to = 'nse') %>% -->
<!--   drop_na() -->

<!-- nse_plot_mo <- ggplot(mo_pivot2, aes(month, nse, color = cal_run)) + geom_point(size = 3) + -->
<!--   theme_bw() + -->
<!--   scale_color_manual(name="cal_run", values=c("grid_ensemble"="maroon", -->
<!--                                             "statn_ensemble"="darkturquoise", -->
<!--                                             "grid_all"="purple1", -->
<!--                                             "statn_all"='springgreen4', -->
<!--                                             "all_ensemble" = "orange")) -->

<!-- ggplotly(nse_plot_mo) -->

<!-- ``` -->

<!-- # Catchment results -->

<!-- ## Precip -->

<!-- ```{r catch precip, echo = F} -->
<!-- setwd("/Users/megansears/Documents/Repos/fourmile/catchment_output") -->

<!-- statn <- read_csv('statnall.csv') -->
<!-- grid <- read_csv('gridall.csv') -->

<!-- grid <- grid %>% mutate(site = 'grid') -->
<!-- statn <- statn %>% mutate(site = 'statn') -->

<!-- catchment <- bind_rows(grid, statn) %>% -->
<!--   mutate(year = as.factor(year(time))) %>% -->
<!--   group_by(year, site) %>% -->
<!--   mutate(cumulative_p = cumsum(precip)) -->

<!-- #plot cumulative precip by year -->
<!-- plot_cumprecip <- function(calyear) { -->
<!--   {catchment %>% -->
<!--     filter(year == calyear) %>% -->
<!--     ggplot(aes(x = time, -->
<!--                y = cumulative_p, color = site)) + -->
<!--     geom_line(size=1) + -->
<!--     ggtitle(calyear) + -->
<!--     ylab('Cumulative precip (mm)') + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = 45, hjust=1))} %>% -->
<!--     ggplotly() -->
<!-- } -->

<!-- name <- levels(catchment$year) -->

<!-- precip_plots <- lapply(name, plot_cumprecip) -->

<!-- htmltools::tagList(precip_plots) -->

<!-- precip_plots -->

<!-- ## total P -->
<!-- max_cumulativep <- catchment %>% -->
<!--   group_by(year, site) %>% -->
<!--   summarize(total_cumulativep = max(cumulative_p)) -->

<!-- ggplot(max_cumulativep, aes(x=year, y=total_cumulativep, color=site)) + -->
<!--   geom_point(size=3) + -->
<!--   ggtitle('Total cumulative precip (mm) by year') -->

<!-- ``` -->

<!-- ## Tmean -->

<!-- ```{r tmean catch, echo = F} -->
<!-- ## tmean -->
<!-- tmean <- catchment %>% -->
<!--   group_by(year, site) %>% -->
<!--   summarize(tmean = mean(tmean)) %>% -->
<!--   ggplot(aes(x = year, y=tmean, color = site)) + -->
<!--   geom_point(size=3) + -->
<!--   ggtitle('Average temp by year') -->

<!-- tmean -->

<!-- ``` -->

# RD and RG

```{r rd and rg, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 14, fig.asp = 1, out.width = "100%"}

gw_gridall <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_all.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_all')

gw_statnall <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_all.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_all')

gw_grid2011 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2011.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2011')

gw_grid2012 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2012.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2012')

gw_grid2013 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2013.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2013')

gw_grid2014 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2014.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2014')

gw_grid2015 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2015.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2015')

gw_grid2016 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2016.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2016')

gw_grid2017 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2017.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2017')

gw_grid2018 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2018.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2018')

gw_grid2019 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2019.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2019')

gw_grid2020 <- read_csv(here('38HRU_sept2013_updates/catchment_output/grid_2020.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'grid_2020')

gw_statn2011 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2011.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2011')

gw_statn2012 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2012.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2012')

gw_statn2013 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2013.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2013')

gw_statn2014 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2014.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2014')

gw_statn2015 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2015.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2015')

gw_statn2016 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2016.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2016')

gw_statn2017 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2017.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2017')

gw_statn2018 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2018.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2018')

gw_statn2019 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2019.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2019')

gw_statn2020 <- read_csv(here('38HRU_sept2013_updates/catchment_output/statn_2020.csv')) %>%
  dplyr::select(time, catchmentRD1_w, catchmentRD2_w, catchmentRG1_w, catchmentRG2_w) %>%
  mutate(source = 'statn_2020')

gw <- bind_rows(
  gw_gridall, gw_statnall,
  gw_grid2011, gw_grid2012,
  gw_grid2013, gw_grid2014,
  gw_grid2015, gw_grid2016,
  gw_grid2017, gw_grid2018,
  gw_grid2019, gw_grid2020,
  gw_statn2011, gw_statn2012,
  gw_statn2013, gw_statn2014,
  gw_statn2015, gw_statn2016,
  gw_statn2017, gw_statn2018,
  gw_statn2019, gw_statn2020
) %>%
  mutate(date = mdy(time),
         year = year(date))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRD1_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RD1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRD2_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RD2') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRG1_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RG1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=catchmentRG2_w), size =1) +
  facet_wrap(~year, scales = 'free', ncol=2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('RG2') +
  theme(text = element_text(size = 20))

gw <- gw %>%
  arrange(date) %>%
  group_by(year, source) %>%
  mutate(RD1cumsum = cumsum(catchmentRD1_w),
         RD2cumsum = cumsum(catchmentRD2_w),
         RG1cumsum = cumsum(catchmentRG1_w),
         RG2cumsum = cumsum(catchmentRG2_w))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RD1cumsum), size =1) +
  facet_wrap(~year, scales = 'free', ncol=2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RD1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RD2cumsum), size = 1) +
  facet_wrap(~year, scales = 'free', ncol=2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RD2') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RG1cumsum), size  = 1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RG1') +
  theme(text = element_text(size = 20))

ggplot(gw, aes(x=date, color = source)) +
  geom_line(aes(y=RG2cumsum), size =1) +
  facet_wrap(~year, scales = 'free', ncol = 2) +
  scale_x_date(labels = date_format("%b")) +
  ggtitle('Cumulative RG2') +
  theme(text = element_text(size = 20))

```

```{r rd and rg sums, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 22, fig.asp = 1.5, out.width = "100%"}

##################
# all 4 with facets

ggplot(gw, aes(x=date)) +
  geom_line(aes(y=RG2cumsum, color = 'cumulative RG2'), size =1) +
  geom_line(aes(y=RG1cumsum, color = 'cumulative RG1'), size =1) +
  geom_line(aes(y=RD1cumsum, color = 'cumulative RD1'), size =1) +
  geom_line(aes(y=RD2cumsum, color = 'cumulative RD2'), size =1) +
  scale_x_date(labels = date_format("%b")) +
  facet_wrap(~year + source, scales = 'free_x', ncol=2) +
  ggtitle('Cumulative RD and RG') +
  theme(text = element_text(size = 24))

```

```{r gw boxplot, echo = FALSE, message = FALSE, warning = FALSE}

gw_maxcumsum <- gw %>%
  group_by(source, year) %>%
  summarize(yr_cum_RD1 = max(RD1cumsum),
            yr_cum_RD2 = max(RD2cumsum),
            yr_cum_RG1 = max(RG1cumsum),
            yr_cum_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  ggtitle('Max annual cumulative sums')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~flow_type, scales = 'free') +
  ggtitle('Max annual cumulative sums')

```

### RD and RG by season

```{r flow season, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 12, fig.asp = 0.75, out.width = "100%"}

#group by season and get max sums by season
flow_season <- gw %>%
  dplyr::select(RD1cumsum, RD2cumsum, RG1cumsum, RG2cumsum, date, year, source) %>%
  mutate(month = month(date)) %>%
  mutate(season = case_when(month %in% c(3,4,5,6) ~ 'snowmelt',
                            month %in% c(7,8,9,10) ~ 'monsoon',
                            month %in% c(11,12,1,2) ~ 'baseflow')) %>%
  group_by(source, year, season) %>%
  summarize(yr_season_RD1 = max(RD1cumsum),
            yr_season_RD2 = max(RD2cumsum),
            yr_season_RG1 = max(RG1cumsum),
            yr_season_RG2 = max(RG2cumsum)) %>%
            pivot_longer(!c(year, source, season), names_to = 'flow_type', values_to = 'flow_cms')

flow_season %>%
  ggplot(aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~season, scales = 'free') +
  ggtitle('Max annual season cumulative sums') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(text = element_text(size = 20))

flow_season %>%
  ggplot(aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~season + flow_type, scales = 'free') +
  ggtitle('Max annual season cumulative sums') +
  theme_bw() +
  theme(text = element_text(size = 20))

```

# HRU results

```{r cumulative tmean for HRUs, echo = F}
hrus <- st_read('/Users/megansears/Documents/Repos/fourmile/38HRU_model/hrus_final.shp')

mapview(hrus)

locations <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/fourmile_sensors.csv')

locations <- locations %>%
  st_as_sf(., coords=c('y','x'), crs = 4326)

#grid first
setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/HRU_output/grid")

read_ages_grid <- function(dataframe){
  read_csv(dataframe, skip = 3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'grid')

}

#pull in all filenames with HRU in the name
filenames <- list.files(".", pattern="HRU", full.names=F)

#read all the csvs using read_ages function
dataframes <- lapply(filenames, read_ages_grid)

#name all the csvs based on first 6 characters
names(dataframes) <- substr(filenames, 1, 6)

#bind all and make 1 df
hru_grids <- dataframes %>%
  bind_rows()

#statn next
setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/HRU_output/statn")

read_ages_statn <- function(dataframe){
  read_csv(dataframe, skip = 3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'statn')

}

#pull in all filenames with HRU in the name
filenames <- list.files(".", pattern="HRU", full.names=F)

#read all the csvs using read_ages function
dataframes <- lapply(filenames, read_ages_statn)

#name all the csvs based on first 6 characters
names(dataframes) <- substr(filenames, 1, 6)

#bind all and make 1 df
hru_statn <- dataframes %>%
  bind_rows()

# #get rid of a bunch of data
# hru_grid_tmean <- hru_grids %>%
#   dplyr::select(c(time, ID, tmean)) %>%
#   rename(tmean_grid = tmean)  %>%
#   group_by(ID) %>%
#   summarize(tmean_all_grid = mean(tmean_grid))
# 
# hru_statn_tmean <- hru_statn %>%
#   dplyr::select(c(time, ID, tmean)) %>%
#   rename(tmean_statn = tmean) %>%
#   group_by(ID) %>%
#   summarize(tmean_all_statn = mean(tmean_statn))
# 
# tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('ID')) %>%
#   mutate(diff = tmean_all_statn - tmean_all_grid) %>%
#   rename(value = ID)
# 
# hrus_map_tmean <- left_join(hrus, tmean_hrus, by = 'value')
# 
# tmap_mode('view')
# 
# locations_t <- locations %>%
#   dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
#                         'Betasso', 'Niwot_Snotel',
#                         'University Camp SNTL'))
# 
# tm_shape(hrus_map_tmean) +
#   tm_polygons('diff', id = 'value',
#               breaks = c(-6,-4,-2,0,2,4,6)) +
# tm_shape(locations_t) +
#   tm_symbols(col='blue', size=4)

```

## Tmean difference
Station - grid

```{r}

#get rid of a bunch of data
hru_grid_tmean <- hru_grids %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_grid = tmean)  %>%
  group_by(ID) %>%
  summarize(tmean_all_grid = mean(tmean_grid))

hru_statn_tmean <- hru_statn %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_statn = tmean) %>%
  group_by(ID) %>%
  summarize(tmean_all_statn = mean(tmean_statn))

tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('ID')) %>%
  mutate(diff = tmean_all_statn - tmean_all_grid) %>%
  rename(value = ID)

hrus_map_tmean <- left_join(hrus, tmean_hrus, by = 'value')

tmean <- ggplot() +
  geom_sf(data = hrus_map_tmean, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Tmean difference (°C)",
                       limits = c(-5, 2)) +
  theme_bw(base_size=13.5) +
  theme_void() +
  theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())

tmean

```

## Tmax difference

```{r tmax, echo = F}

# #get rid of a bunch of data
# hru_grid_tmax <- hru_grids %>%
#   dplyr::select(c(time, ID, tmax)) %>%
#   rename(tmax_grid = tmax)  %>%
#   group_by(ID) %>%
#   summarize(tmax_all_grid = mean(tmax_grid))
# 
# hru_statn_tmax <- hru_statn %>%
#   dplyr::select(c(time, ID, tmax)) %>%
#   rename(tmax_statn = tmax) %>%
#   group_by(ID) %>%
#   summarize(tmax_all_statn = mean(tmax_statn))
# 
# tmax_hrus <- left_join(hru_grid_tmax, hru_statn_tmax, by = c('ID')) %>%
#   mutate(diff = tmax_all_statn - tmax_all_grid) %>%
#   rename(value = ID)
# 
# hrus_map_tmax <- left_join(hrus, tmax_hrus, by = 'value')
# 
# tmap_mode('view')
# 
# locations_t <- locations %>%
#   dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
#                         'Betasso', 'Niwot_Snotel',
#                         'University Camp SNTL'))
# 
# tm_shape(hrus_map_tmax) +
#   tm_polygons('diff', id = 'value',
#               breaks = c(-6,-4,-2,0,2,4,6)) +
# tm_shape(locations_t) +
#   tm_symbols(col='blue', size=4)

# T max
#get rid of a bunch of data
hru_grid_tmax <- hru_grids %>%
  dplyr::select(c(time, ID, tmax)) %>%
  rename(tmax_grid = tmax)  %>%
  group_by(ID) %>%
  summarize(tmax_all_grid = mean(tmax_grid))

hru_statn_tmax <- hru_statn %>%
  dplyr::select(c(time, ID, tmax)) %>%
  rename(tmax_statn = tmax) %>%
  group_by(ID) %>%
  summarize(tmax_all_statn = mean(tmax_statn))

tmax_hrus <- left_join(hru_grid_tmax, hru_statn_tmax, by = c('ID')) %>%
  mutate(diff = tmax_all_statn - tmax_all_grid) %>%
  rename(value = ID)

hrus_map_tmax <- left_join(hrus, tmax_hrus, by = 'value')

tmax <- ggplot() +
  geom_sf(data = hrus_map_tmax, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Tmax difference (°C)",
                       limits = c(-5, 2)) +
  theme_bw(base_size=13.5) +
  theme_void() +
  theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())

tmax

```

## Tmin difference

```{r tmin, echo = F}

# #get rid of a bunch of data
# hru_grid_tmin <- hru_grids %>%
#   dplyr::select(c(time, ID, tmin)) %>%
#   rename(tmin_grid = tmin)  %>%
#   group_by(ID) %>%
#   summarize(tmin_all_grid = mean(tmin_grid))
# 
# hru_statn_tmin <- hru_statn %>%
#   dplyr::select(c(time, ID, tmin)) %>%
#   rename(tmin_statn = tmin) %>%
#   group_by(ID) %>%
#   summarize(tmin_all_statn = mean(tmin_statn))
# 
# tmin_hrus <- left_join(hru_grid_tmin, hru_statn_tmin, by = c('ID')) %>%
#   mutate(diff = tmin_all_statn - tmin_all_grid) %>%
#   rename(value = ID)
# 
# hrus_map_tmin <- left_join(hrus, tmin_hrus, by = 'value')
# 
# tmap_mode('view')
# 
# locations_t <- locations %>%
#   dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
#                         'Betasso', 'Niwot_Snotel',
#                         'University Camp SNTL'))
# 
# tm_shape(hrus_map_tmin) +
#   tm_polygons('diff', id = 'value',
#               breaks = c(-6,-4,-2,0,2,4,6)) +
# tm_shape(locations_t) +
#   tm_symbols(col='blue', size=4)

# Tmin
#get rid of a bunch of data
hru_grid_tmin <- hru_grids %>%
  dplyr::select(c(time, ID, tmin)) %>%
  rename(tmin_grid = tmin)  %>%
  group_by(ID) %>%
  summarize(tmin_all_grid = mean(tmin_grid))

hru_statn_tmin <- hru_statn %>%
  dplyr::select(c(time, ID, tmin)) %>%
  rename(tmin_statn = tmin) %>%
  group_by(ID) %>%
  summarize(tmin_all_statn = mean(tmin_statn))

tmin_hrus <- left_join(hru_grid_tmin, hru_statn_tmin, by = c('ID')) %>%
  mutate(diff = tmin_all_statn - tmin_all_grid) %>%
  rename(value = ID)

hrus_map_tmin <- left_join(hrus, tmin_hrus, by = 'value')

tmin <- ggplot() +
  geom_sf(data = hrus_map_tmin, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Tmin difference (°C)",
                       limits = c(-5, 2)) +
  theme_bw(base_size=13.5) +
  theme_void() +
  theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())

tmin

```

## Cumulative Precip difference

2011-2020 average of daily station-grid precip for DOY 365 (or 366)

```{r cum precip, echo = F}
# #get rid of a bunch of data
# hru_grid_precip <- hru_grids %>%
#   dplyr::select(c(time, ID, precip)) %>%
#   rename(precip_grid = precip)
# 
# hru_statn_precip <- hru_statn %>%
#   dplyr::select(c(time, ID, precip)) %>%
#   rename(precip_statn = precip)
# 
# precip_hrus <- left_join(hru_grid_precip, hru_statn_precip, by = c('time', 'ID')) %>%
#   mutate(diff = precip_statn - precip_grid,
#          year = year(time)) %>%
#   group_by(year, ID) %>%
#   mutate(total_diff = cumsum(diff)) %>%
#   filter(time == max(time)) %>%
#   group_by(ID) %>%
#   summarize(precip_avg_cumdiff = mean(total_diff)) %>%
#   rename(value = ID)
# 
# hrus_map_precip <- left_join(hrus, precip_hrus, by = 'value')
# 
# tmap_mode('view')
# 
# locations_p <- locations %>%
#   filter(type == 'precip')
# 
# tm_shape(hrus_map_precip) +
#   tm_polygons('precip_avg_cumdiff', midpoint = NA, palette = 'YlGnBu', id = 'value') +
# tm_shape(locations_p) +
#   tm_symbols(col='blue', size=4)

# Cumulative p
hru_grid_precip <- hru_grids %>%
  dplyr::select(c(time, ID, precip)) %>%
  rename(precip_grid = precip)

hru_statn_precip <- hru_statn %>%
  dplyr::select(c(time, ID, precip)) %>%
  rename(precip_statn = precip)

precip_hrus <- left_join(hru_grid_precip, hru_statn_precip, by = c('time', 'ID')) %>%
  mutate(diff = precip_statn - precip_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>%
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(precip_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

hrus_map_precip <- left_join(hrus, precip_hrus, by = 'value')

precip <- ggplot() +
  geom_sf(data = hrus_map_precip, aes(fill = precip_avg_cumdiff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "mako", name = "Cumulative P Difference (mm)") +
  theme_bw(base_size=13.5) +
  theme_void() +
  theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())

precip

```

## Solar radiation difference

2011-2020 average of daily station-grid solar rad

```{r cum sol rad, echo = F}

# #get rid of a bunch of data
# hru_grid_solrad <- hru_grids %>%
#   dplyr::select(c(time, ID, solRad)) %>%
#   rename(solrad_grid = solRad)  %>%
#   group_by(ID) %>%
#   summarize(solrad_all_grid = mean(solrad_grid))
# 
# hru_statn_solrad <- hru_statn %>%
#   dplyr::select(c(time, ID, solRad)) %>%
#   rename(solrad_statn = solRad) %>%
#   group_by(ID) %>%
#   summarize(solrad_all_statn = mean(solrad_statn))
# 
# solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('ID')) %>%
#   mutate(diff = solrad_all_statn - solrad_all_grid) %>%
#   rename(value = ID)
# 
# hrus_map_solrad <- left_join(hrus, solrad_hrus, by = 'value')
# 
# tmap_mode('view')
# 
# locations_solrad <- locations %>%
#   dplyr::filter(name %in% c('Gordon_South',
#                         'Betasso', 'Niwot_Snotel'))
# 
# tm_shape(hrus_map_solrad) +
#   tm_polygons('diff', midpoint = NA, palette = 'YlOrRd', id = 'value') +
# tm_shape(locations_solrad) +
#   tm_symbols(col='blue', size=4)

# solar rad
hru_grid_solrad <- hru_grids %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_grid = solRad)  %>%
  group_by(ID) %>%
  summarize(solrad_all_grid = mean(solrad_grid))

hru_statn_solrad <- hru_statn %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_statn = solRad) %>%
  group_by(ID) %>%
  summarize(solrad_all_statn = mean(solrad_statn))

solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('ID')) %>%
  mutate(diff = solrad_all_statn - solrad_all_grid) %>%
  rename(value = ID)

hrus_map_solrad <- left_join(hrus, solrad_hrus, by = 'value')

sr <- ggplot() +
  geom_sf(data = hrus_map_solrad, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "cividis", name = expression("Solar rad difference (MJ/m"^"2" * ")")) +
  theme_bw(base_size=13.5) +
  theme_void() +
  theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()) +
  guides(
    fill = guide_colorbar(
      label.theme = element_text(size = 10)))

sr

```

## RH difference

```{r rh, echo = F}

# #get rid of a bunch of data
# hru_grid_rh <- hru_grids %>%
#   dplyr::select(c(time, ID, rhum)) %>%
#   rename(rh_grid = rhum)  %>%
#   group_by(ID) %>%
#   summarize(rh_all_grid = mean(rh_grid))
# 
# hru_statn_rh <- hru_statn %>%
#   dplyr::select(c(time, ID, rhum)) %>%
#   rename(rh_statn = rhum) %>%
#   group_by(ID) %>%
#   summarize(rh_all_statn = mean(rh_statn))
# 
# rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('ID')) %>%
#   mutate(diff = rh_all_statn - rh_all_grid) %>%
#   rename(value = ID)
# 
# hrus_map_rh <- left_join(hrus, rh_hrus, by = 'value')
# 
# tmap_mode('view')
# 
# locations_rh <- locations %>%
#   dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
#                         'Betasso', 'Niwot_Snotel'))
# 
# tm_shape(hrus_map_rh) +
#   tm_polygons('diff', midpoint = NA, id = 'value') +
# tm_shape(locations_rh) +
#   tm_symbols(col='blue', size=4)

# RH
#get rid of a bunch of data
hru_grid_rh <- hru_grids %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_grid = rhum)  %>%
  group_by(ID) %>%
  summarize(rh_all_grid = mean(rh_grid))

hru_statn_rh <- hru_statn %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_statn = rhum) %>%
  group_by(ID) %>%
  summarize(rh_all_statn = mean(rh_statn))

rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('ID')) %>%
  mutate(diff = rh_all_statn - rh_all_grid) %>%
  rename(value = ID)

hrus_map_rh <- left_join(hrus, rh_hrus, by = 'value')

rh <- ggplot() +
  geom_sf(data = hrus_map_rh, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "viridis", name = "RH difference (%)") +
  theme_bw(base_size=13.5) +
  theme_void() +
  theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())

rh

```

```{r}

# lots <- ggarrange(precip,
#                   sr,
#                   rh,
#                   tmean,
#                   tmax,
#                   tmin) 
#           #nrow = 4)
#           #ncol= 0)
# lots

```

## Average soil sat comparison

```{r soilsat, echo = F}

#get rid of a bunch of data
hru_grid_soilsat <- hru_grids %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_grid = soilSat)  %>%
  group_by(ID) %>%
  summarize(soilsat_all_grid = mean(soilsat_grid))

hru_statn_soilsat <- hru_statn %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_statn = soilSat) %>%
  group_by(ID) %>%
  summarize(soilsat_all_statn = mean(soilsat_statn))

soilsat_hrus <- left_join(hru_grid_soilsat, hru_statn_soilsat, by = c('ID')) %>%
  rename(value = ID)

hrus_map_soilsat <- left_join(hrus, soilsat_hrus, by = 'value')

tmap_mode('plot')

soilavg_grid <- tm_shape(hrus_map_soilsat) +
  tm_polygons('soilsat_all_grid', midpoint = NA, id = 'value',
              breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average soil sat')

soilavg_statn <- tm_shape(hrus_map_soilsat) +
  tm_polygons('soilsat_all_statn', midpoint = NA, id = 'value',
              breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4)) +
  tm_legend(legend.outside = T)

tmap_arrange(soilavg_grid, soilavg_statn)

```

## Average max soil sat comparison

```{r max soilsat, echo = F}

#get rid of a bunch of data
hru_grid_soilsat_max <- hru_grids %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_grid = soilSat)  %>%
  mutate(year = year(time)) %>%
  group_by(ID, year) %>%
  summarize(soilsat_yrmax_grid = max(soilsat_grid)) %>%
  summarize(avg_max_grid = mean(soilsat_yrmax_grid))

hru_statn_soilsat_max <- hru_statn %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_statn = soilSat) %>%
  mutate(year = year(time)) %>%
  group_by(ID, year) %>%
  summarize(soilsat_yrmax_statn = max(soilsat_statn)) %>%
  summarize(avg_max_statn = mean(soilsat_yrmax_statn))

soilsat_hrus_max <- left_join(hru_grid_soilsat_max, hru_statn_soilsat_max, by = c('ID')) %>%
  rename(value = ID)

hrus_map_soilsat_max <- left_join(hrus, soilsat_hrus_max, by = 'value')

tmap_mode('plot')

soilavg_grid_max <- tm_shape(hrus_map_soilsat_max) +
  tm_polygons('avg_max_grid', midpoint = NA, id = 'value',
              breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average max soil sat',
            title.position = c('left', 'top'))

soilavg_statn_max <- tm_shape(hrus_map_soilsat_max) +
  tm_polygons('avg_max_statn', midpoint = NA, id = 'value',
              breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8)) +
  tm_legend(legend.outside = T)

tmap_arrange(soilavg_grid_max, soilavg_statn_max)

```

## Average ET comparison

```{r et, echo = F}
#get rid of a bunch of data
hru_grid_ET <- hru_grids %>%
  dplyr::select(c(time, ID, actET)) %>%
  rename(ET_grid = actET)  %>%
  group_by(ID) %>%
  summarize(ET_all_grid = mean(ET_grid))

hru_statn_ET <- hru_statn %>%
  dplyr::select(c(time, ID, actET)) %>%
  rename(ET_statn = actET) %>%
  group_by(ID) %>%
  summarize(ET_all_statn = mean(ET_statn))

ET_hrus <- left_join(hru_grid_ET, hru_statn_ET, by = c('ID')) %>%
  rename(value = ID)

hrus_map_ET <- left_join(hrus, ET_hrus, by = 'value')

tmap_mode('plot')

ETavg_grid <- tm_shape(hrus_map_ET) +
  tm_polygons('ET_all_grid', midpoint = NA, id = 'value',
              breaks = c(0.0,0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average ET')

ETavg_statn <- tm_shape(hrus_map_ET) +
  tm_polygons('ET_all_statn', midpoint = NA, id = 'value',
             breaks = c(0.0, 0.1, 0.2, 0.4, 0.6, 0.8, 1.0, 1.2)) +
  tm_legend(legend.outside = T)

tmap_arrange(ETavg_grid, ETavg_statn)

```

## Average ET difference

```{r et diff, echo = F}

ET_hrus <- ET_hrus %>%
  mutate(diff = ET_all_statn - ET_all_grid)

hrus_map_et_diff <- left_join(hrus, ET_hrus, by = 'value')

tmap_mode('plot')

tm_shape(hrus_map_et_diff) +
  tm_polygons('diff', midpoint = NA, id = 'value')  +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average ET difference')

```

## Average LAI comparison

```{r lai, echo = F}

#get rid of a bunch of data
hru_grid_LAI <- hru_grids %>%
  dplyr::select(c(time, ID, LAI)) %>%
  rename(LAI_grid = LAI)  %>%
  group_by(ID) %>%
  summarize(LAI_all_grid = mean(LAI_grid))

hru_statn_LAI <- hru_statn %>%
  dplyr::select(c(time, ID, LAI)) %>%
  rename(LAI_statn = LAI) %>%
  group_by(ID) %>%
  summarize(LAI_all_statn = mean(LAI_statn))

LAI_hrus <- left_join(hru_grid_LAI, hru_statn_LAI, by = c('ID')) %>%
  rename(value = ID)

hrus_map_LAI <- left_join(hrus, LAI_hrus, by = 'value')

tmap_mode('plot')

LAIavg_grid <- tm_shape(hrus_map_LAI) +
  tm_polygons('LAI_all_grid', midpoint = NA, id = 'value',
              breaks = c(1,2,3,4,5)) +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average LAI')

LAIavg_statn <- tm_shape(hrus_map_LAI) +
  tm_polygons('LAI_all_statn', midpoint = NA, id = 'value',
              breaks = c(1,2,3,4,5)) +
  tm_legend(legend.outside = T)

tmap_arrange(LAIavg_grid, LAIavg_statn)

```

# Join param & RD/RG

```{r}

join <- gw_maxcumsum %>%
  rename(input_data = source)

params1 <- params %>%
  dplyr::select(-input_data) %>%
  rename(input_data = cal) %>%
  dplyr::select(-c(year_missing))

#
params_weighted1 <- params_weighted %>%
  mutate(input_data = paste(source, year_missing, sep = "_")) %>% 
  pivot_wider(names_from = 'param',
              values_from = 'area_weighted_average') %>%
  rename_with(~ paste0(., '_weighted'), 4:9) %>%
  dplyr::select(-c(year_missing, source))

join <- left_join(join, params1, by = c('input_data')) %>%
  pivot_wider(., names_from='param', values_from='value') %>%
  pivot_wider(., names_from = 'flow_type', values_from = 'flow_cms')

join <- left_join(join, params_weighted1, by = 'input_data')

write_csv(join, '/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/flow_params.csv')

```