### 2024-11-26 state_vars.Rmd
looking at state variables with climate inputs

---
title: "Fourmile state variables"
author: "Megan Sears"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F, 
  warning  = F)

library(tidyverse)
library(ggplot2); theme_set(theme_bw(base_size = 20))
library(plotly)
library(knitr)
library(kableExtra)

```

```{r, include = F}

statn <- read_csv('./38HRU_sept2013_updates/catchment_output/Catchment_statnall.csv') %>%
  mutate(date = mdy(time),
         source = 'statn') %>%
  select(-time)

grid <- read_csv('./38HRU_sept2013_updates/catchment_output/Catchment_gridall.csv') %>%
  mutate(date = mdy(time),
         source = 'grid') %>%
  select(-time)

both <- bind_rows(statn, grid)

```

# Catchment

## Inputs

```{r}

temp <- ggplot(both, 
               aes(date, 
                   tmean,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(temp)

tmean_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_tmean = mean(tmean))

ggplot(tmean_avg, aes(x=month, y=mean_tmean,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean tmean (C)')

precip <- ggplot(both, 
               aes(date, 
                   precip,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(precip)

precip_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month, year) %>%
  summarize(sum_mo = sum(precip)) %>%
  ungroup() %>% 
  group_by(source, month) %>%
  summarize(mean_p = mean(sum_mo))

ggplot(precip_avg, aes(x=month, y=mean_p,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean precip (mm)')

solrad_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_solrad = mean(solRad))

ggplot(solrad_avg, aes(x=month, y=mean_solrad,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean solrad')

```

## Interception
- Average interception: grid > statn
- Monthly average: Statn > grid 1:4 and 10:12, grid > statn 5:9
- During the summer rain interception is greater from grid
- In the winter snow interception is greater from statn

```{r}
int_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_interception = mean(interception))

ggplot(int_avg, aes(x=month, y=mean_interception,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean interception')

interception <- ggplot(both, 
               aes(date, 
                   interception,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(interception)

```

## ET

- Average ET: grid > statn
- Monthly average ET: grid > statn 1:6 and 9:12, statn > grid 7:8

```{r}

et_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_et = mean(actET...21))

ggplot(et_avg, aes(x=month, y=mean_et,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean ET (mm)')

et <- ggplot(both, 
               aes(date, 
                   actET...21,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='ET (mm)')

ggplotly(et)

```

## SWE

- Grid > Statn
- SWE starts accumulating faster in Sept for statn than grid
- But grid > statn all other months

```{r}

swe_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_swe = mean(snowTotSWE))

ggplot(swe_avg, aes(x=month, y=mean_swe,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean SWE (l/day)')

snowTotSWE <- ggplot(both, 
                     aes(date, 
                         snowTotSWE,
                         color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='SWE (l/day)')

ggplotly(snowTotSWE)

```

## Soil sat (whole profile)

- Statn > grid all months

```{r}

soil_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_soil = mean(soilSat))

ggplot(soil_avg, aes(x=month, y=mean_soil,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean soil sat')

soil <- ggplot(both, 
               aes(date, 
                   soilSat,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(soil)

```

## Surface runoff

- RG1: when there is surface runoff, statn is slightly higher
- Except for September (Sep 2013 flood heavily influences this avg)
- Many of the surface runoff events are happening Mar-May and Jul

```{r}

rd1_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rd1 = mean(outRD1))

ggplot(rd1_avg, aes(x=month, y=mean_rd1,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean surface runoff (l)')

rd1 <- ggplot(both, 
               aes(date, 
                   outRD1,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='Surface runoff (l)')

ggplotly(rd1)

```

## Interflow

- Statn > grid Jan to Apr
- Grid > statn May, Jun, Oct
- Other months are similar

```{r}

rd2_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rd2 = mean(outRD2))

ggplot(rd2_avg, aes(x=month, y=mean_rd2,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean interflow (l)')

rd2 <- ggplot(both, 
               aes(date, 
                   outRD2,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='Interflow (l)')

ggplotly(rd2)

```

## Groundwater

- RG1 = shallow groundwater
- RG1: during snowmelt runoff to fall baseflow grid > statn
- RG1: statn > grid during winter

- RG2 = groundwater baseflow
- RG2: Very small relative to RG1
- RG2: Grid > statn

```{r}
gw1_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rg1 = mean(outRG1))

ggplot(gw1_avg, aes(x=month, y=mean_rg1,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean RG1 (unit?)')

gw1 <- ggplot(both, 
               aes(date, 
                   outRG1,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(gw1)

################################################################
gw2_avg <- both %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rg2 = mean(outRG2))

ggplot(gw2_avg, aes(x=month, y=mean_rg2,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean RG2 (unit?)')

gw2 <- ggplot(both, 
               aes(date, 
                   outRG2,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(gw2)

```

# High elevation HRU

HRU18

```{r}

statn_hi <- read_csv('./38HRU_sept2013_updates/HRU_18_statn.csv') %>% 
  mutate(date = mdy(time),
         source = 'statn') %>%
  select(-time)

grid_hi <- read_csv('./38HRU_sept2013_updates/HRU_18_grid.csv') %>%
  mutate(date = mdy(time),
         source = 'grid') %>%
  select(-time)

both_hi <- bind_rows(statn_hi, grid_hi)

```

## Inputs

```{r}

tmean_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_tmean = mean(tmean))

ggplot(tmean_avg, aes(x=month, y=mean_tmean,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean tmean (C)')

temp <- ggplot(both_hi, 
               aes(date, 
                   tmean,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(temp)
############
precip_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month, year) %>%
  summarize(sum_mo = sum(precip)) %>%
  ungroup() %>% 
  group_by(source, month) %>%
  summarize(mean_p = mean(sum_mo))

ggplot(precip_avg, aes(x=month, y=mean_p,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean precip (mm)')

precip <- ggplot(both_hi, 
               aes(date, 
                   precip,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(precip)
#################

solrad_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_solrad = mean(solRad))

ggplot(solrad_avg, aes(x=month, y=mean_solrad,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean solrad (MJ/m2/day)')

```

## Interception

- Station > grid except for Jul

```{r}
int_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_interception = mean(interception))

ggplot(int_avg, aes(x=month, y=mean_interception,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean interception')

interception <- ggplot(both_hi, 
               aes(date, 
                   interception,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(interception)

```

## ET

- Grid > statn except for Mar

```{r}

et_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_et = mean(actET))

ggplot(et_avg, aes(x=month, y=mean_et,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean ET (mm)')

et <- ggplot(both_hi, 
               aes(date, 
                   actET,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='ET (mm)')

ggplotly(et)

```

## SWE

- Grid > statn

```{r}

swe_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_swe = mean(snowTotSWE))

ggplot(swe_avg, aes(x=month, y=mean_swe,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean SWE (l/day)')

snowTotSWE <- ggplot(both_hi, 
                     aes(date, 
                         snowTotSWE,
                         color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='SWE (l/day)')

ggplotly(snowTotSWE)

```

## Soil sat (whole profile)

- Statn > grid

```{r}

soil_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_soil = mean(soilSat))

ggplot(soil_avg, aes(x=month, y=mean_soil,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean soil sat')

soil <- ggplot(both_hi, 
               aes(date, 
                   soilSat,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(soil)

```

## Surface runoff

- No surface runoff

```{r}

rd1_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rd1 = mean(outRD1))

ggplot(rd1_avg, aes(x=month, y=mean_rd1,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean surface runoff (l)')

rd1 <- ggplot(both_hi, 
               aes(date, 
                   outRD1,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='Surface runoff (l)')

ggplotly(rd1)

```

## Interflow

- Statn > grid Jan to Apr
- Grid > statn May to Jun
- Other months are similar

```{r}

rd2_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rd2 = mean(outRD2))

ggplot(rd2_avg, aes(x=month, y=mean_rd2,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean interflow (l)')

rd2 <- ggplot(both_hi, 
               aes(date, 
                   outRD2,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='Interflow (l)')

ggplotly(rd2)

```

## Groundwater

- RG1 = shallow groundwater
- RG1: Jan to May statn > grid
- RG1: grid > statn Jun, Jul, Oct
- RG1: similar rest of year

- RG2 = groundwater baseflow
- RG2: Very small relative to RG1
- RG2: Grid > statn

```{r}
gw1_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rg1 = mean(outRG1))

ggplot(gw1_avg, aes(x=month, y=mean_rg1,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean RG1 (unit?)')

gw1 <- ggplot(both_hi, 
               aes(date, 
                   outRG1,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(gw1)

################################################################
gw2_avg <- both_hi %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rg2 = mean(outRG2))

ggplot(gw2_avg, aes(x=month, y=mean_rg2,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean RG2 (unit?)')

gw2 <- ggplot(both_hi, 
               aes(date, 
                   outRG2,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(gw2)

```

# Low elevation HRU

HRU19

```{r}

statn_lo <- read_csv('./38HRU_sept2013_updates/HRU_19_statn.csv') %>% 
  mutate(date = mdy(time),
         source = 'statn') %>%
  select(-time)

grid_lo <- read_csv('./38HRU_sept2013_updates/HRU_19_grid.csv') %>%
  mutate(date = mdy(time),
         source = 'grid') %>%
  select(-time)

both_lo <- bind_rows(statn_lo, grid_lo)

```

## Inputs

```{r}

tmean_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_tmean = mean(tmean))

ggplot(tmean_avg, aes(x=month, y=mean_tmean,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean tmean (C)')

temp <- ggplot(both_lo, 
               aes(date, 
                   tmean,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(temp)
############
precip_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month, year) %>%
  summarize(sum_mo = sum(precip)) %>%
  ungroup() %>% 
  group_by(source, month) %>%
  summarize(mean_p = mean(sum_mo))

ggplot(precip_avg, aes(x=month, y=mean_p,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean precip (mm)')

precip <- ggplot(both_lo, 
               aes(date, 
                   precip,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(precip)
#################

solrad_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_solrad = mean(solRad))

ggplot(solrad_avg, aes(x=month, y=mean_solrad,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean solrad (MJ/m2/day)')

```

## Interception

- Statn > grid Oct to Apr
- Grid > statn May to Sep

```{r}
int_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_interception = mean(interception))

ggplot(int_avg, aes(x=month, y=mean_interception,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean interception')

interception <- ggplot(both_lo, 
               aes(date, 
                   interception,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(interception)

```

## ET

- Grid > statn Sep to May

```{r}

et_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_et = mean(actET))

ggplot(et_avg, aes(x=month, y=mean_et,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean ET (mm)')

et <- ggplot(both_lo, 
               aes(date, 
                   actET,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='ET (mm)')

ggplotly(et)

```

## SWE

- Grid > statn Oct to Feb
- Statn > grid Mar to Apr and Sep

```{r}

swe_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_swe = mean(snowTotSWE))

ggplot(swe_avg, aes(x=month, y=mean_swe,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean SWE (l/day)')

snowTotSWE <- ggplot(both_lo, 
                     aes(date, 
                         snowTotSWE,
                         color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='SWE (l/day)')

ggplotly(snowTotSWE)

```

## Soil sat (whole profile)

- Statn > grid

```{r}

soil_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_soil = mean(soilSat))

ggplot(soil_avg, aes(x=month, y=mean_soil,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean soil sat')

soil <- ggplot(both_lo, 
               aes(date, 
                   soilSat,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(soil)

```

## Surface runoff

- Very little except for Sep 2013 (grid > statn)

```{r}

rd1_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rd1 = mean(outRD1))

ggplot(rd1_avg, aes(x=month, y=mean_rd1,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean surface runoff (l)')

rd1 <- ggplot(both_lo, 
               aes(date, 
                   outRD1,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='Surface runoff (l)')

ggplotly(rd1)

```

## Interflow

- Statn > grid Jan, May, June, July, Sep

```{r}

rd2_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rd2 = mean(outRD2))

ggplot(rd2_avg, aes(x=month, y=mean_rd2,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean interflow (l)')

rd2 <- ggplot(both_lo, 
               aes(date, 
                   outRD2,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y='Interflow (l)')

ggplotly(rd2)

```

## Groundwater

- RG1 = shallow groundwater
- RG1: Jan, Feb, May, Aug, Sep, Dec statn > grid

- RG2 = groundwater baseflow
- RG2: Very small relative to RG1
- RG2: Grid > statn

```{r}
gw1_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rg1 = mean(outRG1))

ggplot(gw1_avg, aes(x=month, y=mean_rg1,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean RG1 (unit?)')

gw1 <- ggplot(both_lo, 
               aes(date, 
                   outRG1,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(gw1)

################################################################
gw2_avg <- both_lo %>%
  mutate(month = month(date),
         year = year(date)) %>% 
  group_by(source, month) %>%
  summarize(mean_rg2 = mean(outRG2))

ggplot(gw2_avg, aes(x=month, y=mean_rg2,
                 color=source)) +
  geom_point(size=3) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(y= 'Monthly mean RG2 (unit?)')

gw2 <- ggplot(both_lo, 
               aes(date, 
                   outRG2,
                   color = source)) +
  geom_line() +
  scale_color_manual(values = c("orange", "blue"))

ggplotly(gw2)

```