### 2025-01-17 spatial_figs.Rmd
updated figure 2

---
title: "spatial_figures"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: html_document
---

```{r open package, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(dplyr)
library(lubridate)
library(plotly)
library(ggplot2)
library(here)
library(sf)
library(mapview)
library(raster)
library(sf)
library(terra)
library(viridis)
library(ggpubr)

```

# Figure 1a

```{r}

# update the HRUs
hru_new <- vect('/Users/megansears/Documents/Repos/fourmile/data/GIS/hrus_final.shp')
plot(hru_new)

hru_new <- st_read('/Users/megansears/Documents/Repos/fourmile/data/GIS/hrus_final.shp')

mapview(hru_new) 
#mapview(hru_test)

locations <- read_csv('/Users/megansears/Documents/Repos/fourmile/data/GIS/locations_all.csv')

locations <- locations %>%
  st_as_sf(., coords=c('lon','lat'), crs = 4326)

mapview(locations) + mapview(hru_new)

# Read the DEM raster file
dem_raster <- raster('/Users/megansears/Documents/Repos/fourmile/data/GIS/USGS_13_n41w106_20230314 (1).tif')
dem_raster2 <- raster('/Users/megansears/Documents/Repos/fourmile/data/GIS/USGS_13_n40w106_20230602.tif')

dem_raster <- merge(dem_raster, dem_raster2)

dem_raster1 <- rast(dem_raster)

# extract min, max
min <- terra::extract(dem_raster1, hru_new, 'min', touches=T)
min(min$layer)

max <- terra::extract(dem_raster1, hru_new, 'max', touches=T)
max(max$layer)

mapview(dem_raster) + mapview(hru_new) + mapview(locations)

#######
bbox <- st_bbox(locations)

# # Bounding box cuts off some points so expand it
xrange <- bbox$xmax - bbox$xmin # range of x values
yrange <- bbox$ymax - bbox$ymin # range of y values
# Expand the extent
bbox[1] <- bbox[1] - (0.1 * xrange) # xmin - left
bbox[3] <- bbox[3] + (0.1 * xrange) # xmax - right
bbox[2] <- bbox[2] - (0.1 * yrange) # ymin - bottom
bbox[4] <- bbox[4] + (0.1 * yrange) # ymax - top
# 
#######

hru_new <- st_transform(hru_new, crs(dem_raster))

dem_cropped <- crop(dem_raster, extent(bbox))

mapview(dem_cropped) + mapview(hru_new) + mapview(locations)

# Convert the raster to a data frame
dem_df <- as.data.frame(dem_cropped, xy = TRUE)

fire <- st_read('/Users/megansears/Documents/Repos/fourmile/data/GIS/co4005110538520100906_20100621_20110624_burn_bndy.shp')
mapview(fire) + mapview(hru_new)

fire <- st_transform(fire, crs(dem_raster))

## inset data

library(spData)
data("us_states", package = "spData")

# Extract the geometry of Colorado
wus <- us_states[us_states$REGION == "West", ]

point <- locations[locations$source == 'USGS1',]

# Create a ggplot with western US states and the bounding box
ggm1 <- ggplot() + 
  geom_sf(data = wus, fill = "white") + 
  geom_sf(data = point, shape = 8, size = 3, color = "red", fill = "red") +  # Add star point
  theme_void()

ggm1

##

library(ggspatial)

# Define custom shapes for each value of 'display_use'
custom_shapes <- c("Daymet" = 15, "Prism" = 16, "Precip" = 17, "Multi-sensor" = 18, "Streamflow" = 7)  # Replace "Type1", "Type2", etc. with actual values

terrain <- ggplot() +
  geom_raster(data = dem_df, 
              aes(x = x, y = y, fill = layer)) +  
  scale_fill_gradientn(colors = terrain.colors(200),  
    name = "Elevation (m)") +
   geom_sf(data = fire, 
          aes(color = "Fire"), 
          fill = NA, 
          lwd = 0.75) +
  geom_sf(data = hru_new, 
          aes(color = "HRU"), 
          fill = NA, 
          lwd = 0.5) +
  geom_sf(data = locations, 
          aes(shape = factor(display_use)), 
          size = 4) + 
  scale_shape_manual(values = custom_shapes) +
   scale_color_manual(name = NULL,
                      values = c(Fire = "red", HRU = "black"), 
                     labels = c(Fire = "Fourmile Canyon fire", HRU = "HRUs")) + 
  labs(shape = element_blank()) +
  theme_bw(base_size = 25) +
  theme(axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.spacing.y = unit(0.01, "cm"))

terrain
ggsave('./figures/fig1a.png', dpi=600)
##

reference_location <- data.frame(
  x = -105.2705,  # Longitude for Boulder
  y = 40.0150,    # Latitude for Boulder
  name = "Boulder, CO"
)

# Your existing terrain plot
terrain2 <- ggplot() +
  geom_raster(data = dem_df, 
              aes(x = x, y = y, fill = layer)) +  
  scale_fill_gradientn(colors = terrain.colors(200),  
                       name = "Elevation (m)") +
  geom_sf(data = fire, 
          aes(color = "Fire"), 
          fill = NA, 
          lwd = 0.75) +
  geom_sf(data = hru_new, 
          aes(color = "HRU"), 
          fill = NA, 
          lwd = 0.5) +
  geom_sf(data = locations, 
          aes(shape = factor(display_use)), 
          size = 4) + 
  scale_shape_manual(values = custom_shapes) +
  scale_color_manual(name = NULL,
                     values = c(Fire = "red", HRU = "black"), 
                     labels = c(Fire = "Fourmile Canyon fire", HRU = "HRUs")) + 
  labs(shape = element_blank()) +
  theme_bw(base_size = 25) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.spacing.y = unit(0.01, "cm")) +
  geom_point(data = reference_location, aes(x = x, y = y), color = "blue", size = 4, shape = 21, fill = "yellow") + # Reference marker
  geom_text(data = reference_location, aes(x = x, y = y, label = name), vjust = -1, color = "blue") # Reference label

terrain2
##

ggplot() +
  geom_raster(data = dem_df, aes(x = x, y = y, fill = layer)) +
  scale_fill_viridis_c(name = "Elevation (m)", 
                       option = "viridis",direction=-1) +
  geom_sf(data = fire, 
          aes(color = "Fire"), 
          fill = NA, 
          lwd = 0.75) +
  geom_sf(data = hru_new, 
          aes(color = "HRU"), 
          fill = NA, 
          lwd = 0.5) +
  geom_sf(data = locations, 
          aes(shape = factor(display_use)), 
          size = 4) + 
  scale_shape_manual(values = custom_shapes) +
   scale_color_manual(name = NULL,
                      values = c(Fire = "red", HRU = "black"), 
                     labels = c(Fire = "Fourmile Canyon fire", HRU = "HRUs")) + 
  labs(shape = element_blank()) +
  theme_bw(base_size = 25) +
  theme(axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.spacing.y = unit(0.01, "cm"))

ggsave('./figures/fig1a_2.png', dpi=600)

###
library(cowplot)

ggdraw() +
  draw_plot(terrain) +
  draw_plot(ggm1, x = 0.042, y = 0.63, width = 0.16, height = 0.16)
  #draw_plot(ggm1, x = 0.585, y = 0.56, width = 0.25, height = 0.25)

ggsave('./figures/fig1a.png', dpi=600)

```

# Figure 1b-e -- NEED TO REDO WITH WHOLE CATCHMENT DATA and UPDATED MODELS

```{r}
grid <- read_csv("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/grid_all_Catchment.csv",
                 skip=3) %>%
   slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(date = ymd(time)) %>%
  dplyr::select(-time) %>%
  mutate_if(is.character, as.numeric) %>% 
  mutate(site = 'grid')

statn <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/statn_all_Catchment.csv',
                 skip=3) %>%
   slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(date = ymd(time)) %>%
  dplyr::select(-time) %>%
  mutate_if(is.character, as.numeric) %>% 
  mutate(site = 'statn')

all <- bind_rows(grid, statn) %>%
  arrange(date)

catchment_mo_mean <- all %>%
  mutate(month = month(date)) %>%
  group_by(site, month) %>%
  summarise(across(precip:solRad, ~ mean(.x, na.rm = TRUE))) %>%
  dplyr::select(site, month,
                precip_mean = precip, tmean_mean = tmean, 
                solrad_mean = solRad, rh_mean = rhum)

catchment_mo_sd <- all %>%
  mutate(month = month(date)) %>%
  group_by(site, month) %>%
  summarise(across(precip:solRad, ~ sd(.x, na.rm = TRUE))) %>%
  dplyr::select(site, month,
                precip_sd = precip, tmean_sd = tmean, 
                solrad_sd = solRad, rh_sd = rhum)

catchment_mo <- left_join(catchment_mo_mean, catchment_mo_sd, 
                          by = c('month', 'site'))

catchment_mo2 <- catchment_mo %>%
  mutate(month_name = month.abb[month]) %>%
  mutate(month_name = factor(month_name, 
                             levels = c('Jan', 'Feb', 'Mar',
                                        'Apr', 'May', 'Jun',
                                        'Jul', 'Aug', 'Sep',
                                        'Oct', 'Nov', 'Dec')))

p_cols <- ggplot(catchment_mo2, aes(x = month_name, y = precip_mean,
                                    color = site, fill = site)) +
  geom_col(position = 'dodge') +
  geom_errorbar(aes(ymin = 0, ymax = precip_mean + precip_sd), 
                position = position_dodge(width = 0.6), width = 0.2) +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Rename color legend items
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Rename fill legend items
  theme_bw(base_size = 20) +
  labs(y = 'Precipitation (mm)',
       x = element_blank()) +
  guides(color = guide_legend(title = NULL),
         fill = guide_legend(title = NULL)) +
    theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank()
  )

p_cols

ggsave("P.png", plot = p_cols, width = 8, height = 6, dpi = 300)

# Temperature plot
t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
                                      color = site, group = site)) +
  geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size=12) +
  labs(y = expression("Mean temperature " ( degree~C)),
       x = element_blank()) +
  labs(color = NULL, fill = NULL)

# Solar radiation plot
sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
                                         color = site, group = site)) +
  geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size = 12) +
  labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
       x = element_blank()) +
  labs(color = NULL, fill = NULL)

# Relative humidity plot
rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
                                        color = site, group = site)) +
  geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size = 12) +
  labs(y = 'RH (%)', x= element_blank()) +
  labs(color = NULL, fill = NULL)

# Print the plots
t_moplot
sol_moplot
rh_moplot

# t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   theme_bw(base_size=14) +
#   labs(y = expression("Mean temperature " ( degree~C)),
#        x = element_blank())
# 
# t_moplot
# 
# sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   theme_bw(base_size = 14) +
#   labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
#        x = element_blank()) 
# 
# sol_moplot
# 
# 
# rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
#                                       color = site, group = site)) +
#   geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.2) +
#   geom_line() +
#   geom_point() +
#   theme_bw(base_size = 14) +
#   labs(y = 'RH (%)', x= element_blank()) 
# rh_moplot
library(ggpubr)

quad <- ggarrange(p_cols, 
          t_moplot,
          sol_moplot,
          rh_moplot,
          common.legend = T,
          legend = 'bottom',
          ncol=2,
          nrow=2) 
          #nrow = 4)
          #ncol= 0)
quad

ggsave("./figures/quad.png", plot = quad, dpi = 600)

quad_plus <- ggarrange(flow_plot,quad,
                       nrow=2)
quad_plus

ggsave("./figures/quad_plus.png", plot = quad_plus, dpi = 600)

```

# difference - UPDATE file sources!!

```{r}

#grid first
setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/HRU_output/grid")

read_ages_grid <- function(dataframe){
  read_csv(dataframe, skip = 3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'grid')

}

#pull in all filenames with HRU in the name
filenames <- list.files(".", pattern="HRU", full.names=F)

#read all the csvs using read_ages function
dataframes <- lapply(filenames, read_ages_grid)

#name all the csvs based on first 6 characters
names(dataframes) <- substr(filenames, 1, 6)

#bind all and make 1 df
hru_grids <- dataframes %>%
  bind_rows()

#statn next
setwd("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/HRU_output/statn")

read_ages_statn <- function(dataframe){
  read_csv(dataframe, skip = 3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'statn')

}

#pull in all filenames with HRU in the name
filenames <- list.files(".", pattern="HRU", full.names=F)

#read all the csvs using read_ages function
dataframes <- lapply(filenames, read_ages_statn)

#name all the csvs based on first 6 characters
names(dataframes) <- substr(filenames, 1, 6)

#bind all and make 1 df
hru_statn <- dataframes %>%
  bind_rows()

#get rid of a bunch of data
hru_grid_tmean <- hru_grids %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_grid = tmean)  %>%
  group_by(ID) %>%
  summarize(tmean_all_grid = mean(tmean_grid))

hru_statn_tmean <- hru_statn %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_statn = tmean) %>%
  group_by(ID) %>%
  summarize(tmean_all_statn = mean(tmean_statn))

tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('ID')) %>%
  mutate(diff = tmean_all_statn - tmean_all_grid) %>%
  rename(value = ID)

hrus_map_tmean <- left_join(hru_new, tmean_hrus, by = 'value')

# T max
#get rid of a bunch of data
hru_grid_tmax <- hru_grids %>%
  dplyr::select(c(time, ID, tmax)) %>%
  rename(tmax_grid = tmax)  %>%
  group_by(ID) %>%
  summarize(tmax_all_grid = mean(tmax_grid))

hru_statn_tmax <- hru_statn %>%
  dplyr::select(c(time, ID, tmax)) %>%
  rename(tmax_statn = tmax) %>%
  group_by(ID) %>%
  summarize(tmax_all_statn = mean(tmax_statn))

tmax_hrus <- left_join(hru_grid_tmax, hru_statn_tmax, by = c('ID')) %>%
  mutate(diff = tmax_all_statn - tmax_all_grid) %>%
  rename(value = ID)

hrus_map_tmax <- left_join(hru_new, tmax_hrus, by = 'value')

# Tmin
#get rid of a bunch of data
hru_grid_tmin <- hru_grids %>%
  dplyr::select(c(time, ID, tmin)) %>%
  rename(tmin_grid = tmin)  %>%
  group_by(ID) %>%
  summarize(tmin_all_grid = mean(tmin_grid))

hru_statn_tmin <- hru_statn %>%
  dplyr::select(c(time, ID, tmin)) %>%
  rename(tmin_statn = tmin) %>%
  group_by(ID) %>%
  summarize(tmin_all_statn = mean(tmin_statn))

tmin_hrus <- left_join(hru_grid_tmin, hru_statn_tmin, by = c('ID')) %>%
  mutate(diff = tmin_all_statn - tmin_all_grid) %>%
  rename(value = ID)

hrus_map_tmin <- left_join(hru_new, tmin_hrus, by = 'value')

setwd("~/Documents/Repos/fourmile")

# Cumulative p
hru_grid_precip <- hru_grids %>%
  dplyr::select(c(time, ID, precip)) %>%
  rename(precip_grid = precip)

hru_statn_precip <- hru_statn %>%
  dplyr::select(c(time, ID, precip)) %>%
  rename(precip_statn = precip)

precip_hrus <- left_join(hru_grid_precip, hru_statn_precip, by = c('time', 'ID')) %>%
  mutate(diff = precip_statn - precip_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>%
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(precip_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

hrus_map_precip <- left_join(hru_new, precip_hrus, by = 'value')

# solar rad
hru_grid_solrad <- hru_grids %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_grid = solRad)  %>%
  group_by(ID) %>%
  summarize(solrad_all_grid = mean(solrad_grid))

hru_statn_solrad <- hru_statn %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_statn = solRad) %>%
  group_by(ID) %>%
  summarize(solrad_all_statn = mean(solrad_statn))

solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('ID')) %>%
  mutate(diff = solrad_all_statn - solrad_all_grid) %>%
  rename(value = ID)

hrus_map_solrad <- left_join(hru_new, solrad_hrus, by = 'value')

# RH
#get rid of a bunch of data
hru_grid_rh <- hru_grids %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_grid = rhum)  %>%
  group_by(ID) %>%
  summarize(rh_all_grid = mean(rh_grid))

hru_statn_rh <- hru_statn %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_statn = rhum) %>%
  group_by(ID) %>%
  summarize(rh_all_statn = mean(rh_statn))

rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('ID')) %>%
  mutate(diff = rh_all_statn - rh_all_grid) %>%
  rename(value = ID)

hrus_map_rh <- left_join(hru_new, rh_hrus, by = 'value')

####
rh <- ggplot() +
  geom_sf(data = hrus_map_rh, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "viridis", name = "RH (%)") +
  theme_bw(base_size=15.5) +
  theme_void() + bottom_theme

rh

sr <- ggplot() +
  geom_sf(data = hrus_map_solrad, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "cividis", name = expression("Srad (MJ/m"^"2" * ")")) +
  theme_bw(base_size=15.5) +
  theme_void() + top_theme

sr

precip <- ggplot() +
  geom_sf(data = hrus_map_precip, aes(fill = precip_avg_cumdiff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "mako", name = "Cumulative P (mm)") +
  theme_bw(base_size=15.5) +
  theme_void() + top_theme

precip

tmean <- ggplot() +
  geom_sf(data = hrus_map_tmean, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Tmean (°C)",
                       limits = c(-5, 2)) +
  theme_bw(base_size=15.5) +
  theme_void() + bottom_theme

tmean

tmin <- ggplot() +
  geom_sf(data = hrus_map_tmin, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Tmin (°C)",
                       limits = c(-5, 2)) +
  theme_bw(base_size=15.5) +
  theme_void() + bottom_theme

tmin

tmax <- ggplot() +
  geom_sf(data = hrus_map_tmax, aes(fill = diff), color = "black", lwd = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "Tmax (°C)",
                       limits = c(-5, 2)) +
  theme_bw(base_size=15.5) +
  theme_void() + bottom_theme

tmax

bottom_theme <-   theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    legend.key.width = unit(.75, "cm"),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = unit(c(-6, 0, 0, 0), "cm"), 
    panel.spacing = unit(0, "lines"))

top_theme <-   theme(
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    legend.key.width = unit(.75, "cm"),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = unit(c(0, 0, -6, 0), "cm"), 
    panel.spacing = unit(0, "lines"))

lots <- ggarrange(
  precip, sr, rh, tmean,
  heights = 1)

lots

ggsave("./figures/diff_update2.png", plot = lots, dpi = 600)

```

# Pgrid vs Pstatn

```{r}

statn_catch <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/output/4milecreek38HRUs-statnall_#1/out/Catchment.csv',
                        skip=3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'statn')

grid_catch <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_model/output/4milecreek38HRUs-gridall_#1/out/Catchment.csv',
                       skip=3) %>%
  slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(time = ymd(time)) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(source = 'grid')

catchment_p <- bind_rows(statn_catch, grid_catch) %>%
  dplyr::select(precip, time, source) %>%
  pivot_wider(., names_from=source, values_from=precip)

p_comp <- ggplot(catchment_p, aes(x=statn, y=grid)) + geom_point(size = 3) +
  theme_bw(base_size=25) +
  labs(x = 'Station precip (mm)',
       y = 'Grid precip (mm)') +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")

p_comp
ggsave("./figures/p_compare.png", plot = p_comp, dpi = 600)

```

# Obs hydrograph

```{r}

setwd('/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates')

flow <- read_csv('discharge_all.csv') %>%
  replace(.==-9999, NA) %>%
  rename(date = time) %>%
  mutate(date = mdy(date)) %>%
  dplyr::select(date, obs) %>%
  mutate(year = year(date),
         DOY = yday(date)) #%>%
  #drop_na(obs)

library(viridis)

flow_plot <- ggplot(flow, aes(x = DOY, y = obs, color = as.factor(year),
                              linetype = as.factor(year))) + 
  geom_line(size = 1.1) +
  theme_bw(base_size = 20) +
  labs(y = 'Discharge (cms)', color = 'Year') +
  scale_color_viridis_d(option = "mako") +
     theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(color = "black"))

flow_plot

ggsave("./figures/flow_obs_plot.png", dpi = 600)

summary(flow)

flow_plot2 <- ggplot(flow, aes(x = date, y = obs)) + 
  geom_line(size = 1.1) +
  theme_bw(base_size = 20) +
  labs(y = 'Discharge (cms)',
         x = element_blank()) +
   theme(
    legend.position = "none",
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(color = "black"))

flow_plot2

ggsave("/Users/megansears/Documents/Repos/fourmile/figures/flow_obs_plot2.png", dpi = 600, width = 12, height = 4, units = "in")

```

# statn vs grid met

```{r}

grid <- read_csv("/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/grid_all_Catchment.csv",
                 skip=3) %>%
   slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(date = ymd(time)) %>%
  dplyr::select(-time) %>%
  mutate_if(is.character, as.numeric) %>% 
  mutate(site = 'grid')

statn <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_sept2013_updates/statn_all_Catchment.csv',
                 skip=3) %>%
   slice(3:n()) %>%
  dplyr::select(-starts_with('@'),
                -1) %>%
  mutate(date = ymd(time)) %>%
  dplyr::select(-time) %>%
  mutate_if(is.character, as.numeric) %>% 
  mutate(site = 'statn')

all <- bind_rows(grid, statn) %>%
  arrange(date)

catchment_mo_mean <- all %>%
  mutate(month = month(date)) %>%
  group_by(site, month) %>%
  summarise(across(tmean:solRad, ~ mean(.x, na.rm = TRUE))) %>%
  dplyr::select(site, month, tmean_mean = tmean, 
                solrad_mean = solRad, rh_mean = rhum)

catchment_mo_sum <- all %>%
  dplyr::select(date, precip, site) %>%
  mutate(month = month(date),
         year = year(date)) %>%
  group_by(month, year, site) %>%
  summarize(p_sum = sum(precip)) %>%
  group_by(month, site) %>%
  summarize(avg_p = mean(p_sum),
          sd_p = sd(p_sum))

catchment_mo_sd <- all %>%
  mutate(month = month(date)) %>%
  group_by(site, month) %>%
  summarise(across(tmean:solRad, ~ sd(.x, na.rm = TRUE))) %>%
  dplyr::select(site, month, tmean_sd = tmean, 
                solrad_sd = solRad, rh_sd = rhum)

catchment_mo <- left_join(catchment_mo_mean, catchment_mo_sd, 
                          by = c('month', 'site'))

catchment_mo <- left_join(catchment_mo, catchment_mo_sum)

catchment_mo2 <- catchment_mo %>%
  mutate(month_name = month.abb[month]) %>%
  mutate(month_name = factor(month_name, 
                             levels = c('Jan', 'Feb', 'Mar',
                                        'Apr', 'May', 'Jun',
                                        'Jul', 'Aug', 'Sep',
                                        'Oct', 'Nov', 'Dec')))

p_cols <- ggplot(catchment_mo2, aes(x = month_name, y = avg_p,
                                    color = site, fill = site)) +
  geom_col(position = 'dodge') +
  geom_errorbar(aes(ymin = 0, ymax = avg_p + sd_p), 
                position = position_dodge(width = 0.6), width = 0.2) +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Rename color legend items
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Rename fill legend items
  theme_bw(base_size = 20) +
  labs(y = 'Precipitation (mm)',
       x = element_blank()) +
  guides(color = guide_legend(title = NULL),
         fill = guide_legend(title = NULL)) +
    theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank()
  )

p_cols

ggsave("/Users/megansears/Documents/Repos/fourmile/figures/precip.png", plot = p_cols, width = 8, height = 6, dpi = 300)

# Temperature plot
t_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = tmean_mean, 
                                      color = site, group = site)) +
  geom_ribbon(aes(ymin = tmean_mean - tmean_sd, ymax = tmean_mean + tmean_sd, fill = site), alpha = 0.4) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size=20) +
  labs(y = expression("Air temperature " ( degree~C)),
       x = element_blank()) +
  labs(color = NULL, fill = NULL) +
      theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank())

# Solar radiation plot
sol_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = solrad_mean, 
                                         color = site, group = site)) +
  geom_ribbon(aes(ymin = solrad_mean - solrad_sd, ymax = solrad_mean + solrad_sd, fill = site), alpha = 0.4) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size = 20) +
  labs(y = expression("Solar radiation (MJ/m"^"2" * ")"),
       x = element_blank()) +
  labs(color = NULL, fill = NULL) +
      theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank())

# Relative humidity plot
rh_moplot <- ggplot(catchment_mo2, aes(x = month_name, y = rh_mean, 
                                        color = site, group = site)) +
  geom_ribbon(aes(ymin = rh_mean - rh_sd, ymax = rh_mean + rh_sd, fill = site), alpha = 0.4) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("blue", "orange"),
                     labels = c("Gridded", "Observed")) +  # Set line color to blue and orange and rename
  scale_fill_manual(values = c("blue", "orange"),
                    labels = c("Gridded", "Observed")) +   # Set ribbon fill color to blue and orange and rename
  theme_bw(base_size = 20) +
  labs(y = 'RH (%)', x= element_blank()) +
  labs(color = NULL, fill = NULL) +
      theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.text = element_blank())

library(ggpubr)

quad <- ggarrange(p_cols, 
          t_moplot,
          sol_moplot,
          rh_moplot,
          common.legend = T,
          legend = 'bottom',
          ncol=2,
          nrow=2) 
          #nrow = 4)
          #ncol= 0)
quad

ggsave("/Users/megansears/Documents/Repos/fourmile/figures/quad.png", plot = quad, dpi = 600)

quad_plus <- ggarrange(flow_plot2,quad,
                       nrow=2)
quad_plus

ggsave("./figures/quad_plus2.png", plot = quad_plus, dpi = 600)

```