### 2025-03-25 burn_sev.Rmd
burn severity and snow persistence calcs

---
title: "burn_sev"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

library(stars)
library(tidyverse)
library(mapview)

```

```{r}

# for CPF only
dnbr <- read_stars('./data/GIS/co4005110538520100906_20100621_20110624_dnbr.tif')

burn_cat <- st_apply(dnbr, 1:2, function(x) case_when(
  x < -251  ~ 6,
  x > -250 & x < -101 ~ 7,
  x > -100 & x < 99 ~ 1,
  x > 100 & x < 269 ~ 2,
  x > 270 & x < 439 ~ 3,
  x > 440 & x < 659 ~ 4,
  x > 600 ~ 5,
))

sites <- st_read('/Users/megansears/Documents/Repos/fourmile/data/GIS/Fourmile Watershed.shp')

# need to add area to sites (only some sites have it calculated)
sites$area_m2 <- st_area(sites)

sites <- sites %>%  
  mutate(area_km2 = as.numeric(area_m2) / 1e6)

burn_cat <- st_transform(burn_cat, st_crs(sites))

burn_cat_poly <- st_as_sf(burn_cat, as_points = F, merge=F)

mapview(burn_cat_poly)

intersect <- st_intersection(burn_cat_poly, sites)

mapview(intersect)

intersect$area_m2 <- st_area(intersect)

intersect_summ <- intersect %>%
  group_by(co4005110538520100906_20100621_20110624_dnbr.tif) %>%
  summarize(sum_area = sum(area_m2)) %>%
  mutate(area_km2 = as.numeric(sum_area) / 1e6) %>%
  as.data.frame() %>% 
  dplyr::select(-geometry)

percents <- intersect_summ %>%
  mutate(frac_catchment = area_km2/67.2218) %>% 
  mutate(actual_percent = frac_catchment*100)

write_csv(intersect_summ, './data/final_model_inputs/cpf_burncat_percent.csv')

```

# snow persistence

```{r}

sp <- read_stars('/Users/megansears/Documents/Repos/fourmile/data/GIS/SP_mean.tif')

sites <- st_transform(sites, st_crs(sp))

sp <- st_crop(sp, sites)

mapview(sp) + mapview(sites)

sp <- sp %>%
  as.data.frame() %>%
  drop_na()

SP_cats <- sp %>%
  mutate(sp_ct = case_when(
  SP_mean.tif < 50 ~ "Int",
  SP_mean.tif < 75 ~ 'Trans',
  SP_mean.tif >= 75 ~ "Persis"
  )) %>%
  group_by(sp_ct) %>%
  summarize(count = n()) %>%
  mutate(frac = count/316)

```