### 2025-04-04 flow_figs.Rmd
prepping for hydrograph fig

---
title: "hydrograph figures"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(plotly)

```

```{r}

flow <- read_csv('./38HRU_dec2024_updates/discharge_all.csv') %>%
  replace(.==-9999, NA) %>%
  rename(date = time) %>%
  mutate(date = mdy(date))

flow_pivot <- flow %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date))) %>%
  filter(cal_run %in% c('grid_all',
                        'statn_all',
                        'obs'))

flowEg <- flow_pivot %>%
  filter(year %in% c(2012, 2016, 2017)) %>%
  ggplot(., aes(x=date,
                y=q_m3s, 
                color = cal_run)) +
  geom_line() +
  facet_wrap(~year,
             scales = 'free_x') +
  theme_bw(base_size=20)

flowEg

```