### 2025-04-07 flow_figs.Rmd
added a 2016 hydrograph

---
title: "hydrograph figures"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(plotly)

```

```{r}

flow <- read_csv('./38HRU_dec2024_updates/discharge_all.csv') %>%
  replace(.==-9999, NA) %>%
  rename(date = time) %>%
  mutate(date = mdy(date)) %>%
  drop_na()

flow_pivot <- flow %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date))) %>%
  filter(cal_run %in% c('grid_all',
                        'statn_all',
                        'obs')) %>%
  mutate(cal_run = case_match(cal_run,
                            'grid_all' ~ 'Gridded',
                            'statn_all' ~ 'Station',
                            'obs' ~ 'Observed'))

flowEg <- flow_pivot %>%
  filter(year %in% c(2016)) %>%
  ggplot(aes(x = date,
             y = q_m3s, 
             color = cal_run,
             linetype = cal_run)) + 
  geom_line(size = 1, alpha = 1) + 
  theme_bw(base_size = 20) +
  scale_color_manual(values = c("Gridded" = "#0047AB", 
                                "Observed" = "black", 
                                "Station" = "orange")) +
  scale_linetype_manual(values = c("Gridded" = "solid",
                                   "Observed" = "dashed", 
                                   "Station" = "solid")) +
  labs(x = element_blank(),
       y = "Streamflow (cms)") +
   scale_x_date(date_labels = "%b",
               date_breaks = "1 month") +
  theme(
    legend.position = c(0.9, 0.9),
    legend.text = element_text(size = 16),
    legend.key.width = unit(2, "lines"),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black")
  )

flowEg

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/2016_flowEg.png',
       dpi=600,
       width = 12,
       height = 8)

```