### 2025-10-22 StatnPrecip.Rmd
edited statn precip: no TB !6:9, proper prism shift for !6:9, updated TBs for 2016-2020 to be proper sum

---
title: "precip_statn"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)

```

# Precip gauges
GH, LM, SE, SP, ST were missing 2016-2020 so that is in next chunk

```{r}

prec <- read_csv('/Users/megansears/Documents/Repos/fourmile/data/Archive/files_to_edit/precipAll_statn.csv')

#daily precip
precall <- prec %>%
  mutate(date = dmy(date),
         doy = yday(date)) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  filter(year > 2009) %>% 
  mutate_at(c('Gordon_South', 'Gordon_North',
              'GH', 'LM', 'SE', 'SP', 'ST', 'NADP_SL'),
            ~(ifelse(!month %in% c(6:9),
                               NA, .))) # make NA if not in 6:9 month

# need to add betasso
bet_rain <- bet_met %>%
  dplyr::select(c(5,6)) %>%
  rename(Betasso_rain = 1) %>%
  mutate(date = as.Date(DATE_TIME)) %>%
  group_by(date) %>%
  summarize(Betasso = sum(Betasso_rain)) %>%
  filter(date > '2009-12-31') %>%
  mutate(month = month(date)) %>%
  mutate(Betasso = if_else(!month %in% 6:9, NA_real_, Betasso)) # make NA if not in 6:9 month

precall <- full_join(precall, bet_rain, by = 'date')
# precall includes all the precip gauges now
# it also has filtered the TBs out for months !6:9
# but the TBs are missing 2016 - 2020 data which is in updated_precip_2016to2020.xlsx

# formatting to get ready for Ages
precallFormat <- precall %>%
  select(-c(doy, year, month.x, month.y)) %>%
  mutate(across(everything(), ~ replace_na(., as.integer(-9999)))) %>%
  mutate(date = format(date, '%d.%m.%Y'))

write.csv(precallFormat, 
          '/Users/megansears/Documents/Repos/fourmile/data/Archive/files_to_edit/precipStatn_noWinterTB_TBmissing16&20.csv')

```

# TBs that were missing 2016-2020

```{r}

missingP <- read_csv('./data/Archive/files_to_edit/updated_precip_2016to20_mm.csv')

# for TBs: "(two records a day if no precipitation, but additional data if there was precipitation- one record for each tip of the tipping bucket)"
# above is from Sheila Murphy in the UCFCD directions (saved in this folder)

gh <- missingP %>%
  select(date = GH_date, GH) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  group_by(date) %>% # group by date since several tips per day can happen
  summarize(GH = sum(GH)) %>% # sum by day
  filter(date < '2021-01-01') %>% # filter anything later than 1 Jan 2021
  complete(date =
             seq.Date(min(date), 
                      max(ymd('2020-12-31')), 
                      by = "day")) # fill in missing dates

lm <- missingP %>%
  select(date = LM_date, LM) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  group_by(date) %>%
  summarize(LM = sum(LM)) %>%
  filter(date < '2021-01-01') %>%
  complete(date =
             seq.Date(min(date), 
                      max(ymd('2020-12-31')), 
                      by = "day"))

se <- missingP %>%
  select(date = SE_date, SE) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  group_by(date) %>%
  summarize(SE = sum(SE)) %>%
  filter(date < '2021-01-01') %>%
  complete(date =
             seq.Date(min(date), 
                      max(ymd('2020-12-31')), 
                      by = "day"))

sp <- missingP %>%
  select(date = SP_date, SP) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  group_by(date) %>%
  summarize(SP = sum(SP)) %>%
  filter(date < '2021-01-01') %>%
  complete(date =
             seq.Date(min(date), 
                      max(ymd('2020-12-31')), 
                      by = "day"))

st <- missingP %>%
  select(date = ST_date, ST) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  group_by(date) %>%
  summarize(ST = sum(ST)) %>%
  filter(date < '2021-01-01') %>%
  complete(date =
             seq.Date(min(date), 
                      max(ymd('2020-12-31')), 
                      by = "day"))

# combine as all TBs as 1 df
dfs <- list(gh,
            lm,
            se,
            sp,
            st)

# Perform left joins in sequence
tbAll <- reduce(dfs, 
                left_join, 
                by = "date")

# tbAll is GH,LM,SE,SP,ST tipping buckets for 2016-2020
# after looking at a summary of tbAll the following TBs have too high readings randomly
# GH, SP ST
# LM and and SE maxes were 26 and 27 mm per day
# GH had two readings at 949 and 403 - the next smallest max was 29.5 so filtered it by 30
# SP had one reading at 1208, next smallest max was 25.4 so filtered by 26
# ST had many readings greater than 30 (some very large), if there was no precip at any of the other gauges
# and it was larger than ~25 then I assumed it was not real so used 24.5 as filter 
# since the next max that also occurred when other gauges had readings

tbAllFormat <- tbAll %>%
  mutate(GH = if_else(GH > 30.0, NA, GH),
         SP = if_else(SP > 26.0, NA, SP),
         ST = if_else(ST > 24.5, NA, ST))

# now start to format for Ages inputs
# -9999 for NA
# date should be dd.mm.yyyy
tbAllFormat <- tbAllFormat %>%
  mutate(across(everything(), ~ replace_na(., as.integer(-9999)))) %>%
  mutate(month = month(date)) %>%
  mutate(
    across(2:6, ~ if_else(!month %in% 6:9, -9999, .))) %>%
  select(-month) %>% 
  mutate(date = format(date, '%d.%m.%Y'))

write.csv(tbAllFormat,
          '/Users/megansears/Documents/Repos/fourmile/data/Archive/files_to_edit/tb16to20_formatted.csv')

```

# Prism supplement

```{r}

prismSupp <- read_csv('/Users/megansears/Documents/Repos/fourmile/data/Archive/files_to_edit/prismForStatn.csv') %>%
  mutate(date = dmy(date)) %>%
  mutate(month = month(date)) %>%
  mutate(
    across(2:8, ~ if_else(month %in% 6:9, -9999, .))) %>%
  select(-month) %>%
  mutate(date = format(date, '%d.%m.%Y'))

write.csv(prismSupp,
          '/Users/megansears/Documents/Repos/fourmile/data/Archive/files_to_edit/prismForStatn_formatted.csv')

```